---
title: "lolliplot_geoclone"
author: "dcha"
date: "2026-09-01"
output: html_document
---

---
title: "ampli_seq"
author: "dcha"
date: "2026-07-27"
output: html_document
---

```{r}
library(ggplot2)
library(dplyr)
library(readr)
library(tibble)
```



### Remove samples that have low depth and heart samples
```{r}
# Load final_summary.vcf as 'raw'
raw<-read_tsv(file="../raw_data/ampli_benchmark_n650_s14.tsv")
```

```{r}
see_read_depth_sample <- function(vcf) {

  plot_data <- vcf %>%
    dplyr::mutate(
      total_depth = REF_COUNT + ALT_COUNT,
      log10_depth = log10(total_depth)
    ) %>%
    dplyr::filter(is.finite(log10_depth))

  summary_data <- plot_data %>%
    dplyr::group_by(ID) %>%
    dplyr::summarise(
      median = median(total_depth, na.rm = TRUE),
      mean = mean(total_depth, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    tidyr::pivot_longer(
      cols = c(median, mean),
      names_to = "statistic",
      values_to = "depth"
    ) %>%
    dplyr::mutate(
      log10_depth = log10(depth),
      label = sprintf("%.0f", depth),
      label_y = dplyr::if_else(
        statistic == "median",
        log10_depth - 0.13,
        log10_depth + 0.13
      )
    )

  plot_data %>%
    ggplot2::ggplot(
      ggplot2::aes(
        x = ID,
        y = log10_depth
      )
    ) +
    ggplot2::geom_violin() +
    ggplot2::geom_point(
      data = summary_data,
      ggplot2::aes(color = statistic),
      size = 1.5
    ) +
    ggplot2::geom_text(
      data = summary_data,
      ggplot2::aes(
        y = label_y,
        label = label,
        color = statistic
      ),
      size = 3.0,
      show.legend = FALSE
    ) +
    ggplot2::geom_hline(
      yintercept = 3,
      linetype = 2,
      color = "blue"
    ) +
    ggplot2::scale_color_manual(
      name = NULL,
      values = c(
        median = "red",
        mean = "blue"
      ),
      breaks = c("median", "mean")
    ) +
    ggplot2::labs(
      x = NULL,
      y = expression(log[10]("Read depth"))
    ) +
    ggplot2::theme_bw() +
    ggplot2::theme(
      axis.text.x = ggplot2::element_text(
        face = "bold",
        size = 8,
        angle = 90,
        hjust = 1
      ),
      legend.position = "top"
    )
}

depth_plot <- see_read_depth_sample(raw)
depth_plot
```



#### Here we go with proper filtering
```{r}
library(ggplot2)
library(dplyr)
library(readr)
library(tibble)

proper_filter <- function(data) {
  data %>%
    dplyr::filter(
      SEGDUP == FALSE,
      HOMOPOLYMER == FALSE,
      DINUCLEOTIDE == FALSE,
      NEAR_INDEL == FALSE,
      SIMPLE_REPEAT == 0
    )
}

annot_simple <- read_tsv("../raw_data/ampli_raw_annot.simple_repeat.tsv")
no_jgg2 <- annot_simple %>% dplyr::filter(ID!="JGG-control")
no_jgg_good_variants2 <- no_jgg2 %>% proper_filter(.)

source_count_df2 <- no_jgg_good_variants2 %>%
  dplyr::count(SOURCE, name = "num") %>%
  dplyr::mutate(
    num = num / 13
  ) %>%
  dplyr::rename(
    source = SOURCE
  )

```

### Germline het curation

### Upper VAF threshold by germline het candidates
```{r}
het_plot_data <- no_jgg_good_variants2 %>%
  dplyr::filter(
    SOURCE == "GermlineHeterozygous",
    !is.na(UPPER_CI)
  ) %>%
  dplyr::mutate(
    tissue = stringr::str_remove(ID, "^7614-"),
    SUBMIT_NAME = forcats::fct_reorder(
      SUBMIT_NAME,
      UPPER_CI,
      .fun = median,
      .desc = FALSE
    )
  )

het_upper_ci_cutoff <- stats::quantile(
  het_plot_data$UPPER_CI,
  probs = 0.025,
  na.rm = TRUE,
  names = FALSE
)

het_upper_ci_plot <- het_plot_data %>%
  ggplot2::ggplot(
    ggplot2::aes(
      x = SUBMIT_NAME,
      y = UPPER_CI,
      color = SUBMIT_NAME
    )
  ) +
  ggplot2::geom_jitter(
    width = 0.15,
    height = 0,
    size = 2,
    alpha = 0.8
  ) +
  ggrepel::geom_text_repel(
    data = het_plot_data %>%
      dplyr::filter(
        UPPER_CI < 0.4 | UPPER_CI > 0.8
      ),
    ggplot2::aes(label = tissue),
    size = 2.5,
    direction = "y",
    box.padding = 0.2,
    point.padding = 0.1,
    min.segment.length = 0,
    max.overlaps = Inf,
    show.legend = FALSE
  ) +
  ggplot2::stat_summary(
    fun = median,
    geom = "point",
    color = "black",
    shape = 18,
    size = 3
  ) +
  ggplot2::geom_hline(
    ggplot2::aes(
      yintercept = 0.4,
      linetype = "UPPER_CI = 0.4"
    ),
    color = "grey30",
    linewidth = 0.8
  ) +
  ggplot2::geom_hline(
    ggplot2::aes(
      yintercept = 0.8,
      linetype = "UPPER_CI = 0.8"
    ),
    color = "grey15",
    linewidth = 0.8
  ) +
  ggplot2::geom_hline(
    ggplot2::aes(
      yintercept = het_upper_ci_cutoff,
      linetype = "Het UPPER_CI 2.5% cutoff"
    ),
    color = "red",
    linewidth = 0.8
  ) +
  ggplot2::scale_linetype_manual(
    name = NULL,
    values = c(
      "UPPER_CI = 0.4" = "dotted",     "UPPER_CI = 0.8" = "dotdash",
      "Het UPPER_CI 2.5% cutoff" = "dashed"
    )
  ) +
  ggplot2::labs(
    x = "Germline heterozygous variant",
    y = "Upper 95% CI"
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    axis.text.x = ggplot2::element_text(
      angle = 90,
      hjust = 1,
      vjust = 0.5,
      size = 7
    ),
    legend.position = "top"
  ) +
  ggplot2::guides(
    color = "none"
  )

het_upper_ci_cutoff
het_upper_ci_plot
```

```{r}
no_occipital <- no_jgg_good_variants2 %>% filter(ID!="7614-R-O") %>% filter(ID!="7614-L-O") %>% filter(SUBMIT_NAME!="chr19_53787466_C_G-het")

```

### Upper VAF threshold by germline het candidates after filtering
```{r}
het_cutoff_result <- no_occipital %>%
  dplyr::filter(
    SOURCE == "GermlineHeterozygous",
    !is.na(MAF),
    !is.na(LOWER_CI),
    !is.na(UPPER_CI)
  ) %>%
  {
    het_data <- .

    somatic_upper_af_cutoff <- stats::quantile(
      het_data$UPPER_CI,
      probs = 0.025,
      na.rm = TRUE,
      names = FALSE
    )

    cutoff_summary <- het_data %>%
      dplyr::summarise(
        n_variants = dplyr::n_distinct(SUBMIT_NAME),
        n_tissue_variant_rows = dplyr::n(),

        maf_q025 = stats::quantile(MAF, 0.025, na.rm = TRUE),
        maf_q975 = stats::quantile(MAF, 0.975, na.rm = TRUE),

        lower_ci_q025 = stats::quantile(LOWER_CI, 0.025, na.rm = TRUE),
        lower_ci_q975 = stats::quantile(LOWER_CI, 0.975, na.rm = TRUE),

        upper_ci_q025 = stats::quantile(UPPER_CI, 0.025, na.rm = TRUE),
        upper_ci_q975 = stats::quantile(UPPER_CI, 0.975, na.rm = TRUE),

        somatic_upper_af_cutoff = somatic_upper_af_cutoff
      )

    cutoff_plot <- het_data %>%
      tidyr::pivot_longer(
        cols = c(MAF, LOWER_CI, UPPER_CI),
        names_to = "metric",
        values_to = "value"
      ) %>%
      ggplot2::ggplot(
        ggplot2::aes(x = value)
      ) +
      ggplot2::geom_histogram(
        bins = 100,
        color = "black",
        fill = "grey80"
      ) +
      ggplot2::facet_wrap(
        ~metric,
        scales = "free_y",
        nrow = 1
      ) +
      ggplot2::geom_vline(
        data = tibble::tibble(
          metric = "UPPER_CI",
          cutoff = somatic_upper_af_cutoff
        ),
        ggplot2::aes(xintercept = cutoff),
        color = "red",
        linetype = 2,
        linewidth = 0.8
      ) +
      ggplot2::labs(
        x = "Allele fraction",
        y = "Count"
      ) +
      ggplot2::theme_bw()

    list(
      het_data = het_data,
      cutoff = cutoff_summary,
      plot = cutoff_plot
    )
  }

het_cutoff_result$cutoff
het_cutoff_result$plot
```

```{r}
hom_cutoff_result <- no_occipital %>%
  dplyr::filter(
    SOURCE == "GermlineRefOrAltHomozygous",
    (REF_COUNT + ALT_COUNT) > 30,
    !is.na(MAF),
    !is.na(LOWER_CI),
    !is.na(UPPER_CI)
  ) %>%
  {
    hom_data <- .

    lower_ci_cutoff_75 <- stats::quantile(
      hom_data$LOWER_CI,
      probs = 0.75,
      na.rm = TRUE,
      names = FALSE
    )

    lower_ci_cutoff_95 <- stats::quantile(
      hom_data$LOWER_CI,
      probs = 0.95,
      na.rm = TRUE,
      names = FALSE
    )

    cutoff_summary <- hom_data %>%
      dplyr::summarise(
        n_variants = dplyr::n_distinct(SUBMIT_NAME),
        n_tissue_variant_rows = dplyr::n(),
        maf_q025 = stats::quantile(MAF, 0.025, na.rm = TRUE),
        maf_q975 = stats::quantile(MAF, 0.975, na.rm = TRUE),
        lower_ci_q025 = stats::quantile(LOWER_CI, 0.025, na.rm = TRUE),
        lower_ci_q75 = lower_ci_cutoff_75,
        lower_ci_q95 = lower_ci_cutoff_95,
        lower_ci_q975 = stats::quantile(LOWER_CI, 0.975, na.rm = TRUE),
        upper_ci_q025 = stats::quantile(UPPER_CI, 0.025, na.rm = TRUE),
        upper_ci_q975 = stats::quantile(UPPER_CI, 0.975, na.rm = TRUE),
        somatic_lower_af_cutoff = lower_ci_cutoff_95
      )

    plot_data <- hom_data %>%
      tidyr::pivot_longer(
        cols = c(MAF, LOWER_CI, UPPER_CI),
        names_to = "metric",
        values_to = "value"
      )

    plot75 <- plot_data %>%
      ggplot2::ggplot(
        ggplot2::aes(x = value)
      ) +
      ggplot2::geom_histogram(
        bins = 100,
        color = "black",
        fill = "grey80"
      ) +
      ggplot2::facet_wrap(
        ~metric,
        scales = "free_y",
        nrow = 1
      ) +
      ggplot2::geom_vline(
        data = tibble::tibble(
          metric = "LOWER_CI",
          cutoff = lower_ci_cutoff_75
        ),
        ggplot2::aes(xintercept = cutoff),
        color = "red",
        linetype = 2,
        linewidth = 0.8
      ) +
      ggplot2::labs(
        title = "Homozygous-reference LOWER_CI 75th percentile",
        subtitle = paste0(
          "Cutoff = ",
          format(lower_ci_cutoff_75, scientific = TRUE, digits = 4)
        ),
        x = "Allele fraction",
        y = "Count"
      ) +
      ggplot2::theme_bw()

    plot95 <- plot_data %>%
      ggplot2::ggplot(
        ggplot2::aes(x = value)
      ) +
      ggplot2::geom_histogram(
        bins = 100,
        color = "black",
        fill = "grey80"
      ) +
      ggplot2::facet_wrap(
        ~metric,
        scales = "free_y",
        nrow = 1
      ) +
      ggplot2::geom_vline(
        data = tibble::tibble(
          metric = "LOWER_CI",
          cutoff = lower_ci_cutoff_95
        ),
        ggplot2::aes(xintercept = cutoff),
        color = "red",
        linetype = 2,
        linewidth = 0.8
      ) +
      ggplot2::labs(
        title = "Homozygous-reference LOWER_CI 95th percentile",
        subtitle = paste0(
          "Cutoff = ",
          format(lower_ci_cutoff_95, scientific = TRUE, digits = 4)
        ),
        x = "Allele fraction",
        y = "Count"
      ) +
      ggplot2::theme_bw()

    list(
      hom_data = hom_data,
      cutoff = cutoff_summary,
      plot75 = plot75,
      plot95 = plot95
    )
  }

hom_cutoff_result$cutoff %>%
  dplyr::select(lower_ci_q75, lower_ci_q95)

hom_cutoff_result$plot75
```


### mosaic_cand with lower bound 75th
```{r}
mosaic_cand_lowerAF_ci_75 <- no_occipital %>%
  dplyr::filter(
    !SOURCE %in% c(
      "GermlineHeterozygous",
      "GermlineRefOrAltHomozygous"
    ),
    !is.na(LOWER_CI),
    !is.na(UPPER_CI),
    !is.na(NORMAL_LOWER_CI),
    LOWER_CI > hom_cutoff_result$cutoff$lower_ci_q75,
    UPPER_CI < het_cutoff_result$cutoff$somatic_upper_af_cutoff,
    NORMAL_LOWER_CI < hom_cutoff_result$cutoff$lower_ci_q75
  )

cat(
  "number of rows:",
  nrow(mosaic_cand_lowerAF_ci_75),
  "\n"
)

cat(
  "number of variants:",
  dplyr::n_distinct(mosaic_cand_lowerAF_ci_75$SUBMIT_NAME),
  "\n"
)

mosaic_cand_lowerAF_ci_75 %>%
  dplyr::count(SOURCE, sort = TRUE)

mosaic_cand_lowerAF_ci_75 %>%
  dplyr::select(
    ID,
    CHROM,
    POS,
    REF,
    ALT,
    SOURCE,
    SUBMIT_NAME,
    REF_COUNT,
    ALT_COUNT,
    MAF,
    LOWER_CI,
    UPPER_CI,
    NORMAL_REF_COUNT,
    NORMAL_ALT_COUNT,
    NORMAL_MAF,
    NORMAL_LOWER_CI,
    NORMAL_UPPER_CI
  ) %>%
  head()
```


```{r}
#mosaic_added %>% write_tsv(., "flexible_mosaic_added_ampli.tsv")
```

```{r}
mosaic_added <- read_tsv("../raw_data/flexible_mosaic_added_ampli.tsv")

```

### Lolli-plot

```{r}
make_variant_plot_list <- function(data,
                                   flag_col = "mosaic_lowerAF_ci75",
                                   lower_cutoff = hom_cutoff_result$cutoff$lower_ci_q75,
                                   upper_cutoff = het_cutoff_result$cutoff$somatic_upper_af_cutoff) {

  tissue_levels <- c(
    "L-PF", "L-F", "L-P", "L-T",
    "R-PF", "R-F", "R-P", "R-T",
    "L-K", "R-K", "H"
  )

  upper_cutoff <- as.numeric(upper_cutoff)[1]

  if (!is.null(lower_cutoff)) {
    lower_cutoff <- as.numeric(lower_cutoff)[1]
  }

  plot_data <- data %>%
    dplyr::mutate(
      tissue = stringr::str_remove(ID, "^7614-"),
      tissue = dplyr::recode(
        tissue,
        "L-PFC" = "L-PF",
        "R-PFC" = "R-PF",
        "L-Kid" = "L-K",
        "R-Kid" = "R-K",
        "Heart" = "H"
      ),
      tissue = factor(tissue, levels = tissue_levels)
    ) %>%
    dplyr::filter(!is.na(tissue))

  selected_variants <- plot_data %>%
    dplyr::filter(.data[[flag_col]] == 1L) %>%
    dplyr::distinct(SUBMIT_NAME) %>%
    dplyr::pull(SUBMIT_NAME)

  plot_data <- plot_data %>%
    dplyr::filter(SUBMIT_NAME %in% selected_variants) %>%
    dplyr::mutate(
      point_color = dplyr::if_else(
        .data[[flag_col]] == 1L,
        "Mosaic",
        "Other"
      )
    )

  y_break_values <- c(
    0.01,
    0.05,
    0.1,
    0.2,
    upper_cutoff,
    1
  )

  if (!is.null(lower_cutoff)) {
    y_break_values <- c(
      lower_cutoff,
      y_break_values
    )
  }

  y_break_values <- sort(unique(y_break_values))

  y_break_labels <- vapply(
    y_break_values,
    function(x) {
      if (!is.null(lower_cutoff) && isTRUE(all.equal(x, lower_cutoff))) {
        format(x, scientific = TRUE, digits = 2)
      } else {
        format(x, scientific = FALSE, trim = TRUE)
      }
    },
    character(1)
  )

  plot_list <- plot_data %>%
    split(.$SUBMIT_NAME) %>%
    purrr::imap(
      function(temp, variant_name) {

        variant_source <- unique(temp$SOURCE)

        plot <- ggplot2::ggplot(
          temp,
          ggplot2::aes(
            x = tissue,
            y = sqrt(MAF)
          )
        ) +
          ggplot2::geom_linerange(
            ggplot2::aes(
              ymin = sqrt(LOWER_CI),
              ymax = sqrt(UPPER_CI)
            ),
            color = "grey75",
            linewidth = 1.2
          ) +
          ggplot2::geom_point(
            ggplot2::aes(color = point_color),
            size = 3
          ) +
          ggplot2::geom_hline(
            yintercept = sqrt(upper_cutoff),
            linetype = 2,
            color = "black",
            linewidth = 0.8
          ) +
          ggplot2::scale_color_manual(
            values = c(
              "Mosaic" = "#D55E00",
              "Other" = "#56B4E9"
            ),
            guide = "none"
          ) +
          ggplot2::scale_x_discrete(
            drop = FALSE
          ) +
          ggplot2::scale_y_continuous(
            breaks = sqrt(y_break_values),
            labels = y_break_labels,
            limits = c(0, 1)
          ) +
          ggplot2::labs(
            title = variant_name,
            subtitle = paste(
              variant_source,
              collapse = " | "
            ),
            x = NULL,
            y = "sqrt(AF)"
          ) +
          ggplot2::theme_classic() +
          ggplot2::theme(
            plot.title = ggplot2::element_text(
              face = "bold",
              size = 13
            ),
            plot.subtitle = ggplot2::element_text(
              size = 9
            ),
            axis.text.x = ggplot2::element_text(
              face = "bold",
              size = 11,
              angle = 45,
              hjust = 1
            ),
            axis.text.y = ggplot2::element_text(
              size = 11
            ),
            axis.title.y = ggplot2::element_text(
              size = 14
            )
          )

        if (!is.null(lower_cutoff)) {
          plot <- plot +
            ggplot2::geom_hline(
              yintercept = sqrt(lower_cutoff),
              linetype = 2,
              color = "black",
              linewidth = 0.8
            )
        }

        plot
      }
    )

  plot_list
}
```

```{r}
plot_list_ci75 <- make_variant_plot_list(
  data = mosaic_added,
  flag_col = "mosaic_lowerAF_ci75",
  lower_cutoff = hom_cutoff_result$cutoff$lower_ci_q75,
  upper_cutoff = het_cutoff_result$cutoff$somatic_upper_af_cutoff
)

```

### Pick representative variants from `plot_list_ci75`

For candidate selection, a tissue is considered positive only when
`mosaic_lowerAF_ci75 == 1`, matching the orange `Mosaic` points in the
lolliplots. Raw non-zero MAF values that failed the CI75 rule are not counted
as tissue presence.

The four requested patterns are defined as:

1. **Broad with peripheral organs:** positive in heart and both kidneys; rank
   first by the number of positive tissues and then by the weakest positive
   ALT count.
2. **Bilateral brain only:** positive in the largest number of retained left
   and right cortical regions, with no positive heart or kidney sample.
3. **Left brain only:** positive in all four retained left cortical regions,
   with no positive right-brain, heart, or kidney samples.
4. **Focal brain only:** positive in exactly one brain region and no peripheral
   sample.

```{r}
ci75_tissue_levels <- c(
  "L-PF", "L-F", "L-P", "L-T",
  "R-PF", "R-F", "R-P", "R-T",
  "L-K", "R-K", "H"
)

ci75_left_brain <- c("L-PF", "L-F", "L-P", "L-T")
ci75_right_brain <- c("R-PF", "R-F", "R-P", "R-T")
ci75_brain <- c(ci75_left_brain, ci75_right_brain)
ci75_peripheral <- c("L-K", "R-K", "H")

# Restrict the summary to variants that actually occur in plot_list_ci75.
candidate_summary_ci75 <- mosaic_added %>%
  dplyr::mutate(
    tissue = stringr::str_remove(ID, "^7614-"),
    tissue = dplyr::recode(
      tissue,
      "L-PFC" = "L-PF",
      "R-PFC" = "R-PF",
      "L-Kid" = "L-K",
      "R-Kid" = "R-K",
      "Heart" = "H"
    ),
    ci75_present = mosaic_lowerAF_ci75 == 1L,
    total_depth = REF_COUNT + ALT_COUNT
  ) %>%
  dplyr::filter(
    SUBMIT_NAME %in% names(plot_list_ci75),
    tissue %in% ci75_tissue_levels
  ) %>%
  dplyr::arrange(
    SUBMIT_NAME,
    match(tissue, ci75_tissue_levels)
  ) %>%
  dplyr::group_by(SUBMIT_NAME) %>%
  dplyr::summarise(
    source = paste(sort(unique(SOURCE)), collapse = " | "),
    tier = as.integer(
      stringr::str_extract(
        dplyr::first(SUBMIT_NAME),
        "(?<=-t)[123]$"
      )
    ),
    n_present = sum(ci75_present),
    n_left_brain = sum(ci75_present & tissue %in% ci75_left_brain),
    n_right_brain = sum(ci75_present & tissue %in% ci75_right_brain),
    n_brain = sum(ci75_present & tissue %in% ci75_brain),
    n_peripheral = sum(ci75_present & tissue %in% ci75_peripheral),
    heart_present = any(ci75_present & tissue == "H"),
    left_kidney_present = any(ci75_present & tissue == "L-K"),
    right_kidney_present = any(ci75_present & tissue == "R-K"),
    present_tissues = paste(tissue[ci75_present], collapse = ", "),
    min_present_alt_count = min(ALT_COUNT[ci75_present], na.rm = TRUE),
    min_present_depth = min(total_depth[ci75_present], na.rm = TRUE),
    median_present_maf = median(MAF[ci75_present], na.rm = TRUE),
    min_present_lower_ci = min(LOWER_CI[ci75_present], na.rm = TRUE),
    .groups = "drop"
  )

broad_peripheral_candidates_ci75 <- candidate_summary_ci75 %>%
  dplyr::filter(
    heart_present,
    left_kidney_present,
    right_kidney_present
  ) %>%
  dplyr::arrange(
    dplyr::desc(n_present),
    dplyr::desc(min_present_alt_count),
    dplyr::desc(median_present_maf),
    tier
  )

left_brain_only_candidates_ci75 <- candidate_summary_ci75 %>%
  dplyr::filter(
    n_left_brain == length(ci75_left_brain),
    n_right_brain == 0,
    n_peripheral == 0
  ) %>%
  dplyr::arrange(
    dplyr::desc(min_present_alt_count),
    dplyr::desc(median_present_maf),
    tier
  )

bilateral_brain_only_candidates_ci75 <- candidate_summary_ci75 %>%
  dplyr::filter(
    n_left_brain > 0,
    n_right_brain > 0,
    n_peripheral == 0
  ) %>%
  dplyr::arrange(
    dplyr::desc(n_brain),
    dplyr::desc(min_present_alt_count),
    dplyr::desc(median_present_maf),
    tier
  )

focal_brain_only_candidates_ci75 <- candidate_summary_ci75 %>%
  dplyr::filter(
    n_brain == 1,
    n_peripheral == 0
  ) %>%
  dplyr::arrange(
    dplyr::desc(min_present_alt_count),
    dplyr::desc(median_present_maf),
    tier
  )

# Display several alternatives for each requested pattern.
candidate_shortlist_ci75 <- dplyr::bind_rows(
  broad_peripheral_candidates_ci75 %>%
    dplyr::slice_head(n = 3) %>%
    dplyr::mutate(pattern = "broad_with_heart_and_both_kidneys"),
  bilateral_brain_only_candidates_ci75 %>%
    dplyr::slice_head(n = 5) %>%
    dplyr::mutate(pattern = "bilateral_brain_only"),
  left_brain_only_candidates_ci75 %>%
    dplyr::slice_head(n = 5) %>%
    dplyr::mutate(pattern = "all_left_brain_only"),
  focal_brain_only_candidates_ci75 %>%
    dplyr::slice_head(n = 5) %>%
    dplyr::mutate(pattern = "one_brain_region_only")
) %>%
  dplyr::group_by(pattern) %>%
  dplyr::mutate(candidate_rank = dplyr::row_number()) %>%
  dplyr::ungroup() %>%
  dplyr::select(
    pattern,
    candidate_rank,
    SUBMIT_NAME,
    tier,
    source,
    present_tissues,
    n_present,
    min_present_alt_count,
    min_present_depth,
    median_present_maf,
    min_present_lower_ci
  )

candidate_shortlist_ci75 %>%
  print(n = Inf)
```

`min_present_depth` is retained as a review column because the current
`plot_list_ci75` construction does not impose an explicit tissue-depth filter.
In the current data, all three variants positive in heart and both kidneys have
at least one positive peripheral sample below 100 total reads, so a strict
`min_present_depth >= 100` requirement would return no broad candidate.

The first-ranked entry in each class is used below as the default set of
four representatives. The named vector makes it easy to replace any choice
with another row from `candidate_shortlist_ci75`.

```{r}
stopifnot(
  nrow(broad_peripheral_candidates_ci75) > 0,
  nrow(bilateral_brain_only_candidates_ci75) > 0,
  nrow(left_brain_only_candidates_ci75) > 0,
  nrow(focal_brain_only_candidates_ci75) > 0
)

representative_variants_ci75 <- c(
  broad_with_heart_and_kidneys =
    broad_peripheral_candidates_ci75$SUBMIT_NAME[[1]],
  bilateral_brain_only =
    bilateral_brain_only_candidates_ci75$SUBMIT_NAME[[1]],
  left_brain_only =
    left_brain_only_candidates_ci75$SUBMIT_NAME[[1]],
  focal_brain_only =
    focal_brain_only_candidates_ci75$SUBMIT_NAME[[1]]
)

representative_variants_ci75

# These are the four ggplot objects ready for downstream arrangement/export.
representative_plot_list_ci75 <- plot_list_ci75[
  unname(representative_variants_ci75)
]
names(representative_plot_list_ci75) <- names(representative_variants_ci75)
```

### Additional validated tier-3 and PTA-origin examples

Here, "validated" means that the variant is in `plot_list_ci75` and every
CI75-positive tissue has at least 3 ALT reads. For the PTA-origin shortlist,
`SOURCE` must contain `vcf_PTA`; variants that only have `bam_PTA` support but
were discovered by another method are not classified as PTA-origin variants.

```{r}
tier3_validated_candidates_ci75 <- candidate_summary_ci75 %>%
  dplyr::filter(
    tier == 3,
    min_present_alt_count >= 3
  ) %>%
  dplyr::arrange(
    dplyr::desc(n_present),
    dplyr::desc(min_present_alt_count),
    dplyr::desc(median_present_maf)
  )

pta_validated_candidates_ci75 <- candidate_summary_ci75 %>%
  dplyr::filter(
    stringr::str_detect(source, "vcf_PTA"),
    min_present_alt_count >= 3
  ) %>%
  dplyr::arrange(
    dplyr::desc(n_present),
    dplyr::desc(min_present_alt_count),
    dplyr::desc(median_present_maf),
    tier
  )

additional_candidate_shortlist_ci75 <- dplyr::bind_rows(
  tier3_validated_candidates_ci75 %>%
    dplyr::slice_head(n = 5) %>%
    dplyr::mutate(candidate_type = "tier3_validated"),
  pta_validated_candidates_ci75 %>%
    dplyr::slice_head(n = 5) %>%
    dplyr::mutate(candidate_type = "PTA_origin_validated")
) %>%
  dplyr::group_by(candidate_type) %>%
  dplyr::mutate(candidate_rank = dplyr::row_number()) %>%
  dplyr::ungroup() %>%
  dplyr::select(
    candidate_type,
    candidate_rank,
    SUBMIT_NAME,
    tier,
    source,
    present_tissues,
    n_present,
    min_present_alt_count,
    min_present_depth,
    median_present_maf,
    min_present_lower_ci
  )

additional_candidate_shortlist_ci75 %>%
  print(n = Inf)
```

```{r}
stopifnot(
  nrow(tier3_validated_candidates_ci75) > 0,
  nrow(pta_validated_candidates_ci75) > 0
)

additional_representative_variants_ci75 <- c(
  tier3_validated =
    tier3_validated_candidates_ci75$SUBMIT_NAME[[1]],
  PTA_origin_validated =
    pta_validated_candidates_ci75$SUBMIT_NAME[[1]]
)

additional_representative_variants_ci75

additional_representative_plot_list_ci75 <- plot_list_ci75[
  unname(additional_representative_variants_ci75)
]
names(additional_representative_plot_list_ci75) <- names(
  additional_representative_variants_ci75
)

# Optional combined six-example plot list.
extended_representative_variants_ci75 <- c(
  representative_variants_ci75,
  additional_representative_variants_ci75
)

extended_representative_plot_list_ci75 <- plot_list_ci75[
  unname(extended_representative_variants_ci75)
]
names(extended_representative_plot_list_ci75) <- names(
  extended_representative_variants_ci75
)
```

### Figure 2a-style geoclones for the six representatives

The repository implementation of Figure 2a
(`make_AF_plot_variant()`, together with `make_colors()`, `make_hatches()`,
and `make_boxes()`) draws fixed tissue rectangles, uses red opacity
proportional to `sqrt(MAF)`, and optionally hatches samples below a depth
threshold.  The reusable functions below retain those rules while adapting
the layout to the 11 tissues in this experiment.  A tissue is colored only
when `mosaic_lowerAF_ci75 == 1`; a measured CI75-negative tissue is white and
a missing measurement is grey with an X.

The default `normalized_sqrt` mode uses a shared VAF maximum of 0.15 so low
fractions remain visible across all six panels.  Set
`alpha_mode = "original_alpha"` to reproduce the original code's exact
`sqrt(MAF)` red-on-white opacity.  Low-depth hatching is available through
`show_low_depth_hatching = TRUE`, but is off in the final clean figure.
Panel source labels are joined explicitly from
`ampli_raw_annot.simple_repeat.tsv` through the `annot_simple` object.

```{r fig2a-style-geoclone-functions}
geoclone_tissue_geometry <- function() {
  tibble::tribble(
    ~tissue, ~display_group, ~hemisphere, ~xmin, ~xmax, ~ymin, ~ymax,
    "L-PF", "Prefrontal", "Left",  0.00, 1.22, 4.65, 5.35,
    "R-PF", "Prefrontal", "Right", 1.32, 2.54, 4.65, 5.35,
    "L-F",  "Frontal",    "Left",  0.00, 1.22, 3.70, 4.40,
    "R-F",  "Frontal",    "Right", 1.32, 2.54, 3.70, 4.40,
    "L-P",  "Parietal",   "Left",  0.00, 1.22, 2.75, 3.45,
    "R-P",  "Parietal",   "Right", 1.32, 2.54, 2.75, 3.45,
    "L-T",  "Temporal",   "Left",  0.00, 1.22, 1.80, 2.50,
    "R-T",  "Temporal",   "Right", 1.32, 2.54, 1.80, 2.50,
    "H",    "Heart",      "Midline", 0.00, 2.54, 0.75, 1.45,
    "L-K",  "Kidney",     "Left",  0.00, 1.22, -0.20, 0.50,
    "R-K",  "Kidney",     "Right", 1.32, 2.54, -0.20, 0.50
  )
}

normalize_geoclone_tissue <- function(id) {
  tissue <- stringr::str_remove(id, "^7614-")
  dplyr::recode(
    tissue,
    "L-PFC" = "L-PF",
    "R-PFC" = "R-PF",
    "L-Kid" = "L-K",
    "R-Kid" = "R-K",
    "Heart" = "H"
  )
}

format_geoclone_variant <- function(variant) {
  core <- stringr::str_remove(variant, "-t[123]$")
  tier <- toupper(stringr::str_extract(variant, "t[123]$"))
  pieces <- stringr::str_split(core, "_", simplify = TRUE)
  paste0(
    pieces[, 1], ":", pieces[, 2], " ",
    pieces[, 3], ">", pieces[, 4], " (", tier, ")"
  )
}

format_geoclone_source <- function(source, width = 34) {
  breakable <- source %>%
    stringr::str_remove("^tier[123]-") %>%
    stringr::str_replace_all("-", "- ") %>%
    stringr::str_replace_all("_", "_ ")

  stringr::str_wrap(paste0("Source: ", breakable), width = width) %>%
    stringr::str_replace_all("- ", "-") %>%
    stringr::str_replace_all("_ ", "_")
}

prepare_fig2a_geoclone_data <- function(data,
                                        variants,
                                        source_annotation = data,
                                        flag_col = "mosaic_lowerAF_ci75",
                                        vaf_max = 0.15,
                                        alpha_mode = c(
                                          "normalized_sqrt",
                                          "original_alpha"
                                        )) {
  alpha_mode <- match.arg(alpha_mode)
  required_columns <- c(
    "SUBMIT_NAME", "ID", "MAF", "REF_COUNT", "ALT_COUNT",
    flag_col
  )
  missing_columns <- setdiff(required_columns, names(data))
  if (length(missing_columns) > 0L) {
    stop(
      "Missing required columns: ",
      paste(missing_columns, collapse = ", ")
    )
  }
  missing_source_columns <- setdiff(
    c("SUBMIT_NAME", "SOURCE"),
    names(source_annotation)
  )
  if (length(missing_source_columns) > 0L) {
    stop(
      "Missing source-annotation columns: ",
      paste(missing_source_columns, collapse = ", ")
    )
  }
  if (length(variants) == 0L || anyDuplicated(unname(variants))) {
    stop("`variants` must contain unique variant IDs.")
  }
  if (!is.numeric(vaf_max) || length(vaf_max) != 1L ||
      !is.finite(vaf_max) || vaf_max <= 0) {
    stop("`vaf_max` must be one positive finite number.")
  }

  geometry <- geoclone_tissue_geometry()
  variant_key <- tibble::tibble(
    variant_role = names(variants),
    SUBMIT_NAME = unname(variants),
    panel_order = seq_along(variants)
  )
  if (is.null(names(variants)) || any(names(variants) == "")) {
    variant_key$variant_role <- paste0("variant_", variant_key$panel_order)
  }

  source_key <- source_annotation %>%
    dplyr::filter(SUBMIT_NAME %in% variant_key$SUBMIT_NAME) %>%
    dplyr::distinct(SUBMIT_NAME, SOURCE)

  source_conflicts <- source_key %>%
    dplyr::count(SUBMIT_NAME) %>%
    dplyr::filter(n != 1L)
  missing_source_variants <- setdiff(
    variant_key$SUBMIT_NAME,
    source_key$SUBMIT_NAME
  )
  if (nrow(source_conflicts) > 0L || length(missing_source_variants) > 0L ||
      any(is.na(source_key$SOURCE))) {
    stop(
      "Expected one non-missing SOURCE per variant in source_annotation."
    )
  }

  observed <- data %>%
    dplyr::filter(SUBMIT_NAME %in% variant_key$SUBMIT_NAME) %>%
    dplyr::mutate(tissue = normalize_geoclone_tissue(ID)) %>%
    dplyr::filter(tissue %in% geometry$tissue) %>%
    dplyr::select(
      SUBMIT_NAME,
      tissue,
      ID,
      MAF,
      REF_COUNT,
      ALT_COUNT,
      dplyr::all_of(flag_col)
    )

  duplicate_rows <- observed %>%
    dplyr::count(SUBMIT_NAME, tissue) %>%
    dplyr::filter(n != 1L)
  if (nrow(duplicate_rows) > 0L) {
    stop("Expected exactly one source row per observed variant-tissue pair.")
  }

  expected <- tidyr::crossing(
    SUBMIT_NAME = variant_key$SUBMIT_NAME,
    tissue = geometry$tissue
  )

  result <- expected %>%
    dplyr::left_join(observed, by = c("SUBMIT_NAME", "tissue")) %>%
    dplyr::left_join(variant_key, by = "SUBMIT_NAME") %>%
    dplyr::left_join(source_key, by = "SUBMIT_NAME") %>%
    dplyr::left_join(geometry, by = "tissue") %>%
    dplyr::mutate(
      ci75_flag = .data[[flag_col]],
      total_depth = REF_COUNT + ALT_COUNT,
      missing_measurement = is.na(ID) | is.na(MAF) | is.na(ci75_flag),
      ci75_present = !missing_measurement & ci75_flag == 1L,
      plotted_vaf = dplyr::if_else(ci75_present, pmax(MAF, 0), 0),
      fill_value = dplyr::case_when(
        missing_measurement ~ NA_real_,
        alpha_mode == "normalized_sqrt" ~
          sqrt(pmin(plotted_vaf, vaf_max) / vaf_max),
        TRUE ~ sqrt(pmin(plotted_vaf, 1))
      ),
      alpha_mode = alpha_mode
    ) %>%
    dplyr::arrange(panel_order, match(tissue, geometry$tissue))

  stopifnot(
    nrow(result) == length(variants) * nrow(geometry),
    !anyDuplicated(result[c("SUBMIT_NAME", "tissue")]),
    all(is.finite(result$xmin)),
    all(is.finite(result$ymin)),
    all(result$fill_value[!result$missing_measurement] >= 0),
    all(result$fill_value[!result$missing_measurement] <= 1),
    all(result$fill_value[!result$missing_measurement &
                            !result$ci75_present] == 0)
  )

  result
}

# Generate clipped diagonal line segments inside low-depth tissue rectangles.
# This keeps the optional hatch independent of ggpattern or other new packages.
make_geoclone_hatch_segments <- function(plot_data,
                                         depth_threshold = 100,
                                         spacing = 0.16,
                                         x_shift = 0.30) {
  low_depth <- plot_data %>%
    dplyr::filter(
      !missing_measurement,
      !is.na(total_depth),
      total_depth < depth_threshold
    )

  if (nrow(low_depth) == 0L) {
    return(tibble::tibble())
  }

  segments <- lapply(seq_len(nrow(low_depth)), function(i) {
    row <- low_depth[i, ]
    starts <- seq(row$xmin - x_shift, row$xmax, by = spacing)
    x0 <- pmax(starts, row$xmin)
    x1 <- pmin(starts + x_shift, row$xmax)
    y0 <- row$ymin + (x0 - starts) / x_shift * (row$ymax - row$ymin)
    y1 <- row$ymin + (x1 - starts) / x_shift * (row$ymax - row$ymin)

    tibble::tibble(
      panel_label = row$panel_label,
      x = x0,
      xend = x1,
      y = y0,
      yend = y1
    ) %>%
      dplyr::filter(xend > x, yend > y)
  })

  dplyr::bind_rows(segments)
}

plot_fig2a_geoclones <- function(plot_data,
                                 vaf_max = 0.15,
                                 alpha_mode = c(
                                   "normalized_sqrt",
                                   "original_alpha"
                                 ),
                                 ncol = 3,
                                 show_low_depth_hatching = FALSE,
                                 depth_threshold = 100) {
  alpha_mode <- match.arg(alpha_mode)
  role_labels <- c(
    broad_with_heart_and_kidneys = "Broad + peripheral organs",
    bilateral_brain_only = "Bilateral brain only",
    left_brain_only = "Left-brain only",
    focal_brain_only = "Focal brain only",
    tier3_validated = "Validated tier 3",
    PTA_origin_validated = "Validated PTA-origin"
  )

  panel_key <- plot_data %>%
    dplyr::distinct(panel_order, variant_role, SUBMIT_NAME, SOURCE) %>%
    dplyr::arrange(panel_order) %>%
    dplyr::mutate(
      role_label = dplyr::coalesce(
        unname(role_labels[variant_role]),
        stringr::str_replace_all(variant_role, "_", " ")
      ),
      panel_label = paste0(
        role_label, "\n", format_geoclone_variant(SUBMIT_NAME),
        "\n", format_geoclone_source(SOURCE)
      )
    )

  plot_data <- plot_data %>%
    dplyr::left_join(
      panel_key %>% dplyr::select(SUBMIT_NAME, panel_label),
      by = "SUBMIT_NAME"
    ) %>%
    dplyr::mutate(
      panel_label = factor(
        panel_label,
        levels = panel_key$panel_label
      )
    )

  annotations <- panel_key %>%
    tidyr::crossing(
      header = c("Left", "Right"),
      x = c(0.61, 1.93)
    ) %>%
    # crossing() above creates four combinations; retain the paired two.
    dplyr::filter(
      (header == "Left" & x == 0.61) |
        (header == "Right" & x == 1.93)
    ) %>%
    dplyr::mutate(
      panel_label = factor(panel_label, levels = panel_key$panel_label)
    )

  bracket_data <- panel_key %>%
    dplyr::transmute(
      panel_label = factor(panel_label, levels = panel_key$panel_label),
      x = 2.67,
      xend = 2.67,
      y = 1.80,
      yend = 5.35
    )

  if (alpha_mode == "normalized_sqrt") {
    legend_vaf <- unique(c(0, 0.01, 0.05, 0.10, vaf_max))
    legend_breaks <- sqrt(pmin(legend_vaf, vaf_max) / vaf_max)
    legend_title <- paste0("VAF (sqrt scale; capped at ", vaf_max, ")")
  } else {
    legend_vaf <- c(0, 0.01, 0.05, 0.10, 0.15, 0.50, 1)
    legend_breaks <- sqrt(legend_vaf)
    legend_title <- "VAF (original sqrt-alpha scale)"
  }

  figure <- ggplot2::ggplot(plot_data) +
    ggplot2::geom_rect(
      ggplot2::aes(
        xmin = xmin,
        xmax = xmax,
        ymin = ymin,
        ymax = ymax,
        fill = fill_value
      ),
      color = "#171717",
      linewidth = 0.45
    ) +
    ggplot2::geom_segment(
      data = plot_data %>% dplyr::filter(missing_measurement),
      ggplot2::aes(
        x = xmin + 0.15,
        xend = xmax - 0.15,
        y = ymin + 0.12,
        yend = ymax - 0.12
      ),
      inherit.aes = FALSE,
      color = "#666666",
      linewidth = 0.5
    ) +
    ggplot2::geom_segment(
      data = plot_data %>% dplyr::filter(missing_measurement),
      ggplot2::aes(
        x = xmin + 0.15,
        xend = xmax - 0.15,
        y = ymax - 0.12,
        yend = ymin + 0.12
      ),
      inherit.aes = FALSE,
      color = "#666666",
      linewidth = 0.5
    ) +
    ggplot2::geom_text(
      data = annotations,
      ggplot2::aes(x = x, y = 5.68, label = header),
      inherit.aes = FALSE,
      family = "Arial",
      size = 7 / ggplot2::.pt,
      fontface = "bold"
    ) +
    ggplot2::geom_segment(
      data = bracket_data,
      ggplot2::aes(x = x, xend = xend, y = y, yend = yend),
      inherit.aes = FALSE,
      linewidth = 0.45
    ) +
    ggplot2::geom_segment(
      data = bracket_data,
      ggplot2::aes(
        x = x - 0.14,
        xend = x,
        y = y,
        yend = y
      ),
      inherit.aes = FALSE,
      linewidth = 0.45
    ) +
    ggplot2::geom_segment(
      data = bracket_data,
      ggplot2::aes(
        x = x - 0.14,
        xend = x,
        y = yend,
        yend = yend
      ),
      inherit.aes = FALSE,
      linewidth = 0.45
    ) +
    ggplot2::geom_text(
      data = bracket_data,
      ggplot2::aes(x = x + 0.18, y = 3.58, label = "Cortex"),
      inherit.aes = FALSE,
      angle = 270,
      family = "Arial",
      size = 7 / ggplot2::.pt
    ) +
    ggplot2::facet_wrap(~panel_label, ncol = ncol) +
    ggplot2::scale_fill_gradient(
      name = legend_title,
      low = "white",
      high = "#E84A4A",
      limits = c(0, 1),
      breaks = legend_breaks,
      labels = scales::label_number(accuracy = 0.01)(legend_vaf),
      na.value = "#D9D9D9",
      oob = scales::squish
    ) +
    ggplot2::scale_y_continuous(
      breaks = c(5.00, 4.05, 3.10, 2.15, 1.10, 0.15),
      labels = c(
        "Prefrontal", "Frontal", "Parietal", "Temporal", "Heart", "Kidney"
      )
    ) +
    ggplot2::coord_fixed(
      ratio = 0.84,
      xlim = c(-0.05, 2.98),
      ylim = c(-0.38, 5.90),
      clip = "off"
    ) +
    ggplot2::labs(
      x = NULL,
      y = NULL
    ) +
    ggplot2::guides(
      fill = ggplot2::guide_colorbar(
        title.position = "top",
        title.hjust = 0.5,
        barwidth = grid::unit(8.0, "cm"),
        barheight = grid::unit(0.35, "cm")
      )
    ) +
    ggplot2::theme_void(base_size = 7, base_family = "Arial") +
    ggplot2::theme(
      strip.text = ggplot2::element_text(
        face = "bold",
        size = 7,
        lineheight = 1.05,
        margin = ggplot2::margin(b = 4)
      ),
      axis.text.y = ggplot2::element_text(
        color = "black",
        face = "bold",
        size = 7,
        hjust = 1,
        margin = ggplot2::margin(r = 5)
      ),
      legend.position = "bottom",
      legend.title = ggplot2::element_text(size = 7),
      legend.text = ggplot2::element_text(size = 7),
      panel.spacing = grid::unit(1.1, "lines"),
      plot.margin = ggplot2::margin(8, 12, 8, 8)
    )

  if (isTRUE(show_low_depth_hatching)) {
    hatch_data <- make_geoclone_hatch_segments(
      plot_data,
      depth_threshold = depth_threshold
    )
    if (nrow(hatch_data) > 0L) {
      hatch_data$panel_label <- factor(
        hatch_data$panel_label,
        levels = panel_key$panel_label
      )
      figure <- figure +
        ggplot2::geom_segment(
          data = hatch_data,
          ggplot2::aes(x = x, xend = xend, y = y, yend = yend),
          inherit.aes = FALSE,
          color = "#303030",
          linewidth = 0.22,
          alpha = 0.65
        )
    }
  }

  figure
}

export_fig2a_geoclones <- function(figure,
                                   plot_data,
                                   output_dir,
                                   stem = "fig5g_six_variant_geoclones",
                                   width = 13.5,
                                   height = 8.0,
                                   dpi = 300) {
  dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
  pdf_file <- file.path(output_dir, paste0(stem, ".pdf"))
  png_file <- file.path(output_dir, paste0(stem, "_300dpi.png"))
  tsv_file <- file.path(output_dir, paste0(stem, "_plot_data.tsv"))

  ggplot2::ggsave(
    filename = pdf_file,
    plot = figure,
    device = grDevices::cairo_pdf,
    width = width,
    height = height,
    units = "in"
  )

  ragg::agg_png(
    filename = png_file,
    width = width,
    height = height,
    units = "in",
    res = dpi,
    background = "white"
  )
  print(figure)
  grDevices::dev.off()

  export_columns <- c(
    "panel_order", "variant_role", "SUBMIT_NAME", "tissue",
    "display_group", "hemisphere", "SOURCE", "ID", "MAF",
    "REF_COUNT", "ALT_COUNT", "total_depth", "ci75_flag",
    "ci75_present", "missing_measurement", "plotted_vaf", "fill_value",
    "alpha_mode"
  )
  readr::write_tsv(
    plot_data %>% dplyr::select(dplyr::all_of(export_columns)),
    tsv_file,
    na = "NA"
  )

  stopifnot(
    file.exists(pdf_file),
    file.info(pdf_file)$size > 0,
    file.exists(png_file),
    file.info(png_file)$size > 0,
    file.exists(tsv_file),
    length(readr::read_lines(tsv_file)) == nrow(plot_data) + 1L
  )

  list(pdf = pdf_file, png = png_file, tsv = tsv_file)
}
```

```{r fig2a-style-geoclone-export, fig.width=13.5, fig.height=8}
fig2a_geoclone_vaf_max <- 0.15
fig2a_geoclone_alpha_mode <- "normalized_sqrt"

fig2a_geoclone_data <- prepare_fig2a_geoclone_data(
  data = mosaic_added,
  variants = extended_representative_variants_ci75,
  source_annotation = annot_simple,
  flag_col = "mosaic_lowerAF_ci75",
  vaf_max = fig2a_geoclone_vaf_max,
  alpha_mode = fig2a_geoclone_alpha_mode
)

stopifnot(
  nrow(fig2a_geoclone_data) == 66L,
  dplyr::n_distinct(fig2a_geoclone_data$SUBMIT_NAME) == 6L,
  all(table(fig2a_geoclone_data$SUBMIT_NAME) == 11L),
  setequal(
    fig2a_geoclone_data$SUBMIT_NAME,
    unname(extended_representative_variants_ci75)
  ),
  all(
    fig2a_geoclone_data$fill_value[
      !fig2a_geoclone_data$missing_measurement &
        !fig2a_geoclone_data$ci75_present
    ] == 0
  )
)

fig2a_geoclone_six_variant_plot <- plot_fig2a_geoclones(
  plot_data = fig2a_geoclone_data,
  vaf_max = fig2a_geoclone_vaf_max,
  alpha_mode = fig2a_geoclone_alpha_mode,
  ncol = 3,
  show_low_depth_hatching = FALSE
)

geoclone_notebook_dir <- ".."

fig2a_geoclone_output_files <- export_fig2a_geoclones(
  figure = fig2a_geoclone_six_variant_plot,
  plot_data = fig2a_geoclone_data,
  output_dir = file.path(geoclone_notebook_dir, "output")
)

fig2a_geoclone_output_files
fig2a_geoclone_six_variant_plot
```

### Four pattern variants in one row

This compact version contains only the four tissue-distribution patterns and
omits the validated tier-3 and PTA-origin examples.

```{r fig2a-style-four-pattern-export, fig.width=14, fig.height=5.5}
fig2a_geoclone_four_pattern_data <- prepare_fig2a_geoclone_data(
  data = mosaic_added,
  variants = representative_variants_ci75,
  source_annotation = annot_simple,
  flag_col = "mosaic_lowerAF_ci75",
  vaf_max = fig2a_geoclone_vaf_max,
  alpha_mode = fig2a_geoclone_alpha_mode
)

stopifnot(
  nrow(fig2a_geoclone_four_pattern_data) == 44L,
  dplyr::n_distinct(
    fig2a_geoclone_four_pattern_data$SUBMIT_NAME
  ) == 4L,
  all(table(fig2a_geoclone_four_pattern_data$SUBMIT_NAME) == 11L),
  setequal(
    fig2a_geoclone_four_pattern_data$SUBMIT_NAME,
    unname(representative_variants_ci75)
  ),
  !any(
    fig2a_geoclone_four_pattern_data$variant_role %in%
      c("tier3_validated", "PTA_origin_validated")
  )
)

fig2a_geoclone_four_pattern_plot <- plot_fig2a_geoclones(
  plot_data = fig2a_geoclone_four_pattern_data,
  vaf_max = fig2a_geoclone_vaf_max,
  alpha_mode = fig2a_geoclone_alpha_mode,
  ncol = 4,
  show_low_depth_hatching = FALSE
)

fig2a_geoclone_four_pattern_output_files <- export_fig2a_geoclones(
  figure = fig2a_geoclone_four_pattern_plot,
  plot_data = fig2a_geoclone_four_pattern_data,
  output_dir = file.path(geoclone_notebook_dir, "output"),
  stem = "fig5g_four_pattern_geoclones_one_row",
  width = 14,
  height = 5.5,
  dpi = 300
)

fig2a_geoclone_four_pattern_output_files
fig2a_geoclone_four_pattern_plot
```


```{r}
# Save a named list of ggplot objects into a multi-page PDF
save_plot_list_pdf <- function(plot_list, pdf_file, width = 5, height = 6) {
  grDevices::pdf(pdf_file, width = width, height = height)
  on.exit(grDevices::dev.off(), add = TRUE)

  for (plot_name in names(plot_list)) {
    print(plot_list[[plot_name]])
  }
}

# This optional export uses the submission output directory.
# The existing RUN_LEGACY_VOLUMES_EXPORT switch is retained for compatibility.
if (identical(Sys.getenv("RUN_LEGACY_VOLUMES_EXPORT"), "1")) {
  save_plot_list_pdf(
    plot_list = plot_list_ci75,
    pdf_file = "../output/plot_list_ci75_all_variants.pdf"
  )
} else {
  message("Skipping opt-in all-variant PDF export.")
}

```
