---
title: "bench"
author: "DoHyeonCha"
date: "2026-02-23"
output: html_document
---


```{r}
library(readr)
library(dplyr)

# Input
ampli <- read_tsv("chung_ampli_hg19.snv.hg38.tsv", show_col_types = FALSE)
validated <- read_tsv("validated_chung.tsv", show_col_types = FALSE)

# Variants validated in all three: CTX, L, T
blt_keys <- validated %>%
  dplyr::filter(CTX == 1, L == 1, T == 1) %>%
  pull(variant_list) %>%
  unique()

# Add BLT_validated column to ampli table
ampli_out <- ampli %>%
  mutate(BLT_validated = if_else(HG19_KEY %in% blt_keys, 1L, 0L))

# Save
write_tsv(ampli_out, "chung_ampli_hg19.snv.hg38.with_BLT_validated.tsv")

```


```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)

#raw_old<- read_tsv("raw_duplex_pileup.tsv")
#nogerm_old<-read_tsv("no_germline.no_blacklist.duplex_pileup.tsv")
filt_old <- read_tsv("no_unrelated.no_germline.no_blacklist.duplex_pileup.tsv")

#ext_pileup <- read_tsv("raw_duplex_pileup_extended.tsv")
```

```{r}
 
pta_cols <- raw %>% colnames(.) %>% .[str_detect(., "PTA")]


unrel_cols <- raw %>% colnames(.) %>% .[str_detect(.,"unrelated")]

raw_pta_unrel <-  raw %>% dplyr::select(c("chr","pos","ref", "alt", pta_cols, unrel_cols))
```



```{r}
alt_pta_cols <- ext_pileup %>% colnames(.) %>% .[str_detect(., "PTA")] %>% .[str_detect(., "ALT_COUNT")]

ext_pileup_pta_only <- ext_pileup %>% dplyr::select(c("chr","pos","ref", "alt", alt_pta_cols))
```

```{r}

ext_pileup_pta_only %>% dplyr::select(c("chr","pos","ref", "alt"))
raw_pta_unrel %>% dplyr::select(c("chr","pos","ref", "alt"))
                   
```


```{r}
library(dplyr)
library(stringr)

# --------------------------------------------------
# PTA single-neuron ALT count columns (01-16 only)
# --------------------------------------------------
pta16_alt_cols <- ext_pileup %>%
  colnames() %>%
  .[str_detect(., "^PTA_SINGLE_NEURON_[0-9]{2}_ALT_COUNT$")] %>%
  sort()

stopifnot(length(pta16_alt_cols) == 16)

# --------------------------------------------------
# Recalculate PTA existence by ALT_COUNT >= 1
# --------------------------------------------------
pta_exist_mat <- ext_pileup %>%
  dplyr::select(all_of(pta16_alt_cols)) %>%
  as.data.frame()

pta_exist_mat[is.na(pta_exist_mat)] <- 0

pta_exist_bin <- pta_exist_mat >= 1
pta_exist_n <- rowSums(pta_exist_bin)

pta_exist_which <- apply(pta_exist_bin, 1, function(x) {
  idx <- which(x)
  if (length(idx) == 0) "." else paste(idx, collapse = ",")
})

# --------------------------------------------------
# Overwrite / create the PTA BAM-level columns
# --------------------------------------------------
raw2 <- raw %>%
  mutate(
    existing_any_PTA1_16 = as.integer(pta_exist_n >= 1),
    existing_any_PTA1_16_singleton = as.integer(pta_exist_n == 1),
    existing_any_PTA1_16_multiple = as.integer(pta_exist_n >= 2),
    existing_any_PTA1_16_number = as.integer(pta_exist_n),
    existing_any_PTA1_16_which = pta_exist_which
  )
```

```{r}
library(dplyr)
library(stringr)

# --------------------------------------------------
# Unrelated PTA samples to be redefined by ALT_COUNT >= 1
# --------------------------------------------------
pta_unrel_map <- c(
  "HG19_PTA_6566_1_ALT_COUNT" = "HG19_PTA_6566_1",
  "HG19_PTA_6566_2_ALT_COUNT" = "HG19_PTA_6566_2",
  "HG19_PTA_8420_1_ALT_COUNT" = "HG19_PTA_8420_1",
  "HG19_PTA_8420_2_ALT_COUNT" = "HG19_PTA_8420_2"
)

stopifnot(all(names(pta_unrel_map) %in% colnames(ext_pileup)))

pta_unrel_old_names <- unname(pta_unrel_map)

# --------------------------------------------------
# Helper: clean old unrelated list by removing PTA entries
# --------------------------------------------------
clean_unrelated_which <- function(x, remove_names) {
  if (is.na(x) || x == "." || x == "") return(character(0))
  toks <- strsplit(x, ",", fixed = TRUE)[[1]]
  toks <- toks[toks != ""]
  toks <- toks[!toks %in% remove_names]
  toks
}

# --------------------------------------------------
# Helper: derive platform names from final sample names
# --------------------------------------------------
to_platform <- function(x) {
  if (length(x) == 0) return(character(0))
  
  out <- ifelse(
    str_detect(x, "PTA"),
    "PTA",
    ifelse(
      str_detect(x, "PPMSEQ"),
      "PPMSEQ",
      ifelse(
        str_detect(x, "UDSEQ"),
        "UDSEQ",
        ifelse(
          str_detect(x, "NANOSEQ"),
          "NANOSEQ",
          ifelse(
            str_detect(x, "ILLUMINA"),
            "ILLUMINA",
            NA_character_
          )
        )
      )
    )
  )
  
  unique(out[!is.na(out)])
}

# --------------------------------------------------
# Rebuild unrelated fields
# --------------------------------------------------
new_unrelated_which <- vector("character", nrow(raw))

for (i in seq_len(nrow(raw))) {
  # old unrelated entries except PTA
  base_tokens <- clean_unrelated_which(raw$existing_in_unrelated_which[i], pta_unrel_old_names)
  
  # newly defined PTA unrelated entries from ext_pileup ALT_COUNT >= 1
  new_pta_tokens <- unname(
    pta_unrel_map[
      as.numeric(ext_pileup[i, names(pta_unrel_map), drop = TRUE]) >= 1
    ]
  )
  
  # merge and keep stable order:
  # first old non-PTA tokens, then newly added PTA tokens
  final_tokens <- unique(c(base_tokens, new_pta_tokens))
  
  new_unrelated_which[i] <- if (length(final_tokens) == 0) "." else paste(final_tokens, collapse = ",")
}

new_unrelated_platform <- vapply(new_unrelated_which, function(x) {
  toks <- clean_unrelated_which(x, character(0))
  plats <- to_platform(toks)
  if (length(plats) == 0) "." else paste(plats, collapse = ",")
}, character(1))

new_unrelated_flag <- as.integer(new_unrelated_which != ".")

# --------------------------------------------------
# Overwrite/create columns
# --------------------------------------------------
raw3 <- raw2 %>%
  mutate(
    existing_in_unrelated = new_unrelated_flag,
    existing_in_unrelated_platform = new_unrelated_platform,
    existing_in_unrelated_which = new_unrelated_which
  )

```

```{r}
library(dplyr)

pta_update_cols <- c(
  "existing_any_PTA1_16",
  "existing_any_PTA1_16_singleton",
  "existing_any_PTA1_16_multiple",
  "existing_any_PTA1_16_number",
  "existing_any_PTA1_16_which",
  "existing_in_unrelated",
  "existing_in_unrelated_platform",
  "existing_in_unrelated_which"
)

# Optional safety checks
stopifnot(all(colnames(nogerm) == colnames(raw)))
stopifnot(all(pta_update_cols %in% colnames(raw3)))

# Build lookup table from raw3
raw3_lookup <- raw3 %>%
  dplyr::select(chr, pos, ref, alt, all_of(pta_update_cols))

# Replace the PTA-related BAM evidence fields in nogerm
nogerm3 <- nogerm %>%
  dplyr::select(-all_of(pta_update_cols)) %>%
  left_join(raw3_lookup, by = c("chr", "pos", "ref", "alt")) %>%
  dplyr::select(all_of(colnames(nogerm)))

# Safety checks
stopifnot(nrow(nogerm3) == nrow(nogerm))
stopifnot(all(colnames(nogerm3) == colnames(nogerm)))

# Check whether any key failed to match
n_unmatched <- nogerm3 %>%
  filter(is.na(existing_any_PTA1_16)) %>%
  nrow()

cat("Unmatched rows in nogerm3:", n_unmatched, "\n")

```

```{r}
library(dplyr)

filt3 <- nogerm3 %>%
  filter(existing_in_unrelated == 0)

# checks
cat("nrow(filt):", nrow(filt), "\n")
cat("nrow(filt3):", nrow(filt3), "\n")

filt3$existing_in_unrelated %>% table()

```

```{r}
raw3 %>% write_tsv("raw_duplex_pileup_newPTA.tsv")
nogerm3 %>% write_tsv("no_germline.no_blacklist.duplex_pileup_newPTA.tsv")
filt3 %>% write_tsv("no_unrelated.no_germline.no_blacklist.duplex_pileup_newPTA.tsv")

```

```{r}
library(readr)
library(dplyr)

# Read Chung amplicon table
chung_ampli <- read_tsv("chung_ampli_hg19.snv.hg38.tsv", show_col_types = FALSE)

# Build lookup table by hg19 key
chung_key <- chung_ampli %>%
  transmute(
    chr = CHROM,
    pos = POS,
    ref = REF,
    alt = ALT,
    chung_ampli = CATEGORY
  ) %>%
  distinct(chr, pos, ref, alt, .keep_all = TRUE)

# Safety check for duplicated keys in Chung table
dup_chung <- chung_key %>%
  count(chr, pos, ref, alt) %>%
  filter(n > 1)

cat("Duplicated keys in Chung table:", nrow(dup_chung), "\n")

# Helper
add_chung_ampli <- function(df, lookup) {
  df %>%
    left_join(lookup, by = c("chr", "pos", "ref", "alt")) %>%
    mutate(chung_ampli = ifelse(is.na(chung_ampli), ".", chung_ampli)) %>%
    relocate(chung_ampli, .after = last_col())
}

# Apply
raw4 <- add_chung_ampli(raw, chung_key)
nogerm4 <- add_chung_ampli(nogerm, chung_key)
filt4 <- add_chung_ampli(filt, chung_key)

# Quick summaries
raw4$chung_ampli %>% table()
nogerm4$chung_ampli %>% table()
filt4$chung_ampli %>% table()

# Optional sanity checks
cat("ncol(raw)  ->", ncol(raw),  " / ncol(raw4)  ->", ncol(raw4),  "\n")
cat("ncol(nogerm)->", ncol(nogerm)," / ncol(nogerm4)->", ncol(nogerm4), "\n")
cat("ncol(filt) ->", ncol(filt), " / ncol(filt4) ->", ncol(filt4), "\n")

```
```{r}
raw4 %>% write_tsv("raw_duplex_pileup_newPTA.tsv")
nogerm4 %>% write_tsv("no_germline.no_blacklist.duplex_pileup_newPTA.tsv")
filt4 %>% write_tsv("no_unrelated.no_germline.no_blacklist.duplex_pileup_newPTA.tsv")

```

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)

raw <- read_tsv("raw_duplex_pileup_newPTA.tsv")
nogerm<-read_tsv("no_germline.no_blacklist.duplex_pileup_newPTA.tsv")
filt <- read_tsv("no_unrelated.no_germline.no_blacklist.duplex_pileup_newPTA.tsv")

ext_pileup <- read_tsv("raw_duplex_pileup_extended.tsv")
```

```{r}
bamExistCols <- raw %>% colnames(.) %>% .[str_detect(., "exist")] %>% .[!str_detect(.,"PTA")]
bulkExist_raw <- raw %>% dplyr::select(c("chr", "pos", "ref", "alt", bamExistCols))
```

```{r}
library(stringr)

platform_from_col <- function(x) {
  str_remove(x, "^is_existing_lower_ci_0001_")
}

bam_platforms <- bamExistCols %>%
  .[str_detect(., "^is_existing")] %>%
  platform_from_col() %>%
  unique()

unrelated_bams <- raw$existing_in_unrelated_which %>%
  .[. != "."] %>%                          # remove "."
  str_split(",") %>%
  unlist() %>%
  unique()

```

```{r}
related_bams_alt_count <- ext_pileup %>% dplyr::select(c("chr", "pos", "ref", "alt", paste0(bam_platforms, "_ALT_COUNT")))
unrelated_bams_alt_count <- ext_pileup %>% dplyr::select(c("chr", "pos", "ref", "alt", paste0(unrelated_bams, "_ALT_COUNT")
))

```

```{r}
library(dplyr)

# ALT count column names
alt_cols <- paste0(bam_platforms, "_ALT_COUNT")

# new column names
new_cols <- paste0("is_altge1_", bam_platforms)

# generate
alt_ge1_df <- related_bams_alt_count %>%
  mutate(
    across(
      all_of(alt_cols),
      ~ as.integer(. >= 1),
      .names = "is_altge1_{.col}"
    )
  ) %>%
  # rename columns to match desired format
  rename_with(
    ~ gsub("_ALT_COUNT$", "", .) ,
    starts_with("is_altge1_")
  ) %>%
  dplyr::select(chr, pos, ref, alt, starts_with("is_altge1_"))

alt_ge1_df
```


```{r}
library(dplyr)
library(stringr)
library(purrr)

# -----------------------------
# unrelated non-PTA BAMs to recompute from ALT_COUNT
# -----------------------------
unrelated_nonpta_bams <- unrelated_bams[!str_detect(unrelated_bams, "PTA")]
unrelated_nonpta_alt_cols <- paste0(unrelated_nonpta_bams, "_ALT_COUNT")

# sanity check
stopifnot(all(unrelated_nonpta_alt_cols %in% colnames(ext_pileup)))

# key check
stopifnot(
  identical(raw$chr, ext_pileup$chr),
  identical(raw$pos, ext_pileup$pos),
  identical(raw$ref, ext_pileup$ref),
  identical(raw$alt, ext_pileup$alt)
)

# -----------------------------
# helper: BAM name -> platform
# -----------------------------
bam_to_platform <- function(x) {
  case_when(
    str_detect(x, "PTA") ~ "PTA",
    str_detect(x, "NANOSEQ") ~ "NANOSEQ",
    str_detect(x, "UDSEQ") ~ "UDSEQ",
    str_detect(x, "PPMSEQ") ~ "PPMSEQ",
    str_detect(x, "ILLUMINA") ~ "ILLUMINA",
    TRUE ~ NA_character_
  )
}

# -----------------------------
# helper: keep only PTA tokens from old which
# -----------------------------
extract_pta_tokens <- function(x) {
  if (is.na(x) || x == ".") return(character(0))
  toks <- str_split(x, ",", simplify = FALSE)[[1]]
  toks <- toks[str_detect(toks, "PTA")]
  unique(toks)
}

old_pta_tokens <- map(raw$existing_in_unrelated_which, extract_pta_tokens)

# -----------------------------
# recompute non-PTA unrelated BAM hits from ALT_COUNT >= 1
# -----------------------------
unrelated_alt_mat <- ext_pileup %>%
  dplyr::select(all_of(unrelated_nonpta_alt_cols)) %>%
  as.matrix()

mode(unrelated_alt_mat) <- "numeric"

new_nonpta_tokens <- apply(
  unrelated_alt_mat,
  1,
  function(v) unrelated_nonpta_bams[which(v >= 1)]
)

# -----------------------------
# combine PTA(old) + nonPTA(new)
# -----------------------------
altge1_tokens <- map2(old_pta_tokens, new_nonpta_tokens, function(pta, nonpta) {
  out <- unique(c(pta, nonpta))
  out[out != ""]
})

altge1_in_unrelated_which <- map_chr(altge1_tokens, function(x) {
  if (length(x) == 0) "." else paste(x, collapse = ",")
})

altge1_in_unrelated_platform <- map_chr(altge1_tokens, function(x) {
  if (length(x) == 0) return(".")
  plats <- unique(bam_to_platform(x))
  plats <- plats[!is.na(plats)]
  if (length(plats) == 0) "." else paste(plats, collapse = ",")
})

altge1_in_unrelated <- as.integer(map_int(altge1_tokens, length) > 0)

# -----------------------------
# final table
# -----------------------------
raw_altge1_unrelated <- raw %>%
  mutate(
    altge1_in_unrelated = altge1_in_unrelated,
    altge1_in_unrelated_platform = altge1_in_unrelated_platform,
    altge1_in_unrelated_which = altge1_in_unrelated_which
  )

# preview
raw_altge1_unrelated %>%
  dplyr::select(
    chr, pos, ref, alt,
    existing_in_unrelated,
    existing_in_unrelated_platform,
    existing_in_unrelated_which,
    altge1_in_unrelated,
    altge1_in_unrelated_platform,
    altge1_in_unrelated_which
  ) %>%
  as.data.frame() %>%
  head(10)

```

```{r}
library(dplyr)
library(stringr)

# check keys
stopifnot(
  identical(raw[c("chr","pos","ref","alt")], alt_ge1_df[c("chr","pos","ref","alt")]),
  identical(raw[c("chr","pos","ref","alt")], raw_altge1_unrelated[c("chr","pos","ref","alt")])
)

# columns to append
altge1_cols_related <- alt_ge1_df %>%
  colnames() %>%
  .[str_detect(., "^is_altge1_")]

altge1_cols_unrelated <- raw_altge1_unrelated %>%
  colnames() %>%
  .[str_detect(., "^altge1_in_unrelated")]

# concatenate
raw2 <- bind_cols(
  raw,
  alt_ge1_df %>% dplyr::select(all_of(altge1_cols_related)),
  raw_altge1_unrelated %>% dplyr::select(all_of(altge1_cols_unrelated))
)

raw2
```

```{r}
library(dplyr)

# ---- prepare ampli table ----
ampli_keyed <- ampli_out %>%
  transmute(
    chr = CHROM,
    pos = POS,
    ref = REF,
    alt = ALT,
    chung_ampli = CATEGORY,
    chung_ampli_BLT_mosaic = as.integer(BLT_validated == 1)
  )

# ---- append to raw2 by variant key ----
raw3 <- raw2 %>%
  left_join(ampli_keyed, by = c("chr", "pos", "ref", "alt")) %>%
  mutate(
    chung_ampli = if_else(is.na(chung_ampli), ".", chung_ampli),
    chung_ampli_BLT_mosaic = if_else(is.na(chung_ampli_BLT_mosaic), 0L, chung_ampli_BLT_mosaic)
  )

raw3

```

```{r}
library(dplyr)
library(stringr)

# 가져올 컬럼들
new_cols <- colnames(raw3) %>%
  .[str_detect(., "altge1|chung_ampli")]

# key + new cols만 추출
raw3_sub <- raw3 %>%
  dplyr::select(chr, pos, ref, alt, all_of(new_cols))

# nogerm에 붙이기
nogerm3 <- nogerm %>%
  left_join(raw3_sub, by = c("chr", "pos", "ref", "alt"))

nogerm3

```

```{r}
filt3 <- nogerm3 %>% filter(altge1_in_unrelated ==0)
filt <- nogerm3 %>% filter(existing_in_unrelated ==0)
```

```{r}
raw3 %>% write_tsv("raw_duplex_pileup_newPTA_altge1.tsv")
nogerm3 %>% write_tsv("no_germline.no_blacklist.duplex_pileup_newPTA_altge1.tsv")
filt %>% write_tsv("no_unrelated.no_germline.no_blacklist.duplex_pileup_newPTA_altge1_unrel_lowci.tsv")
filt3 %>% write_tsv("no_unrelated.no_germline.no_blacklist.duplex_pileup_newPTA_altge1_unrel_altge1.tsv")


```



```{r}

call_cols_full_depth <- c(
  "call_Illumina",
  "call_ppmSeq_multiread",
  #"call_ppmSeq_putative_multiread",
  "call_ppmSeq_singleton_HC",
  "call_UDSeq_multiread",
  "call_UDSeq_singleton",
  "call_NanoSeq_multiread",
  "call_NanoSeq_singleton",
  "call_HiDEFseq_double_stranded",
  "call_HiDEFseq_single_stranded",
  "call_PTA_singleton"
)

call_cols_100x <- c(
  "call_Illumina_100x",
  "call_ppmSeq_multiread_100x",
  #"call_ppmSeq_putative_multiread_100x",
  "call_ppmSeq_singleton_HC_100x",
  "call_UDSeq_multiread",
  "call_UDSeq_singleton",
  "call_NanoSeq_multiread",
  "call_NanoSeq_singleton",
  "call_HiDEFseq_double_stranded",
  "call_HiDEFseq_single_stranded",
  "call_PTA_singleton"
)

call_cols_30x <- c(
  "call_Illumina_30x",
  "call_ppmSeq_multiread_30x",
  #"call_ppmSeq_putative_multiread_30x",
  "call_ppmSeq_singleton_HC_30x",
  "call_UDSeq_30x_multiread",
  "call_UDSeq_30x_singleton",
  "call_NanoSeq_multiread",
  "call_NanoSeq_singleton",
  "call_HiDEFseq_double_stranded",
  "call_HiDEFseq_single_stranded",
  "call_PTA_singleton"
)


call_cols_ppmSeq_multiread <- c(
  "call_ppmSeq_multiread",
  "call_ppmSeq_multiread_30x",
  "call_ppmSeq_multiread_100x",
  "call_ppmSeq_multiread_200x"
)

call_cols_ppmSeq_singleton <- c(
  "call_ppmSeq_singleton_HC",
  "call_ppmSeq_singleton_HC_30x",
  "call_ppmSeq_singleton_HC_100x",
  "call_ppmSeq_singleton_HC_200x"
)


call_cols_illumina <- c(
  "call_Illumina",
  "call_Illumina_30x",
  "call_Illumina_100x",
  "call_Illumina_200x"
)


call_cols_full_depth_clonal_with_PTA <- c(
  "call_Illumina",
  "call_ppmSeq_multiread",
  "call_UDSeq_multiread",
  "call_NanoSeq_multiread",
  "call_HiDEFseq_double_stranded",
  "call_PTA_singleton"
)
call_cols_full_depth_clonal <- c(
  "call_Illumina",
  "call_ppmSeq_multiread",
  "call_UDSeq_multiread",
  "call_NanoSeq_multiread",
  "call_HiDEFseq_double_stranded"
)

call_cols_100x_clonal_with_PTA <- c(
  "call_Illumina_100x",
  "call_ppmSeq_multiread_100x",
  "call_UDSeq_multiread",
  "call_NanoSeq_multiread",
  "call_HiDEFseq_double_stranded",
  "call_PTA_singleton"
)


call_cols_100x_clonal <- c(
  "call_Illumina_100x",
  "call_ppmSeq_multiread_100x",
  "call_UDSeq_multiread",
  "call_NanoSeq_multiread",
  "call_HiDEFseq_double_stranded"
)

call_cols_30x_clonal_with_PTA <- c(
  "call_Illumina_30x",
  "call_ppmSeq_multiread_30x",
  "call_UDSeq_30x_multiread",
  "call_NanoSeq_multiread",
  "call_HiDEFseq_double_stranded",
  "call_PTA_singleton"
)

call_cols_30x_clonal <- c(
  "call_Illumina_30x",
  "call_ppmSeq_multiread_30x",
  "call_UDSeq_30x_multiread",
  "call_NanoSeq_multiread",
  "call_HiDEFseq_double_stranded"
)

get_self_exist_cols_full <- function(callset) {

  if (callset == "call_Illumina")
    return("is_existing_lower_ci_0001_ILLUMINA")

  if (callset %in% c("call_ppmSeq_multiread","call_ppmSeq_singleton_HC"))
    return("is_existing_lower_ci_0001_DUPLEX_PPMSEQ")

  if (callset %in% c("call_UDSeq_multiread","call_UDSeq_singleton"))
    return("is_existing_lower_ci_0001_DUPLEX_UDSEQ")

  if (callset %in% c("call_NanoSeq_multiread","call_NanoSeq_singleton"))
    return("is_existing_lower_ci_0001_DUPLEX_NANOSEQ")

  if (callset %in% c("call_HiDEFseq_double_stranded","call_HiDEFseq_single_stranded"))
    return("is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ")

  if (callset == "call_PTA_singleton")
    return("existing_any_PTA1_16")

  character(0)
}



exist_cols_full <- c(
  "is_existing_lower_ci_0001_ILLUMINA",
  "is_existing_lower_ci_0001_DUPLEX_PPMSEQ",
  "is_existing_lower_ci_0001_DUPLEX_UDSEQ",
  "is_existing_lower_ci_0001_DUPLEX_NANOSEQ",
  "is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ",
  "existing_any_PTA1_16"
)

```


```{r}
call_cols_full_depth <- c(
  "call_Illumina",
  "call_ppmSeq_multiread",
  #"call_ppmSeq_putative_multiread",
  "call_ppmSeq_singleton_HC",
  "call_UDSeq_multiread",
  "call_UDSeq_singleton",
  "call_NanoSeq_multiread",
  "call_NanoSeq_singleton",
  "call_HiDEFseq_double_stranded",
  "call_HiDEFseq_single_stranded",
  "call_PTA_singleton"
)

get_self_exist_cols_full <- function(callset) {

  if (callset == "call_Illumina")
    return("is_existing_lower_ci_0001_ILLUMINA")

  if (callset %in% c("call_ppmSeq_multiread","call_ppmSeq_singleton_HC"))
    return("is_existing_lower_ci_0001_DUPLEX_PPMSEQ")

  if (callset %in% c("call_UDSeq_multiread","call_UDSeq_singleton"))
    return("is_existing_lower_ci_0001_DUPLEX_UDSEQ")

  if (callset %in% c("call_NanoSeq_multiread","call_NanoSeq_singleton"))
    return("is_existing_lower_ci_0001_DUPLEX_NANOSEQ")

  if (callset %in% c("call_HiDEFseq_double_stranded","call_HiDEFseq_single_stranded"))
    return("is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ")

  if (callset == "call_PTA_singleton")
    return("existing_any_PTA1_16")

  character(0)
}




exist_cols_full <- c(
  "is_existing_lower_ci_0001_ILLUMINA",
  "is_existing_lower_ci_0001_DUPLEX_PPMSEQ",
  "is_existing_lower_ci_0001_DUPLEX_UDSEQ",
  "is_existing_lower_ci_0001_DUPLEX_NANOSEQ",
  "is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ",
  "existing_any_PTA1_16"
)

get_self_exist_cols_full_altge1 <- function(callset) {

  if (callset == "call_Illumina")
    return("is_altge1_ILLUMINA")

  if (callset %in% c("call_ppmSeq_multiread","call_ppmSeq_singleton_HC"))
    return("is_altge1_DUPLEX_PPMSEQ")

  if (callset %in% c("call_UDSeq_multiread","call_UDSeq_singleton"))
    return("is_altge1_DUPLEX_UDSEQ")

  if (callset %in% c("call_NanoSeq_multiread","call_NanoSeq_singleton"))
    return("is_altge1_DUPLEX_NANOSEQ")

  if (callset %in% c("call_HiDEFseq_double_stranded","call_HiDEFseq_single_stranded"))
    return("is_altge1_DUPLEX_HIDEFSEQ")

  if (callset == "call_PTA_singleton")
    return(c("existing_any_PTA1_16_singleton","existing_any_PTA1_16_multiple"))

  character(0)
}


exist_cols_full_altge1 <- c(
  "is_altge1_ILLUMINA",
  "is_altge1_DUPLEX_PPMSEQ",
  "is_altge1_DUPLEX_UDSEQ",
  "is_altge1_DUPLEX_NANOSEQ",
  "is_altge1_DUPLEX_HIDEFSEQ",
  "existing_any_PTA1_16"
)


prev_ampli_col <- c("chung_ampli")
```

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)

#raw <- read_tsv("raw_duplex_pileup_newPTA_altge1.tsv")
#nogerm<-read_tsv("no_germline.no_blacklist.duplex_pileup_newPTA_altge1.tsv")
#filt <- read_tsv("no_unrelated.no_germline.no_blacklist.duplex_pileup_newPTA_altge1_unrel_lowci.tsv")
#filt_altge1<-read_tsv("no_unrelated.no_germline.no_blacklist.duplex_pileup_newPTA_altge1_unrel_altge1.tsv")
ext_pileup <- read_tsv("raw_duplex_pileup_extended.tsv")

#filt<-read_tsv("no_unrelated.no_germline.no_blacklist.duplex_pileup_newPTA_altge1_unrel_altge1.tsv")
filt <- read_tsv("no_unrelated.no_germline.no_blacklist.duplex_pileup_newPTA_altge1_unrel_lowci.tsv")

#filt_lowci <- read_tsv("no_unrelated.no_germline.no_blacklist.duplex_pileup_newPTA_altge1_unrel_lowci.tsv")
#filt_altge1 <-read_tsv("no_unrelated.no_germline.no_blacklist.duplex_pileup_newPTA_altge1_unrel_altge1.tsv")
#ext_pileup <- read_tsv("raw_duplex_pileup_extended.tsv")
```

```{r}
#ext_pileup %>% dplyr::select(c(chr,pos,ref,alt, ))
```


### Start here


```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)

raw <- read_tsv("raw_duplex_pileup_newPTA_altge1.tsv")
filt <- read_tsv("no_unrelated.no_germline.no_blacklist.duplex_pileup_newPTA_altge1_unrel_lowci.tsv")
```

### Stratified random selection SEED

```{r}
STRATIFICATION_SEED <- 42

```


```{r}
library(dplyr)



call_cols_full_depth <- c(
  "call_Illumina",
  "call_ppmSeq_multiread",
  "call_ppmSeq_singleton_HC",
  "call_UDSeq_multiread",
  "call_UDSeq_singleton",
  "call_NanoSeq_multiread",
  "call_NanoSeq_singleton",
  "call_HiDEFseq_double_stranded",
  "call_HiDEFseq_single_stranded",
  "call_PTA_singleton"
)

get_self_exist_cols_full <- function(callset) {

  if (callset == "call_Illumina")
    return("is_existing_lower_ci_0001_ILLUMINA")

  if (callset %in% c("call_ppmSeq_multiread","call_ppmSeq_singleton_HC"))
    return("is_existing_lower_ci_0001_DUPLEX_PPMSEQ")

  if (callset %in% c("call_UDSeq_multiread","call_UDSeq_singleton"))
    return("is_existing_lower_ci_0001_DUPLEX_UDSEQ")

  if (callset %in% c("call_NanoSeq_multiread","call_NanoSeq_singleton"))
    return("is_existing_lower_ci_0001_DUPLEX_NANOSEQ")

  if (callset %in% c("call_HiDEFseq_double_stranded","call_HiDEFseq_single_stranded"))
    return("is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ")

  if (callset == "call_PTA_singleton")
    return(c("existing_any_PTA1_16"))

  character(0)
}



exist_cols_full <- c(
  "is_existing_lower_ci_0001_ILLUMINA",
  "is_existing_lower_ci_0001_DUPLEX_PPMSEQ",
  "is_existing_lower_ci_0001_DUPLEX_UDSEQ",
  "is_existing_lower_ci_0001_DUPLEX_NANOSEQ",
  "is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ",
  "existing_any_PTA1_16"
)

```



### Start with "filt" get the tier-wise numbers
```{r}
library(dplyr)
library(purrr)
library(tibble)


tiering_df <- filt

tier_counts <- purrr::map_dfr(call_cols_full_depth, function(cs) {
  
  other_call_cols <- setdiff(call_cols_full_depth, cs)
  self_exist_col <- get_self_exist_cols_full(cs)
  other_exist_cols <- setdiff(exist_cols_full, self_exist_col)

  #self_exist_col <- get_self_exist_cols_full_altge1(cs)
  #other_exist_cols <- setdiff(exist_cols_full_altge1, self_exist_col)

  self_called <- tiering_df[[cs]] == 1
  
  other_called_any <- rowSums(as.data.frame(tiering_df[, other_call_cols, drop = FALSE]), na.rm = TRUE) >= 1
  other_called_none <- rowSums(as.data.frame(tiering_df[, other_call_cols, drop = FALSE]), na.rm = TRUE) == 0
  
  other_exist_any <- rowSums(as.data.frame(tiering_df[, other_exist_cols, drop = FALSE]), na.rm = TRUE) >= 1
  other_exist_none <- rowSums(as.data.frame(tiering_df[, other_exist_cols, drop = FALSE]), na.rm = TRUE) == 0
  
  tibble(
    callset = cs,
    tier1 = sum(self_called & other_called_any, na.rm = TRUE),
    tier2 = sum(self_called & other_called_none & other_exist_any, na.rm = TRUE),
    tier3 = sum(self_called & other_called_none & other_exist_none, na.rm = TRUE),
    total_called = sum(self_called, na.rm = TRUE)
  )
})

tier_counts

```


```{r}
library(dplyr)
library(purrr)
library(tibble)


tier_summary <- purrr::map_dfr(call_cols_full_depth, function(cs) {
  
  other_call_cols <- setdiff(call_cols_full_depth, cs)
  #self_exist_col <- get_self_exist_cols_full_altge1(cs)
  self_exist_col <- get_self_exist_cols_full(cs)

  #other_exist_cols <- setdiff(exist_cols_full_altge1, self_exist_col)
  other_exist_cols <- setdiff(exist_cols_full, self_exist_col)
  self_called <- filt[[cs]] == 1
  other_called_n <- rowSums(as.data.frame(filt[, other_call_cols, drop = FALSE]), na.rm = TRUE)
  other_exist_n <- rowSums(as.data.frame(filt[, other_exist_cols, drop = FALSE]), na.rm = TRUE)
  is_chung_mosaic <- filt$chung_ampli_BLT_mosaic == 1
  
  idx_tier1 <- self_called & (other_called_n >= 1)
  idx_tier2 <- self_called & (other_called_n == 0) & (other_exist_n >= 1)
  idx_tier3 <- self_called & (other_called_n == 0) & (other_exist_n == 0)
  
  tibble(
    callset = cs,
    tier = c("tier1", "tier2", "tier3"),
    number = c(
      sum(idx_tier1, na.rm = TRUE),
      sum(idx_tier2, na.rm = TRUE),
      sum(idx_tier3, na.rm = TRUE)
    ),
    chung_et_al_2024 = c(
      sum(idx_tier1 & is_chung_mosaic, na.rm = TRUE),
      sum(idx_tier2 & is_chung_mosaic, na.rm = TRUE),
      sum(idx_tier3 & is_chung_mosaic, na.rm = TRUE)
    ),
    new = c(
      sum(idx_tier1 & !is_chung_mosaic, na.rm = TRUE),
      sum(idx_tier2 & !is_chung_mosaic, na.rm = TRUE),
      sum(idx_tier3 & !is_chung_mosaic, na.rm = TRUE)
    )
  )
})

tier_summary

```

```{r}
library(dplyr)
library(purrr)
library(tibble)


#exist_cols_noPTA <- setdiff(exist_cols_full_altge1, "existing_any_PTA1_16")

exist_cols_noPTA <- setdiff(exist_cols_full, "existing_any_PTA1_16")



get_self_exist_cols_noPTA <- function(callset) {

  if (callset == "call_Illumina")
    return("is_existing_lower_ci_0001_ILLUMINA")

  if (callset %in% c("call_ppmSeq_multiread","call_ppmSeq_singleton_HC"))
    return("is_existing_lower_ci_0001_DUPLEX_PPMSEQ")

  if (callset %in% c("call_UDSeq_multiread","call_UDSeq_singleton"))
    return("is_existing_lower_ci_0001_DUPLEX_UDSEQ")

  if (callset %in% c("call_NanoSeq_multiread","call_NanoSeq_singleton"))
    return("is_existing_lower_ci_0001_DUPLEX_NANOSEQ")

  if (callset %in% c("call_HiDEFseq_double_stranded","call_HiDEFseq_single_stranded"))
    return("is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ")


  character(0)
}


tier_summary_noPTA <- purrr::map_dfr(call_cols_full_depth, function(cs) {
  
  other_call_cols <- setdiff(call_cols_full_depth, cs)
  
  #self_exist_col_full <- get_self_exist_cols_full_altge1(cs)
  #other_exist_cols_full <- setdiff(exist_cols_full_altge1, self_exist_col_full)
  
  self_exist_col_full <- get_self_exist_cols_full(cs)
  other_exist_cols_full <- setdiff(exist_cols_full, self_exist_col_full)
  
  self_exist_col_noPTA <- get_self_exist_cols_noPTA(cs)
  other_exist_cols_noPTA <- setdiff(exist_cols_noPTA, self_exist_col_noPTA)
  
  self_called <- filt[[cs]] == 1
  other_called_n <- rowSums(as.data.frame(filt[, other_call_cols, drop = FALSE]), na.rm = TRUE)
  
  other_exist_n_full <- rowSums(as.data.frame(filt[, other_exist_cols_full, drop = FALSE]), na.rm = TRUE)
  other_exist_n_noPTA <- rowSums(as.data.frame(filt[, other_exist_cols_noPTA, drop = FALSE]), na.rm = TRUE)
  
  is_chung_mosaic <- filt$chung_ampli_BLT_mosaic == 1
  
  idx_tier1 <- self_called & (other_called_n >= 1)
  idx_tier2 <- self_called & (other_called_n == 0) & (other_exist_n_full >= 1)
  idx_tier2_noPTA <- self_called & (other_called_n == 0) & (other_exist_n_noPTA >= 1)
  idx_tier3 <- self_called & (other_called_n == 0) & (other_exist_n_full == 0)
  
  tibble(
    callset = cs,
    tier = c("tier1", "tier2", "tier2_noPTA", "tier3"),
    number = c(
      sum(idx_tier1, na.rm = TRUE),
      sum(idx_tier2, na.rm = TRUE),
      sum(idx_tier2_noPTA, na.rm = TRUE),
      sum(idx_tier3, na.rm = TRUE)
    ),
    chung_et_al_2024 = c(
      sum(idx_tier1 & is_chung_mosaic, na.rm = TRUE),
      sum(idx_tier2 & is_chung_mosaic, na.rm = TRUE),
      sum(idx_tier2_noPTA & is_chung_mosaic, na.rm = TRUE),
      sum(idx_tier3 & is_chung_mosaic, na.rm = TRUE)
    ),
    new = c(
      sum(idx_tier1 & !is_chung_mosaic, na.rm = TRUE),
      sum(idx_tier2 & !is_chung_mosaic, na.rm = TRUE),
      sum(idx_tier2_noPTA & !is_chung_mosaic, na.rm = TRUE),
      sum(idx_tier3 & !is_chung_mosaic, na.rm = TRUE)
    )
  )
})

tier_summary_noPTA
```



### tier1 selection
```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(purrr)
  library(tibble)
})

set.seed(STRATIFICATION_SEED)

# Exclude HiDEFseq single stranded
callsets_use <- setdiff(call_cols_full_depth, "call_HiDEFseq_single_stranded")

# Fixed variant key from user-provided columns
filt2 <- filt %>%
  mutate(
    variant_key = paste(chr, pos, ref, alt, sep = ":")
  )

tier1_candidates <- purrr::map_dfr(callsets_use, function(cs) {
  
  other_call_cols <- setdiff(call_cols_full_depth, cs)
  
  self_called <- filt2[[cs]] == 1
  other_called_n <- rowSums(as.data.frame(filt2[, other_call_cols, drop = FALSE]), na.rm = TRUE)
  
  idx_tier1 <- self_called & (other_called_n >= 1)
  
  filt2[idx_tier1, , drop = FALSE] %>%
    dplyr::filter(chung_ampli_BLT_mosaic == 0) %>%   # <- 추가
    transmute(
      chr,
      pos,
      ref,
      alt,
      variant_key,
      chung_ampli_BLT_mosaic
    ) %>%
    mutate(callset = cs) %>%
    distinct()
})
# Order by smaller available tier1 first
callset_order <- tier1_candidates %>%
  dplyr::count(callset, name = "avail_n") %>%
  arrange(avail_n, callset)

# Greedy sampling without duplicate variants across callsets
selected_keys <- character(0)
selected_list <- vector("list", length = nrow(callset_order))

for (i in seq_len(nrow(callset_order))) {
  cs <- callset_order$callset[i]
  
  cand_cs <- tier1_candidates %>%
    dplyr::filter(callset == cs, !variant_key %in% selected_keys)
  
  take_n <- min(10, nrow(cand_cs))
  
  if (take_n == 0) {
    selected_list[[i]] <- tibble(
      chr = character(0),
      pos = numeric(0),
      ref = character(0),
      alt = character(0),
      variant_key = character(0),
      callset = character(0),
      chung_ampli_BLT_mosaic = numeric(0)
    )
  } else {
    picked <- cand_cs %>%
      slice_sample(n = take_n)
    
    selected_keys <- c(selected_keys, picked$variant_key)
    selected_list[[i]] <- picked
  }
}

tier1_selected <- bind_rows(selected_list) %>%
  left_join(callset_order, by = "callset") %>%
  arrange(avail_n, callset, chr, pos, ref, alt)

tier1_selected_summary <- tier1_selected %>%
  dplyr::count(callset, name = "selected_n") %>%
  right_join(callset_order, by = "callset") %>%
  mutate(selected_n = dplyr::coalesce(selected_n, 0L)) %>%
  arrange(avail_n, callset)


tier1_selected %>%
  dplyr::count(variant_key) %>%
  dplyr::filter(n > 1)

# Add overlap partners from original tier1 candidate universe
tier1_overlap_annot <- tier1_candidates %>%
  group_by(variant_key) %>%
  summarise(
    all_tier1_callsets = paste(sort(unique(callset)), collapse = ","),
    n_tier1_callsets = dplyr::n_distinct(callset),
    .groups = "drop"
  )

tier1_selected_annot <- tier1_selected %>%
  left_join(tier1_overlap_annot, by = "variant_key") %>%
  rowwise() %>%
  mutate(
    overlap_partner_callsets = paste(
      setdiff(strsplit(all_tier1_callsets, ",", fixed = TRUE)[[1]], callset),
      collapse = ","
    ),
    overlap_partner_callsets = ifelse(overlap_partner_callsets == "", NA_character_, overlap_partner_callsets)
  ) %>%
  ungroup() %>%
  dplyr::select(
    chr, pos, ref, alt, variant_key, callset, avail_n,
    n_tier1_callsets, overlap_partner_callsets, chung_ampli_BLT_mosaic
  ) %>%
  arrange(avail_n, callset, chr, pos, ref, alt)

tier1_selected_annot
```





### tier2 stratification considering PTA
```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(purrr)
  library(stringr)
  library(tidyr)
})

call_cols_full_depth <- c(
  "call_Illumina",
  "call_ppmSeq_multiread",
  "call_ppmSeq_singleton_HC",
  "call_UDSeq_multiread",
  "call_UDSeq_singleton",
  "call_NanoSeq_multiread",
  "call_NanoSeq_singleton",
  "call_HiDEFseq_double_stranded",
  "call_HiDEFseq_single_stranded",
  "call_PTA_singleton"
)

exist_pretty_map <- c(
  "is_existing_lower_ci_0001_ILLUMINA" = "ILLUMINA",
  "is_existing_lower_ci_0001_DUPLEX_PPMSEQ" = "PPMSEQ",
  "is_existing_lower_ci_0001_DUPLEX_UDSEQ" = "UDSEQ",
  "is_existing_lower_ci_0001_DUPLEX_NANOSEQ" = "NANOSEQ",
  "is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ" = "HIDEFSEQ",
  "existing_any_PTA1_16" = "PTA"
)

source_levels <- c("ILLUMINA", "PPMSEQ", "UDSEQ", "NANOSEQ", "HIDEFSEQ", "PTA")

tier2_new_source_considerPTA_df <- purrr::map_dfr(call_cols_full_depth, function(cs) {
  
  other_call_cols <- setdiff(call_cols_full_depth, cs)
  #self_exist_col <- get_self_exist_cols_full_altge1(cs)
  #other_exist_cols <- setdiff(exist_cols_full_altge1, self_exist_col)
  
  self_exist_col <- get_self_exist_cols_full(cs)
  other_exist_cols <- setdiff(exist_cols_full, self_exist_col)  
  
  self_called <- filt[[cs]] == 1
  other_called_n <- rowSums(as.data.frame(filt[, other_call_cols, drop = FALSE]), na.rm = TRUE)
  other_exist_mat <- as.data.frame(filt[, other_exist_cols, drop = FALSE])
  other_exist_n <- rowSums(other_exist_mat, na.rm = TRUE)
  
  idx_tier2_new <- self_called &
    (other_called_n == 0) &
    (other_exist_n >= 1) &
    (filt$chung_ampli_BLT_mosaic != 1)
  
  if (sum(idx_tier2_new, na.rm = TRUE) == 0) {
    return(tibble(callset = character(0), source_group = character(0), n = integer(0)))
  }
  
  source_group_vec <- apply(other_exist_mat[idx_tier2_new, , drop = FALSE], 1, function(x) {
    hit_cols <- names(x)[which(x >= 1)]
    hit_sources <- unname(exist_pretty_map[hit_cols]) %>%
      unique() %>%
      .[order(match(., source_levels))]
    
    bulk_sources <- setdiff(hit_sources, "PTA")
    
    if (length(bulk_sources) == 0) {
      "PTA_only"
    } else if (length(bulk_sources) == 1) {
      bulk_sources
    } else {
      "MULTIPLE_BULK"
    }
  })
  
  tibble(
    callset = cs,
    source_group = source_group_vec
  ) %>%
    dplyr::count(callset, source_group, name = "n")
})

group_levels <- c(
  "PTA_only",
  "ILLUMINA",
  "PPMSEQ",
  "UDSEQ",
  "NANOSEQ",
  "HIDEFSEQ",
  "MULTIPLE_BULK"
)

callset_to_source <- c(
  "call_Illumina" = "ILLUMINA",
  "call_ppmSeq_multiread" = "PPMSEQ",
  "call_ppmSeq_singleton_HC" = "PPMSEQ",
  "call_UDSeq_multiread" = "UDSEQ",
  "call_UDSeq_singleton" = "UDSEQ",
  "call_NanoSeq_multiread" = "NANOSEQ",
  "call_NanoSeq_singleton" = "NANOSEQ",
  "call_HiDEFseq_double_stranded" = "HIDEFSEQ",
  "call_HiDEFseq_single_stranded" = "HIDEFSEQ",
  "call_PTA_singleton" = "PTA"
)

allowed_grid <- tidyr::expand_grid(
  callset = call_cols_full_depth,
  source_group = group_levels
) %>%
  mutate(
    self_source = unname(callset_to_source[as.character(callset)])
  ) %>%
  dplyr::filter(source_group != self_source) %>%
  dplyr::select(-self_source)

tier2_new_source_considerPTA_df <- allowed_grid %>%
  left_join(
    tier2_new_source_considerPTA_df %>%
      mutate(
        callset = as.character(callset),
        source_group = as.character(source_group)
      ),
    by = c("callset", "source_group")
  ) %>%
  mutate(
    n = dplyr::coalesce(n, 0L),
    callset = factor(callset, levels = call_cols_full_depth),
    source_group = factor(source_group, levels = group_levels)
  ) %>%
  arrange(callset, source_group)

tier2_new_source_considerPTA_df

```

#### AmpliSeq allocation number decision
```{r}
tier1_num <- tier1_selected_annot %>% nrow(.) # 55
tier2_PTA_only_num <- 120
tier2_remainder_num <- 300
tier3_num <- 220
germ_althom_num <- 25
germ_het_num <- 60


tier2_PTA_only_num_min_quota <- 5
tier2_remainder_num_min_quota <- 10
tier3_min_quota <- 5


total_expected_num <- tier1_num + tier2_PTA_only_num + tier2_remainder_num + tier3_num + germ_althom_num + germ_het_num


total_expected_num
STRATIFICATION_SEED
```

### tier2 number allocation run
```{r}

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
})


strata_tier2 <- tier2_new_source_considerPTA_df %>%
  dplyr::filter(callset != "call_HiDEFseq_single_stranded") %>%
  dplyr::filter(!(callset == "call_PTA_singleton" & source_group == "PTA_only")) %>%
  mutate(PTA_only = source_group == "PTA_only")

make_target_map <- function(a = 100L, b = 200L, c = 5L, d = 10L) {
  tibble(
    PTA_only = c(TRUE, FALSE),
    target_total = c(a, b),
    min_quota = c(c, d)
  )
}

target_map <- make_target_map(tier2_PTA_only_num, tier2_remainder_num, tier2_PTA_only_num_min_quota, tier2_remainder_num_min_quota)

minscale_integer_alloc <- function(x, total_n, min_n) {
  k <- length(x)
  stopifnot(total_n >= k * min_n)
  
  xmin <- min(x)
  xsum <- sum(x)
  
  if (all(x == xmin)) {
    y <- rep(total_n / k, k)
  } else {
    a <- (total_n - min_n * k) / (xsum - k * xmin)
    b <- min_n - a * xmin
    y <- a * x + b
  }
  
  y_floor <- floor(y)
  remainder <- y - y_floor
  n_to_add <- total_n - sum(y_floor)
  
  y_int <- y_floor
  if (n_to_add > 0) {
    add_idx <- order(remainder, decreasing = TRUE)[seq_len(n_to_add)]
    y_int[add_idx] <- y_int[add_idx] + 1L
  }
  
  as.integer(y_int)
}

# cap after integerization, then redistribute leftover to feasible groups
cap_and_redistribute <- function(df_one_pta) {
  # required columns:
  # callset, step1_num, step1_num_int_raw, n_total, target_total
  
  out <- df_one_pta %>%
    mutate(
      capacity = n_total,
      step1_num_int = pmin(step1_num_int_raw, capacity)
    )
  
  deficit <- unique(out$target_total) - sum(out$step1_num_int)
  
  if (deficit < 0) {
    stop("Capped allocation exceeded target_total, which should not happen.")
  }
  
  if (deficit == 0) {
    return(out)
  }
  
  # redistribute leftover one-by-one
  # priority: larger step1_num first, then larger remaining capacity
  while (deficit > 0) {
    out <- out %>%
      mutate(room_left = capacity - step1_num_int)
    
    eligible <- which(out$room_left > 0)
    if (length(eligible) == 0) {
      stop("No eligible group left for redistribution.")
    }
    
    ord <- eligible[order(-out$step1_num[eligible], -out$room_left[eligible])]
    pick <- ord[1]
    
    out$step1_num_int[pick] <- out$step1_num_int[pick] + 1L
    deficit <- deficit - 1L
  }
  
  out %>%
    dplyr::select(-room_left)
}

# 1. Sqrt-based number allocation
step1_alloc <- strata_tier2 %>%
  group_by(callset, PTA_only) %>%
  dplyr::summarise(n_total = sum(n), .groups = "drop") %>%
  left_join(target_map, by = "PTA_only") %>%
  group_by(PTA_only) %>%
  mutate(
    sqrt_n_total = sqrt(n_total),
    step1_num = target_total[1] * sqrt_n_total / sum(sqrt_n_total)
  ) %>%
  ungroup()

# 2. Min-scaling + integerization
step1_alloc <- step1_alloc %>%
  group_by(PTA_only) %>%
  group_modify(~{
    .x %>%
      mutate(
        step1_num_int_raw = minscale_integer_alloc(
          x = step1_num,
          total_n = .$target_total[1],
          min_n = .$min_quota[1]
        )
      )
  }) %>%
  ungroup()

# 3. Capacity capping + redistribution within each PTA_only group
step1_alloc <- step1_alloc %>%
  group_by(PTA_only) %>%
  group_modify(~ cap_and_redistribute(.x)) %>%
  ungroup()

# attach back to strata_tier2
strata_tier2 <- strata_tier2 %>%
  left_join(
    step1_alloc %>%
      dplyr::select(callset, PTA_only, step1_num, step1_num_int),
    by = c("callset", "PTA_only")
  )

alpha <- 0.5

alloc_step2_power_B <- function(n_vec, total_n, alpha = 0.5) {
  n_vec <- as.integer(n_vec)
  total_n <- as.integer(total_n)

  if (any(is.na(n_vec)) || any(n_vec < 0)) {
    stop("n_vec must be non-negative integers.")
  }
  if (length(total_n) != 1 || is.na(total_n) || total_n < 0) {
    stop("total_n must be a single non-negative integer.")
  }
  if (sum(n_vec) < total_n) {
    stop("Requested total_n exceeds available capacity sum(n_vec).")
  }

  # Strategy B: minimum 1 if n >= 1
  alloc <- ifelse(n_vec >= 1L, 1L, 0L)
  alloc <- pmin(alloc, n_vec)

  if (sum(alloc) > total_n) {
    # If even min-1 is too much, give 1 to the largest groups only
    out <- rep(0L, length(n_vec))
    eligible <- which(n_vec >= 1L)
    w <- n_vec[eligible]^alpha
    ord <- eligible[order(-w, -n_vec[eligible])]
    pick <- ord[seq_len(total_n)]
    out[pick] <- 1L
    return(as.integer(out))
  }

  remaining <- total_n - sum(alloc)
  if (remaining == 0) {
    return(as.integer(alloc))
  }

  # Proportional allocation with capacity cap
  while (remaining > 0) {
    cap_left <- n_vec - alloc
    eligible <- which(cap_left > 0)

    if (length(eligible) == 0) {
      stop("No remaining capacity, but quota still left.")
    }

    w <- n_vec[eligible]^alpha
    if (sum(w) == 0) {
      w <- rep(1, length(eligible))
    }

    y <- remaining * w / sum(w)
    y_floor <- floor(y)
    y_rem <- y - y_floor

    add <- y_floor

    # largest remainder
    n_to_add <- remaining - sum(add)
    if (n_to_add > 0) {
      add_idx <- order(y_rem, decreasing = TRUE)[seq_len(n_to_add)]
      add[add_idx] <- add[add_idx] + 1L
    }

    # apply with cap
    add_capped <- pmin(add, cap_left[eligible])
    alloc[eligible] <- alloc[eligible] + add_capped

    allocated_now <- sum(add_capped)
    if (allocated_now == 0) {
      # safety fallback: add 1 to the largest eligible group
      i <- eligible[order(-w, -cap_left[eligible])][1]
      alloc[i] <- alloc[i] + 1L
      remaining <- remaining - 1L
    } else {
      remaining <- total_n - sum(alloc)
    }
  }

  as.integer(alloc)
}

strata_tier2_strategyB_alpha <- strata_tier2 %>%
  group_by(callset, PTA_only) %>%
  group_modify(~{
    quota <- unique(.x$step1_num_int)
    if (length(quota) != 1) {
      stop("Each callset/PTA_only group must have one unique step1_num_int.")
    }

    .x %>%
      mutate(
        step2_num_int = alloc_step2_power_B(
          n_vec = n,
          total_n = quota,
          alpha = alpha
        )
      )
  }) %>%
  ungroup()

checkB_alpha <- strata_tier2_strategyB_alpha %>%
  group_by(callset, PTA_only) %>%
  dplyr::summarise(
    step1_num_int = unique(step1_num_int),
    step2_sum = sum(step2_num_int),
    total_capacity = sum(n),
    any_exceed = any(step2_num_int > n),
    .groups = "drop"
  )


minscale_integer_alloc <- function(x, total_n, min_n) {
  k <- length(x)
  stopifnot(total_n >= k * min_n)
  
  xmin <- min(x)
  xsum <- sum(x)
  
  if (all(x == xmin)) {
    y <- rep(total_n / k, k)
  } else {
    a <- (total_n - min_n * k) / (xsum - k * xmin)
    b <- min_n - a * xmin
    y <- a * x + b
  }
  
  y_floor <- floor(y)
  remainder <- y - y_floor
  n_to_add <- total_n - sum(y_floor)
  
  y_int <- y_floor
  if (n_to_add > 0) {
    add_idx <- order(remainder, decreasing = TRUE)[seq_len(n_to_add)]
    y_int[add_idx] <- y_int[add_idx] + 1L
  }
  
  as.integer(y_int)
}

cap_and_redistribute <- function(df, total_target) {
  out <- df %>%
    mutate(
      capacity = new,
      alloc_int = pmin(alloc_int_raw, capacity)
    )
  
  deficit <- total_target - sum(out$alloc_int)
  
  if (deficit < 0) {
    stop("Capped allocation exceeded total_target.")
  }
  
  if (deficit == 0) {
    return(out)
  }
  
  while (deficit > 0) {
    out <- out %>%
      mutate(room_left = capacity - alloc_int)
    
    eligible <- which(out$room_left > 0)
    if (length(eligible) == 0) {
      stop("No eligible group left for redistribution.")
    }
    
    ord <- eligible[order(-out$alloc_cont[eligible], -out$room_left[eligible])]
    pick <- ord[1]
    
    out$alloc_int[pick] <- out$alloc_int[pick] + 1L
    deficit <- deficit - 1L
  }
  
  out %>%
    dplyr::select(-room_left)
}

tier3_alloc <- tier_summary %>%
  dplyr::filter(callset != "call_HiDEFseq_single_stranded") %>%
  dplyr::filter(tier == "tier3") %>%
  mutate(
    sqrt_new = sqrt(new),
    alloc_cont = tier3_num * sqrt_new / sum(sqrt_new)
  ) %>%
  mutate(
    alloc_int_raw = minscale_integer_alloc(
      x = alloc_cont,
      total_n = tier3_num,
      min_n = tier3_min_quota
    )
  ) %>%
  cap_and_redistribute(total_target = tier3_num) %>%
  arrange(desc(alloc_int), desc(new))

tier3_alloc

tier3_alloc <- tier3_alloc %>%
  dplyr::mutate(
    callset = factor(
      callset,
      levels = levels(strata_tier2_strategyB_alpha$callset)
    )
  ) %>%
  dplyr::arrange(callset)



ampli_alloc_tbl <- strata_tier2_strategyB_alpha %>%
  dplyr::mutate(tier = "tier2") %>%
  dplyr::select(
    tier,
    callset,
    PTA_only,
    source_group,
    n,
    step1_num,
    step1_num_int,
    step2_num_int
  ) %>%
  dplyr::bind_rows(
    tibble::tibble(
      tier = "tier1",
      callset = "call_shared",
      PTA_only = FALSE,
      source_group = ".",
      n = tier1_num,
      step1_num = tier1_num,
      step1_num_int = tier1_num,
      step2_num_int = tier1_num
    ),
    .
  )


ampli_alloc_tbl <- ampli_alloc_tbl %>%
  dplyr::bind_rows(
    tier3_alloc %>%
      dplyr::transmute(
        tier = "tier3",
        callset = as.character(callset),
        PTA_only = FALSE,
        source_group = ".",
        n = new,
        step1_num = alloc_int,
        step1_num_int = alloc_int,
        step2_num_int = alloc_int
      ),
    tibble::tibble(
      tier = "germline_althom",
      callset = "germline_althom",
      PTA_only = FALSE,
      source_group = ".",
      n = germ_althom_num,
      step1_num = germ_althom_num,
      step1_num_int = germ_althom_num,
      step2_num_int = germ_althom_num
    ),
    tibble::tibble(
      tier = "germline_het",
      callset = "germline_het",
      PTA_only = FALSE,
      source_group = ".",
      n = germ_het_num,
      step1_num = germ_het_num,
      step1_num_int = germ_het_num,
      step2_num_int = germ_het_num
    )
  )

ampli_alloc_tbl %>% write_tsv("ampliseq_allocation_numbers_by_tier_and_strata.tsv")

```


### tier2 stratified selection - check numbers
```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(purrr)
  library(stringr)
  library(tidyr)
})


exist_pretty_map <- c(
  "is_existing_lower_ci_0001_ILLUMINA" = "ILLUMINA",
  "is_existing_lower_ci_0001_DUPLEX_PPMSEQ" = "PPMSEQ",
  "is_existing_lower_ci_0001_DUPLEX_UDSEQ" = "UDSEQ",
  "is_existing_lower_ci_0001_DUPLEX_NANOSEQ" = "NANOSEQ",
  "is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ" = "HIDEFSEQ",
  "existing_any_PTA1_16" = "PTA"
)

source_levels <- c("ILLUMINA", "PPMSEQ", "UDSEQ", "NANOSEQ", "HIDEFSEQ", "PTA")

# Row-level tier2 variant table
tier2_variant_tbl <- purrr::map_dfr(call_cols_full_depth, function(cs) {
  
  other_call_cols <- setdiff(call_cols_full_depth, cs)
  
  self_exist_col <- get_self_exist_cols_full(cs)
  other_exist_cols <- setdiff(exist_cols_full, self_exist_col)
  
  self_called <- filt[[cs]] == 1
  other_called_n <- rowSums(as.data.frame(filt[, other_call_cols, drop = FALSE]), na.rm = TRUE)
  other_exist_mat <- as.data.frame(filt[, other_exist_cols, drop = FALSE])
  other_exist_n <- rowSums(other_exist_mat, na.rm = TRUE)
  
  idx_tier2 <- self_called &
    (other_called_n == 0) &
    (other_exist_n >= 1) &
    (filt$chung_ampli_BLT_mosaic != 1)
  
  if (sum(idx_tier2, na.rm = TRUE) == 0) {
    return(tibble(
      chr = character(0),
      pos = numeric(0),
      ref = character(0),
      alt = character(0),
      callset = character(0),
      source_group = character(0),
      PTA_only = logical(0)
    ))
  }
  
  sub_filt <- filt[idx_tier2, , drop = FALSE]
  sub_exist_mat <- other_exist_mat[idx_tier2, , drop = FALSE]
  
  source_group_vec <- apply(sub_exist_mat, 1, function(x) {
    hit_cols <- names(x)[which(x >= 1)]
    hit_sources <- unname(exist_pretty_map[hit_cols]) %>%
      unique() %>%
      .[order(match(., source_levels))]
    
    bulk_sources <- setdiff(hit_sources, "PTA")
    
    if (length(bulk_sources) == 0) {
      "PTA_only"
    } else if (length(bulk_sources) == 1) {
      bulk_sources
    } else {
      "MULTIPLE_BULK"
    }
  })
  
  tibble(
    chr = sub_filt$chr,
    pos = sub_filt$pos,
    ref = sub_filt$ref,
    alt = sub_filt$alt,
    callset = cs,
    source_group = source_group_vec,
    PTA_only = source_group_vec == "PTA_only"
  )
})

# Apply the same exclusions used for ampli_alloc_tbl tier2
tier2_variant_tbl <- tier2_variant_tbl %>%
  dplyr::filter(callset != "call_HiDEFseq_single_stranded") %>%
  dplyr::filter(!(callset == "call_PTA_singleton" & source_group == "PTA_only"))

# Row-level count
tier2_variant_count_check <- tier2_variant_tbl %>%
  group_by(callset, PTA_only, source_group) %>%
  dplyr::summarise(n_rowlevel = n(), .groups = "drop") %>%
  arrange(callset, PTA_only, source_group)

# Allocation table count
tier2_alloc_count_check <- ampli_alloc_tbl %>%
  dplyr::filter(tier == "tier2") %>%
  dplyr::select(callset, PTA_only, source_group, n, step2_num_int) %>%
  rename(n_alloc_tbl = n) %>%
  arrange(callset, PTA_only, source_group)

# Compare
tier2_check_compare <- tier2_alloc_count_check %>%
  full_join(
    tier2_variant_count_check,
    by = c("callset", "PTA_only", "source_group")
  ) %>%
  mutate(
    n_alloc_tbl = dplyr::coalesce(n_alloc_tbl, 0L),
    n_rowlevel = dplyr::coalesce(n_rowlevel, 0L),
    count_match = (n_alloc_tbl == n_rowlevel),
    feasible = step2_num_int <= n_rowlevel
  ) %>%
  arrange(callset, PTA_only, source_group)

tier2_check_compare

tier2_check_compare %>% dplyr::select(count_match, feasible) %>% table(.)
```

### Tier2 - stratified selection
```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(readr)
})

set.seed(STRATIFICATION_SEED)

# 1. Allocation table in the exact stratum order you want
tier2_alloc_ordered <- ampli_alloc_tbl %>%
  dplyr::filter(tier == "tier2") %>%
  mutate(stratum_order = dplyr::row_number()) %>%
  dplyr::select(stratum_order, callset, PTA_only, source_group, n, step2_num_int)

# 2. Attach stratum order and target number to row-level tier2 variants
tier2_variant_tbl2 <- tier2_variant_tbl %>%
  inner_join(
    tier2_alloc_ordered %>%
      dplyr::select(stratum_order, callset, PTA_only, source_group, step2_num_int),
    by = c("callset", "PTA_only", "source_group")
  )

# 3. Stratified sampling in the same stratum order
tier2_sampled <- tier2_alloc_ordered %>%
  group_split(stratum_order, .keep = TRUE) %>%
  lapply(function(one_stratum) {
    this_order <- one_stratum$stratum_order[1]
    this_callset <- one_stratum$callset[1]
    this_pta <- one_stratum$PTA_only[1]
    this_source <- one_stratum$source_group[1]
    this_n <- one_stratum$step2_num_int[1]
    
    tier2_variant_tbl2 %>%
      dplyr::filter(
        stratum_order == this_order,
        callset == this_callset,
        PTA_only == this_pta,
        source_group == this_source
      ) %>%
      slice_sample(n = this_n) %>%
      mutate(sampled_n = this_n)
  }) %>%
  bind_rows() %>%
  arrange(stratum_order, chr, pos, ref, alt) %>%
  dplyr::select(
    stratum_order,
    chr, pos, ref, alt,
    callset, PTA_only, source_group,
    sampled_n
  )

tier2_sampled
```

### tier2 selection validation
```{r}
tier2_sampled_check <- tier2_sampled %>%
  group_by(stratum_order, callset, PTA_only, source_group) %>%
  dplyr::summarise(n_sampled = dplyr::n(), .groups = "drop") %>%
  left_join(
    tier2_alloc_ordered %>%
      dplyr::select(stratum_order, callset, PTA_only, source_group, step2_num_int),
    by = c("stratum_order", "callset", "PTA_only", "source_group")
  ) %>%
  mutate(match = (n_sampled == step2_num_int)) %>%
  arrange(stratum_order)

tier2_sampled_check





tier2_sampled_check$match %>% table(.)
```

### tier3 selection - check
```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(purrr)
  library(tidyr)
})

call_cols_full_depth_tier3 <- c(
  "call_Illumina",
  "call_ppmSeq_multiread",
  "call_ppmSeq_singleton_HC",
  "call_UDSeq_multiread",
  "call_UDSeq_singleton",
  "call_NanoSeq_multiread",
  "call_NanoSeq_singleton",
  "call_HiDEFseq_double_stranded",
  "call_PTA_singleton"
)

# Row-level tier3 table
tier3_variant_tbl <- purrr::map_dfr(call_cols_full_depth_tier3, function(cs) {
  
  other_call_cols <- setdiff(call_cols_full_depth_tier3, cs)
  
  self_exist_col <- get_self_exist_cols_full(cs)
  other_exist_cols <- setdiff(exist_cols_full, self_exist_col)
  
  self_called <- filt[[cs]] == 1
  other_called_n <- rowSums(as.data.frame(filt[, other_call_cols, drop = FALSE]), na.rm = TRUE)
  other_exist_n <- rowSums(as.data.frame(filt[, other_exist_cols, drop = FALSE]), na.rm = TRUE)
  
  idx_tier3 <- self_called &
    (other_called_n == 0) &
    (other_exist_n == 0) &
    (filt$chung_ampli_BLT_mosaic != 1)
  
  if (sum(idx_tier3, na.rm = TRUE) == 0) {
    return(tibble(
      chr = character(0),
      pos = numeric(0),
      ref = character(0),
      alt = character(0),
      callset = character(0),
      PTA_only = logical(0),
      source_group = character(0)
    ))
  }
  
  sub_filt <- filt[idx_tier3, , drop = FALSE]
  
  tibble(
    chr = sub_filt$chr,
    pos = sub_filt$pos,
    ref = sub_filt$ref,
    alt = sub_filt$alt,
    callset = cs,
    PTA_only = FALSE,
    source_group = "."
  )
})

# Row-level count
tier3_variant_count_check <- tier3_variant_tbl %>%
  group_by(callset, PTA_only, source_group) %>%
  dplyr::summarise(n_rowlevel = dplyr::n(), .groups = "drop") %>%
  arrange(callset)

# Allocation table count
tier3_alloc_count_check <- ampli_alloc_tbl %>%
  dplyr::filter(tier == "tier3") %>%
  dplyr::select(callset, PTA_only, source_group, n, step2_num_int) %>%
  rename(n_alloc_tbl = n) %>%
  arrange(callset)

# Compare
tier3_check_compare <- tier3_alloc_count_check %>%
  full_join(
    tier3_variant_count_check,
    by = c("callset", "PTA_only", "source_group")
  ) %>%
  mutate(
    n_alloc_tbl = dplyr::coalesce(n_alloc_tbl, 0L),
    n_rowlevel = dplyr::coalesce(n_rowlevel, 0L),
    count_match = (n_alloc_tbl == n_rowlevel),
    feasible = step2_num_int <= n_rowlevel
  ) %>%
  arrange(callset)

tier3_check_compare

tier3_check_compare %>% dplyr::select(count_match, feasible) %>% table(.)

```

### Tier3 selection
```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(readr)
})

set.seed(STRATIFICATION_SEED)

# 1. Allocation table in the exact stratum order
tier3_alloc_ordered <- ampli_alloc_tbl %>%
  dplyr::filter(tier == "tier3") %>%
  mutate(stratum_order = dplyr::row_number()) %>%
  dplyr::select(stratum_order, callset, PTA_only, source_group, n, step2_num_int)

# 2. Attach stratum info to row-level tier3 variants
tier3_variant_tbl2 <- tier3_variant_tbl %>%
  inner_join(
    tier3_alloc_ordered %>%
      dplyr::select(stratum_order, callset, PTA_only, source_group, step2_num_int),
    by = c("callset", "PTA_only", "source_group")
  )

# 3. Stratified sampling in the same order
tier3_sampled <- tier3_alloc_ordered %>%
  split(.$stratum_order) %>%
  lapply(function(one_stratum) {
    this_order <- one_stratum$stratum_order[1]
    this_callset <- one_stratum$callset[1]
    this_pta <- one_stratum$PTA_only[1]
    this_source <- one_stratum$source_group[1]
    this_n <- one_stratum$step2_num_int[1]
    
    tier3_variant_tbl2 %>%
      dplyr::filter(
        stratum_order == this_order,
        callset == this_callset,
        PTA_only == this_pta,
        source_group == this_source
      ) %>%
      slice_sample(n = this_n) %>%
      mutate(sampled_n = this_n)
  }) %>%
  bind_rows() %>%
  arrange(stratum_order, chr, pos, ref, alt) %>%
  dplyr::select(
    stratum_order,
    chr, pos, ref, alt,
    callset, PTA_only, source_group,
    sampled_n
  )

tier3_sampled

```

### Tier3 - selection check
```{r}
tier3_sampled_check <- tier3_sampled %>%
  group_by(stratum_order, callset, PTA_only, source_group) %>%
  dplyr::summarise(n_sampled = dplyr::n(), .groups = "drop") %>%
  left_join(
    tier3_alloc_ordered %>%
      dplyr::select(stratum_order, callset, PTA_only, source_group, step2_num_int),
    by = c("stratum_order", "callset", "PTA_only", "source_group")
  ) %>%
  mutate(match = (n_sampled == step2_num_int)) %>%
  arrange(stratum_order)

tier3_sampled_check


tier3_sampled_check$match %>% table(.)
```


### Tier1, 2, 3 BED format
### Germline het and refhom randomly selected to BED
```{r}
### Tier1 (n=50, non-overlap with chung's) to primer BED


library(dplyr)
library(stringr)
library(readr)
library(purrr)

set.seed(STRATIFICATION_SEED)


callset_order <- c(
  "vcf_Illumina",
  "vcf_ppmSeq_multiread",
  "vcf_ppmSeq_singleton_HC",
  "vcf_UDSeq_multiread",
  "vcf_UDSeq_singleton",
  "vcf_NanoSeq_multiread",
  "vcf_NanoSeq_singleton",
  "vcf_HiDEFseq_double_stranded",
  "vcf_HiDEFseq_single_stranded",
  "vcf_PTA_singleton"
)

tier1_bed <- tier1_selected_annot %>%
  mutate(
    start = pos - 20,
    end   = pos + 20,
    all_callsets = map2(callset, overlap_partner_callsets, \(x, y) {
      partners <- str_split(y, ",", simplify = FALSE)[[1]] %>%
        str_trim()
      c(x, partners)
    }),
    all_callsets = map(all_callsets, \(x) {
      x %>%
        str_replace("^call_", "vcf_") %>%
        unique() %>%
        .[order(match(., callset_order))]
    }),
    name = pmap_chr(
      list(chr, pos, ref, alt, all_callsets),
      \(chr, pos, ref, alt, all_callsets) {
        paste0(
          chr, "_", pos, "_", ref, "_", alt,
          "-tier1-",
          paste(all_callsets, collapse = "-")
        )
      }
    )
  ) %>%
  dplyr::select(chr, start, end, name)



tier2_bed <- tier2_sampled %>%
  mutate(
    start = pos - 20,
    end = pos + 20,
    callset_vcf = str_replace(callset, "^call_", "vcf_"),
    name = case_when(
      PTA_only ~ paste0(
        chr, "_", pos, "_", ref, "_", alt,
        "-tier2-",
        callset_vcf,
        "-bam_PTA"
      ),
      TRUE ~ paste0(
        chr, "_", pos, "_", ref, "_", alt,
        "-tier2-",
        callset_vcf,
        "-bam_",
        source_group
      )
    )
  ) %>%
  dplyr::select(chr, start, end, name)

tier3_bed <- tier3_sampled %>%
  mutate(
    start = pos - 20,
    end = pos + 20,
    callset_vcf = str_replace(callset, "^call_", "vcf_"),
    name = paste0(
      chr, "_", pos, "_", ref, "_", alt,
      "-tier3-",
      callset_vcf
    )
  ) %>%
  dplyr::select(chr, start, end, name)


prev_id01_ampli <- read_tsv("prev_id01_germline.with_vaf.tsv", show_col_types = FALSE)

het_prev_id01_ampli <- prev_id01_ampli %>%
    dplyr::filter(
        CATEGORY == "Het",
        dplyr::between(ppmseq_vaf, 0.46, 0.54),
        dplyr::between(illumina_vaf, 0.46, 0.54)
    )
hom_prev_id01_ampli <- prev_id01_ampli %>%
    dplyr::filter(
        CATEGORY == "Alt_Homo")



# expected to already exist:
# - het_prev_id01_ampli
# - hom_prev_id01_ampli
# - germ_het_num
# - germ_althom_num

germline_sampled <- bind_rows(
  het_prev_id01_ampli %>%
    slice_sample(n = germ_het_num),
  hom_prev_id01_ampli %>%
    slice_sample(n = germ_althom_num)
)

germline_bed <- germline_sampled %>%
  mutate(
    start = POS - 20,
    end = POS + 20,
    category_label = case_when(
      CATEGORY == "Het" ~ "GermlineHeterozygous",
      CATEGORY == "Alt_Homo" ~ "GermlineRefOrAltHomozygous",
      TRUE ~ CATEGORY
    ),
    name = paste0(
      CHROM, "_", POS, "_", REF, "_", ALT, "-",
      category_label
    )
  ) %>%
  dplyr::select(
    chr = CHROM,
    start,
    end,
    name
  )

germline_sampled
germline_bed



chrom_levels <- c(
  paste0("chr", 1:22),
  "chrX"
)

benchmark_stratified_7614_BLT_ampliseq_panel <- bind_rows(
  tier1_bed,
  tier2_bed,
  tier3_bed,
  germline_bed
) %>%
  mutate(
    chr = factor(chr, levels = chrom_levels, ordered = TRUE)
  ) %>%
  arrange(chr, start) %>%
  mutate(chr = as.character(chr))

readr::write_tsv(
  benchmark_stratified_7614_BLT_ampliseq_panel,
  "benchmark_stratified_7614_BLT_ampliseq_panel_fullcol4.bed",
  col_names = FALSE
)

benchmark_stratified_7614_BLT_ampliseq_panel_short <- benchmark_stratified_7614_BLT_ampliseq_panel %>%
  mutate(
    name = name %>%
      stringr::str_replace("-tier1-.*$", "-t1") %>%
      stringr::str_replace("-tier2-.*$", "-t2") %>%
      stringr::str_replace("-tier3-.*$", "-t3") %>%
      stringr::str_replace("-GermlineHeterozygous$", "-het") %>%
      stringr::str_replace("-GermlineRefOrAltHomozygous$", "-hom")
  )

readr::write_tsv(
  benchmark_stratified_7614_BLT_ampliseq_panel_short,
  "benchmark_stratified_7614_BLT_ampliseq_panel_short.bed",
  col_names = FALSE
)

```




### Comparison with Illumina DesignStudio panels
```{r}
res_dir <- "design_n780_seed12345/"

long_panel <- read_tsv(paste0(res_dir, "benchmark_stratified_7614_BLT_ampliseq_panel_fullcol4.bed"), col_names = F) %>% setNames(c("chr", "start", "end", "name"))
short_panel <- read_tsv(paste0(res_dir, "benchmark_stratified_7614_BLT_ampliseq_panel_short.bed"), col_names = F) %>% setNames(c("chr", "start", "end", "short_name"))

full_ampli_design <- long_panel
full_ampli_design$short_name <- short_panel$short_name


designed <- read_tsv(paste0(res_dir,"illumina_result.tsv"))
designed_joined <- designed %>%
    left_join(
        full_ampli_design %>%
            dplyr::select(short_name, name),
        by = c("Name" = "short_name")
    )

ampli_numbers <- read_tsv(paste0(res_dir, "ampliseq_allocation_numbers_by_tier_and_strata.tsv"), show_col_types = FALSE)

submitted_tbl <- designed_joined %>%
  dplyr::filter(Num_Amplicons == 1) %>%
  mutate(
    tier = case_when(
      str_detect(name, "-tier1-") ~ "tier1",
      str_detect(name, "-tier2-") ~ "tier2",
      str_detect(name, "-tier3-") ~ "tier3",
      str_detect(name, "GermlineRefOrAltHomozygous$") ~ "germline_althom",
      str_detect(name, "GermlineHeterozygous$") ~ "germline_het",
      TRUE ~ NA_character_
    ),
    callset = case_when(
      tier == "tier1" ~ "call_shared",
      tier == "tier2" ~ str_match(name, "-tier2-vcf_(.+)-bam_([^_]+(?:_[^_]+)*)$")[,2],
      tier == "tier3" ~ str_match(name, "-tier3-vcf_(.+)$")[,2],
      tier == "germline_althom" ~ "germline_althom",
      tier == "germline_het" ~ "germline_het",
      TRUE ~ NA_character_
    ),
    source_raw = case_when(
      tier == "tier2" ~ str_match(name, "-tier2-vcf_(.+)-bam_([^_]+(?:_[^_]+)*)$")[,3],
      TRUE ~ "."
    ),
    PTA_only = case_when(
      tier == "tier2" & source_raw == "PTA" ~ TRUE,
      TRUE ~ FALSE
    ),
    source_group = case_when(
      tier == "tier2" & source_raw == "PTA" ~ "PTA_only",
      tier == "tier2" ~ source_raw,
      TRUE ~ "."
    ),
    callset = if_else(str_starts(callset, "call_") | str_starts(callset, "germline_"),
                      callset,
                      paste0("call_", callset))
  ) %>%
  dplyr::count(tier, callset, PTA_only, source_group, name = "submitted_num")

ampli_numbers2 <- ampli_numbers %>%
  left_join(
    submitted_tbl,
    by = c("tier", "callset", "PTA_only", "source_group")
  ) %>%
  mutate(
    submitted_num = coalesce(submitted_num, 0L)
  ) %>%
  relocate(submitted_num, .after = step2_num_int)

ampli_numbers2


ampli_numbers2 %>% write_tsv(., paste0(res_dir, "before_after_ampli_num.tsv"))

```


### Comparison with Illumina DesignStudio panels - merged
```{r}
res_dir <- "merged_design_n784/"



long_panel <- read_tsv(paste0(res_dir, "best_ampli_design_n784_stratified_manual.bed"), col_names = F) %>% setNames(c("chr", "start", "end", "name", "seed"))
short_panel <- read_tsv(paste0(res_dir, "best_ampli_design_n784_stratified_manual.shortname.bed"), col_names = F) %>% setNames(c("chr", "start", "end", "short_name"))

full_ampli_design <- long_panel
full_ampli_design$short_name <- short_panel$short_name


designed <- read_tsv(paste0(res_dir,"illumina_result.tsv"))
designed_joined <- designed %>%
    left_join(
        full_ampli_design %>%
            dplyr::select(short_name, name),
        by = c("Name" = "short_name")
    )

ampli_numbers <- read_tsv(paste0(res_dir, "ampliseq_allocation_numbers_by_tier_and_strata.tsv"), show_col_types = FALSE)

submitted_tbl <- designed_joined %>%
  dplyr::filter(Num_Amplicons == 1) %>%
  mutate(
    tier = case_when(
      str_detect(name, "-tier1-") ~ "tier1",
      str_detect(name, "-tier2-") ~ "tier2",
      str_detect(name, "-tier3-") ~ "tier3",
      str_detect(name, "GermlineRefOrAltHomozygous$") ~ "germline_althom",
      str_detect(name, "GermlineHeterozygous$") ~ "germline_het",
      TRUE ~ NA_character_
    ),
    callset = case_when(
      tier == "tier1" ~ "call_shared",
      tier == "tier2" ~ str_match(name, "-tier2-vcf_(.+)-bam_([^_]+(?:_[^_]+)*)$")[,2],
      tier == "tier3" ~ str_match(name, "-tier3-vcf_(.+)$")[,2],
      tier == "germline_althom" ~ "germline_althom",
      tier == "germline_het" ~ "germline_het",
      TRUE ~ NA_character_
    ),
    source_raw = case_when(
      tier == "tier2" ~ str_match(name, "-tier2-vcf_(.+)-bam_([^_]+(?:_[^_]+)*)$")[,3],
      TRUE ~ "."
    ),
    PTA_only = case_when(
      tier == "tier2" & source_raw == "PTA" ~ TRUE,
      TRUE ~ FALSE
    ),
    source_group = case_when(
      tier == "tier2" & source_raw == "PTA" ~ "PTA_only",
      tier == "tier2" ~ source_raw,
      TRUE ~ "."
    ),
    callset = if_else(str_starts(callset, "call_") | str_starts(callset, "germline_"),
                      callset,
                      paste0("call_", callset))
  ) %>%
  dplyr::count(tier, callset, PTA_only, source_group, name = "submitted_num")

ampli_numbers2 <- ampli_numbers %>%
  left_join(
    submitted_tbl,
    by = c("tier", "callset", "PTA_only", "source_group")
  ) %>%
  mutate(
    submitted_num = coalesce(submitted_num, 0L)
  ) %>%
  relocate(submitted_num, .after = step2_num_int)

ampli_numbers2


ampli_numbers2 %>% write_tsv(., paste0(res_dir, "before_after_ampli_num.tsv"))

```


```{r}
library(dplyr)
library(readr)
library(stringr)
library(purrr)

files <- c(
  "design_n780_seed42/before_after_ampli_num.tsv",
  "design_n780_seed123/before_after_ampli_num.tsv",
  "design_n780_seed2026/before_after_ampli_num.tsv",
  "design_n780_seed12345/before_after_ampli_num.tsv"
)

seed_order <- c("42", "123", "2026", "12345")

# Canonical stratum order from the first TSV
stratum_order_tbl <- readr::read_tsv(files[1], show_col_types = FALSE) %>%
  mutate(stratum_order = dplyr::row_number()) %>%
  dplyr::select(stratum_order, tier, callset, PTA_only, source_group)

# Read all files
dat_all <- purrr::map_dfr(files, function(f) {
  seed_now <- stringr::str_extract(f, "seed[0-9]+") %>%
    stringr::str_remove("seed")

  readr::read_tsv(f, show_col_types = FALSE) %>%
    mutate(
      seed = seed_now,
      submitted_num = as.integer(submitted_num),
      step2_num_int = as.integer(step2_num_int),
      design_rate_num = dplyr::if_else(step2_num_int == 0, NA_real_, submitted_num / step2_num_int)
    ) %>%
    left_join(stratum_order_tbl, by = c("tier", "callset", "PTA_only", "source_group"))
})

# Summarise by stratum
best_by_stratum <- dat_all %>%
  mutate(seed = factor(seed, levels = seed_order)) %>%
  group_by(stratum_order, tier, callset, PTA_only, source_group) %>%
  summarise(
    step2_num_int = dplyr::first(step2_num_int),
    submitted_num = paste(submitted_num[order(seed)], collapse = "_"),
    max_rate = suppressWarnings(max(design_rate_num, na.rm = TRUE)),
    best_seed = {
      mr <- suppressWarnings(max(design_rate_num, na.rm = TRUE))
      if (is.infinite(mr)) {
        "."
      } else {
        paste(as.character(seed[design_rate_num == mr]), collapse = "_")
      }
    },
    best_design_rate = {
      mr <- suppressWarnings(max(design_rate_num, na.rm = TRUE))
      if (is.infinite(mr)) "." else sprintf("%.6f", mr)
    },
    .groups = "drop"
  ) %>%
  arrange(stratum_order) %>%
  mutate(
    PTA_only = as.character(PTA_only),
    step2_num_int = as.character(step2_num_int)
  ) %>%
  dplyr::select(
    tier,
    callset,
    PTA_only,
    source_group,
    step2_num_int,
    submitted_num,
    best_seed,
    best_design_rate
  )

print(best_by_stratum, n = Inf)

readr::write_tsv(best_by_stratum, "best_seed_by_stratum.tsv", na = "")
```


```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(purrr)
})

set.seed(1)

callsets_use <- c(
  "call_Illumina",
  "call_ppmSeq_multiread",
  "call_ppmSeq_singleton_HC",
  "call_UDSeq_multiread",
  "call_UDSeq_singleton",
  "call_NanoSeq_multiread",
  "call_NanoSeq_singleton",
  "call_HiDEFseq_double_stranded",
  "call_PTA_singleton"
)

filt2 <- filt %>%
  mutate(
    variant_key = paste(chr, pos, ref, alt, sep = ":")
  )

tier3_candidates <- purrr::map_dfr(callsets_use, function(cs) {
  
  other_call_cols <- setdiff(call_cols_full_depth, cs)
  self_exist_col_full <- get_self_exist_cols_full_altge1(cs)
  other_exist_cols_full <- setdiff(exist_cols_full_altge1, self_exist_col_full)
  
  self_called <- filt2[[cs]] == 1
  other_called_n <- rowSums(as.data.frame(filt2[, other_call_cols, drop = FALSE]), na.rm = TRUE)
  other_exist_n_full <- rowSums(as.data.frame(filt2[, other_exist_cols_full, drop = FALSE]), na.rm = TRUE)
  
  idx <- self_called &
    (other_called_n == 0) &
    (other_exist_n_full == 0) &
    (filt2$chung_ampli_BLT_mosaic == 0)
  
  filt2[idx, , drop = FALSE] %>%
    transmute(
      chr,
      pos,
      ref,
      alt,
      variant_key,
      callset = cs
    ) %>%
    distinct()
})

tier3_selected <- tier3_candidates %>%
  group_by(callset) %>%
  slice_sample(prop = 1) %>%
  slice_head(n = 20) %>%
  ungroup() %>%
  arrange(callset, chr, pos, ref, alt)

tier3_selected_summary <- tier3_candidates %>%
  count(callset, name = "avail_n") %>%
  left_join(
    tier3_selected %>% count(callset, name = "selected_n"),
    by = "callset"
  ) %>%
  mutate(
    selected_n = dplyr::coalesce(selected_n, 0L)
  ) %>%
  arrange(avail_n)

tier3_selected
tier3_selected_summary
```

```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(stringr)
  library(UpSetR)
})

# ---- Build matrix for UpSetR (missing cols -> 0) ----
make_mat_strata <- function(df, cols) {
  out <- lapply(cols, function(x) {
    if (x %in% colnames(df)) as.integer(df[[x]] %in% 1) else rep(0L, nrow(df))
  })
  mat <- as.data.frame(out, check.names = FALSE)
  colnames(mat) <- cols
  rownames(mat) <- NULL
  mat
}

# ---- Tier1 rows: called by >= 2 VCF callsets ----
tier1_idx <- rowSums(as.data.frame(tiering_df[, call_cols_full_depth, drop = FALSE]), na.rm = TRUE) >= 2

filt_tier1 <- tiering_df[tier1_idx, , drop = FALSE]

cat("Tier1 rows:", nrow(filt_tier1), "\n")

# ---- UpSet input ----
mat_calls_tier1 <- make_mat_strata(filt_tier1, call_cols_full_depth)

# Optional prettier labels
colnames(mat_calls_tier1) <- str_replace(colnames(mat_calls_tier1), "^call_", "")

# ---- Plot ----
upset(
  mat_calls_tier1,
  nsets = ncol(mat_calls_tier1),
  nintersects = 200,
  keep.order = TRUE,
  order.by = "freq",
  mainbar.y.label = "Tier1 intersection size",
  sets.x.label = "Tier1 set size"
)

```

```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(stringr)
  library(ggplot2)
  library(ComplexUpset)
})

call_cols_full_depth <- c(
  "call_Illumina",
  "call_ppmSeq_multiread",
  "call_ppmSeq_singleton_HC",
  "call_UDSeq_multiread",
  "call_UDSeq_singleton",
  "call_NanoSeq_multiread",
  "call_NanoSeq_singleton",
  "call_HiDEFseq_double_stranded",
  "call_HiDEFseq_single_stranded",
  "call_PTA_singleton"
)

# ---- Tier1 rows: called by >= 2 VCF callsets ----
tier1_idx <- rowSums(as.data.frame(tiering_df[, call_cols_full_depth, drop = FALSE]), na.rm = TRUE) >= 2

filt_tier1 <- tiering_df[tier1_idx, , drop = FALSE] %>%
  mutate(
    chung_group = ifelse(chung_ampli_BLT_mosaic == 1, "B-L-T_chung_et_al_2024", "new")
  )

cat("Tier1 rows:", nrow(filt_tier1), "\n")

# ---- make 0/1 logical columns for ComplexUpset ----
plot_df <- filt_tier1 %>%
  mutate(across(all_of(call_cols_full_depth), ~ .x %in% 1)) %>%
  rename_with(~ str_replace(.x, "^call_", ""), all_of(call_cols_full_depth))

plot_sets <- str_replace(call_cols_full_depth, "^call_", "")

# ---- stacked upset ----
upset(
  plot_df,
  intersect = plot_sets,
  sort_intersections_by = "cardinality",
  sort_sets = FALSE,
  base_annotations = list(
    "Tier1 intersection size" = (
      ggplot(mapping = aes(fill = chung_group)) +
        geom_bar(stat = "count", position = "stack") +
        geom_text(
          stat = "count",
          aes(label = after_stat(count)),
          position = position_stack(vjust = 0.5),
          size = 3
        ) +
        labs(y = "Tier1 intersection size", fill = NULL)
    )
  ),
  width_ratio = 0.2
)
```

### PTA-including tier2
```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(purrr)
  library(stringr)
  library(ggplot2)
})

call_cols_full_depth <- c(
  "call_Illumina",
  "call_ppmSeq_multiread",
  "call_ppmSeq_singleton_HC",
  "call_UDSeq_multiread",
  "call_UDSeq_singleton",
  "call_NanoSeq_multiread",
  "call_NanoSeq_singleton",
  "call_HiDEFseq_double_stranded",
  "call_HiDEFseq_single_stranded",
  "call_PTA_singleton"
)

exist_pretty_map <- c(
  "is_altge1_ILLUMINA" = "ILLUMINA",
  "is_altge1_DUPLEX_PPMSEQ" = "PPMSEQ",
  "is_altge1_DUPLEX_UDSEQ" = "UDSEQ",
  "is_altge1_DUPLEX_NANOSEQ" = "NANOSEQ",
  "is_altge1_DUPLEX_HIDEFSEQ" = "HIDEFSEQ",
  "existing_any_PTA1_16" = "PTA"
)

tier2_new_source_df <- purrr::map_dfr(call_cols_full_depth, function(cs) {
  
  other_call_cols <- setdiff(call_cols_full_depth, cs)
  self_exist_col <- get_self_exist_cols_full_altge1(cs)
  other_exist_cols <- setdiff(exist_cols_full_altge1, self_exist_col)
  
  self_called <- filt[[cs]] == 1
  other_called_n <- rowSums(as.data.frame(filt[, other_call_cols, drop = FALSE]), na.rm = TRUE)
  other_exist_mat <- as.data.frame(filt[, other_exist_cols, drop = FALSE])
  other_exist_n <- rowSums(other_exist_mat, na.rm = TRUE)
  
  idx_tier2_new <- self_called &
    (other_called_n == 0) &
    (other_exist_n >= 1) &
    (filt$chung_ampli_BLT_mosaic  != 1)
  
  if (sum(idx_tier2_new, na.rm = TRUE) == 0) {
    return(tibble(callset = cs, source = character(0), n = integer(0)))
  }
  
  src_class <- apply(other_exist_mat[idx_tier2_new, , drop = FALSE], 1, function(x) {
    hit_cols <- names(x)[which(x >= 1)]
    if (length(hit_cols) >= 2) {
      "MULTIPLE"
    } else {
      unname(exist_pretty_map[hit_cols])
    }
  })
  
  tibble(callset = cs, source = src_class) %>%
    dplyr::count(callset, source, name = "n")
})


tier2_new_source_df <- tier2_new_source_df %>%
  mutate(
    source = factor(
      source,
      levels = c("MULTIPLE", "ILLUMINA", "PPMSEQ", "UDSEQ", "NANOSEQ", "HIDEFSEQ", "PTA")
    )
  )


source_levels <- c("MULTIPLE", "ILLUMINA", "PPMSEQ", "UDSEQ", "NANOSEQ", "HIDEFSEQ", "PTA")

# self source mapping
callset_to_source <- c(
  "call_Illumina" = "ILLUMINA",
  "call_ppmSeq_multiread" = "PPMSEQ",
  "call_ppmSeq_singleton_HC" = "PPMSEQ",
  "call_UDSeq_multiread" = "UDSEQ",
  "call_UDSeq_singleton" = "UDSEQ",
  "call_NanoSeq_multiread" = "NANOSEQ",
  "call_NanoSeq_singleton" = "NANOSEQ",
  "call_HiDEFseq_double_stranded" = "HIDEFSEQ",
  "call_HiDEFseq_single_stranded" = "HIDEFSEQ",
  "call_PTA_singleton" = "PTA"
)

allowed_grid <- tidyr::expand_grid(
  callset = call_cols_full_depth,
  source = source_levels
) %>%
  mutate(
    self_source = unname(callset_to_source[as.character(callset)])
  ) %>%
  dplyr::filter(source != self_source) %>%
  dplyr::select(-self_source)

tier2_new_source_df2 <- allowed_grid %>%
  left_join(
    tier2_new_source_df %>%
      mutate(
        callset = as.character(callset),
        source = as.character(source)
      ),
    by = c("callset", "source")
  ) %>%
  mutate(
    n = dplyr::coalesce(n, 0L),
    callset = factor(callset, levels = call_cols_full_depth),
    source = factor(source, levels = source_levels)
  ) %>%
  arrange(callset, source)

tier2_new_source_df2
```

```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(purrr)
  library(stringr)
  library(tidyr)
})

call_cols_full_depth <- c(
  "call_Illumina",
  "call_ppmSeq_multiread",
  "call_ppmSeq_singleton_HC",
  "call_UDSeq_multiread",
  "call_UDSeq_singleton",
  "call_NanoSeq_multiread",
  "call_NanoSeq_singleton",
  "call_HiDEFseq_double_stranded",
  "call_HiDEFseq_single_stranded",
  "call_PTA_singleton"
)

exist_pretty_map <- c(
  "is_altge1_ILLUMINA" = "ILLUMINA",
  "is_altge1_DUPLEX_PPMSEQ" = "PPMSEQ",
  "is_altge1_DUPLEX_UDSEQ" = "UDSEQ",
  "is_altge1_DUPLEX_NANOSEQ" = "NANOSEQ",
  "is_altge1_DUPLEX_HIDEFSEQ" = "HIDEFSEQ",
  "existing_any_PTA1_16" = "PTA"
)

source_levels <- c("ILLUMINA", "PPMSEQ", "UDSEQ", "NANOSEQ", "HIDEFSEQ", "PTA")

callset_to_source <- c(
  "call_Illumina" = "ILLUMINA",
  "call_ppmSeq_multiread" = "PPMSEQ",
  "call_ppmSeq_singleton_HC" = "PPMSEQ",
  "call_UDSeq_multiread" = "UDSEQ",
  "call_UDSeq_singleton" = "UDSEQ",
  "call_NanoSeq_multiread" = "NANOSEQ",
  "call_NanoSeq_singleton" = "NANOSEQ",
  "call_HiDEFseq_double_stranded" = "HIDEFSEQ",
  "call_HiDEFseq_single_stranded" = "HIDEFSEQ",
  "call_PTA_singleton" = "PTA"
)

tier2_new_source_combo_df <- purrr::map_dfr(call_cols_full_depth, function(cs) {
  
  other_call_cols <- setdiff(call_cols_full_depth, cs)
  self_exist_col <- get_self_exist_cols_full_altge1(cs)
  other_exist_cols <- setdiff(exist_cols_full_altge1, self_exist_col)
  
  self_called <- filt[[cs]] == 1
  other_called_n <- rowSums(as.data.frame(filt[, other_call_cols, drop = FALSE]), na.rm = TRUE)
  other_exist_mat <- as.data.frame(filt[, other_exist_cols, drop = FALSE])
  other_exist_n <- rowSums(other_exist_mat, na.rm = TRUE)
  
  idx_tier2_new <- self_called &
    (other_called_n == 0) &
    (other_exist_n >= 1) &
    (filt$chung_ampli_BLT_mosaic != 1)
  
  if (sum(idx_tier2_new, na.rm = TRUE) == 0) {
    return(tibble(callset = character(0), source_combo = character(0), n = integer(0)))
  }
  
  combo_vec <- apply(other_exist_mat[idx_tier2_new, , drop = FALSE], 1, function(x) {
    hit_cols <- names(x)[which(x >= 1)]
    hit_sources <- unname(exist_pretty_map[hit_cols]) %>%
      unique() %>%
      .[order(match(., source_levels))]
    
    paste(hit_sources, collapse = "+")
  })
  
  tibble(
    callset = cs,
    source_combo = combo_vec
  ) %>%
    dplyr::count(callset, source_combo, name = "n") %>%
    arrange(callset, desc(n), source_combo)
})

tier2_new_source_combo_df <- tier2_new_source_combo_df %>%
  mutate(
    has_PTA = str_detect(source_combo, "(^|\\+)PTA($|\\+)"),
    n_sources = str_count(source_combo, "\\+") + 1
  )

tier2_new_source_combo_df
```

```{r}
tier2_new_source_considerPTA_df %>% write_tsv("case1_lowci_stratum.tsv")
altge1_tier2_new_source_considerPTA_df %>% write_tsv("case2_altge1_stratum.tsv")


```

```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(purrr)
  library(stringr)
  library(ggplot2)
  library(tidyr)
})

call_cols_full_depth <- c(
  "call_Illumina",
  "call_ppmSeq_multiread",
  "call_ppmSeq_singleton_HC",
  "call_UDSeq_multiread",
  "call_UDSeq_singleton",
  "call_NanoSeq_multiread",
  "call_NanoSeq_singleton",
  "call_HiDEFseq_double_stranded",
  "call_HiDEFseq_single_stranded",
  "call_PTA_singleton"
)

# Use ONLY non-PTA exist columns for source calculation
exist_cols_source_noPTA <- c(
  "is_altge1_ILLUMINA",
  "is_altge1_DUPLEX_PPMSEQ",
  "is_altge1_DUPLEX_UDSEQ",
  "is_altge1_DUPLEX_NANOSEQ",
  "is_altge1_DUPLEX_HIDEFSEQ"
)


exist_pretty_map_noPTA <- c(
  "is_altge1_ILLUMINA" = "ILLUMINA",
  "is_altge1_DUPLEX_PPMSEQ" = "PPMSEQ",
  "is_altge1_DUPLEX_UDSEQ" = "UDSEQ",
  "is_altge1_DUPLEX_NANOSEQ" = "NANOSEQ",
  "is_altge1_DUPLEX_HIDEFSEQ" = "HIDEFSEQ"
)

tier2_new_source_df_noPTA <- purrr::map_dfr(call_cols_full_depth, function(cs) {
  
  other_call_cols <- setdiff(call_cols_full_depth, cs)
  self_exist_col <- get_self_exist_cols_noPTA(cs)
  other_exist_cols <- setdiff(exist_cols_source_noPTA, self_exist_col)
  
  self_called <- filt[[cs]] == 1
  other_called_n <- rowSums(as.data.frame(filt[, other_call_cols, drop = FALSE]), na.rm = TRUE)
  other_exist_mat <- as.data.frame(filt[, other_exist_cols, drop = FALSE])
  other_exist_n <- rowSums(other_exist_mat, na.rm = TRUE)
  
  idx_tier2_new <- self_called &
    (other_called_n == 0) &
    (other_exist_n >= 1) &
    (filt$chung_ampli_BLT_mosaic  != 1)
  
  if (sum(idx_tier2_new, na.rm = TRUE) == 0) {
    return(tibble(callset = cs, source = character(0), n = integer(0)))
  }
  
  src_class <- apply(other_exist_mat[idx_tier2_new, , drop = FALSE], 1, function(x) {
    hit_cols <- names(x)[which(x >= 1)]
    if (length(hit_cols) >= 2) {
      "MULTIPLE"
    } else {
      unname(exist_pretty_map_noPTA[hit_cols])
    }
  })
  
  tibble(callset = cs, source = src_class) %>%
    dplyr::count(callset, source, name = "n")
})

tier2_new_source_df_noPTA <- tier2_new_source_df_noPTA %>%
  mutate(
    source = factor(
      source,
      levels = c("MULTIPLE", "ILLUMINA", "PPMSEQ", "UDSEQ", "NANOSEQ", "HIDEFSEQ")
    )
  )

source_levels_noPTA <- c("MULTIPLE", "ILLUMINA", "PPMSEQ", "UDSEQ", "NANOSEQ", "HIDEFSEQ")

# PTA intentionally omitted here because it is not a source category in noPTA mode
callset_to_source_noPTA <- c(
  "call_Illumina" = "ILLUMINA",
  "call_ppmSeq_multiread" = "PPMSEQ",
  "call_ppmSeq_singleton_HC" = "PPMSEQ",
  "call_UDSeq_multiread" = "UDSEQ",
  "call_UDSeq_singleton" = "UDSEQ",
  "call_NanoSeq_multiread" = "NANOSEQ",
  "call_NanoSeq_singleton" = "NANOSEQ",
  "call_HiDEFseq_double_stranded" = "HIDEFSEQ",
  "call_HiDEFseq_single_stranded" = "HIDEFSEQ"
)

allowed_grid_noPTA <- tidyr::expand_grid(
  callset = call_cols_full_depth,
  source = source_levels_noPTA
) %>%
  mutate(
    self_source = unname(callset_to_source_noPTA[as.character(callset)])
  ) %>%
  dplyr::filter(is.na(self_source) | source != self_source) %>%
  dplyr::select(-self_source)

tier2_new_source_df2_noPTA <- allowed_grid_noPTA %>%
  left_join(
    tier2_new_source_df_noPTA %>%
      mutate(
        callset = as.character(callset),
        source = as.character(source)
      ),
    by = c("callset", "source")
  ) %>%
  mutate(
    n = dplyr::coalesce(n, 0L),
    callset = factor(callset, levels = call_cols_full_depth),
    source = factor(source, levels = source_levels_noPTA)
  ) %>%
  arrange(callset, source)

tier2_new_source_df2_noPTA
```

```{r}
ggplot(tier2_new_source_df2, aes(x = callset, y = n, fill = source)) +
  geom_col(position = "fill", width = 0.8) +
  coord_flip() +
  labs(
    x = NULL,
    y = "Fraction",
    fill = "Other BAM evidence source"
  ) +
  theme_bw(base_size = 12)


ggplot(tier2_new_source_df2_noPTA, aes(x = callset, y = n, fill = source)) +
  geom_col(position = "fill", width = 0.8) +
  coord_flip() +
  labs(
    x = NULL,
    y = "Fraction",
    fill = "Other BAM evidence source"
  ) +
  theme_bw(base_size = 12)

```


### Filtering profile
```{r}
library(grid)


call_cols_full_depth <- c(
  "call_Illumina",
  "call_ppmSeq_multiread",
  "call_ppmSeq_singleton_HC",
  "call_UDSeq_multiread",
  "call_UDSeq_singleton",
  "call_NanoSeq_multiread",
  "call_NanoSeq_singleton",
  "call_HiDEFseq_double_stranded",
  "call_HiDEFseq_single_stranded",
  "call_PTA_singleton"
)

no_black <- raw %>%
  dplyr::filter(
    blacklist_any==0,
    gnomad_common==0,
    in_hmer7==0,
    in_simple_repeats_ucsc==0
  )

count_calls <- function(df, call_cols){
  df %>%
    summarise(
      dplyr::across(dplyr::all_of(call_cols), ~ sum(.x==1, na.rm=TRUE))
    )
}

# ---- Step0: before filtering ----
cnt0 <- count_calls(no_black, call_cols_full_depth)

# ---- Germline filter (is_germline + any germline_like) ----
gll_cols <- colnames(no_black)[stringr::str_detect(
  colnames(no_black),
  "^is_germline_like_upper_ci_052_"
)]

after_germline <- no_black %>%
  dplyr::filter(
    !(is_germline==1 |
      rowSums(dplyr::across(dplyr::all_of(gll_cols), ~ as.integer(.x==1)), na.rm=TRUE) > 0)
  )

cnt1 <- count_calls(after_germline, call_cols_full_depth)

# ---- Unrelated filter ----
after_unrelated_pon <- after_germline %>%
  dplyr::filter(existing_in_unrelated==0)
  #dplyr::filter(altge1_in_unrelated==0)

cnt2 <- count_calls(after_unrelated_pon, call_cols_full_depth)

# ---- Final table (rows = callset) ----
out2 <- dplyr::bind_rows(
  cnt0 %>% mutate(stage="before_filter"),
  cnt1 %>% mutate(stage="after_germline"),
  cnt2 %>% mutate(stage="after_unrelated_pon")
) %>%
  tidyr::pivot_longer(-stage, names_to="callset", values_to="n") %>%
  tidyr::pivot_wider(names_from=stage, values_from=n) %>%
  dplyr::select(
    callset,
    before_filter,
    after_germline,
    after_unrelated_pon
  )

out2


viz_rel_stacked <- out2 %>%
  dplyr::mutate(callset = factor(callset, levels = rev(out2$callset))) %>%
  dplyr::rename(
    before = before_filter,
    after_germline = after_germline,
    after_pon = after_unrelated_pon
  ) %>%
  dplyr::mutate(
    removed_germline = (before - after_germline) / before,
    removed_unrelated_pon      = (after_germline - after_pon) / before,
    kept             = after_pon / before
  ) %>%
  dplyr::select(callset, removed_germline, removed_unrelated_pon, kept) %>%
  tidyr::pivot_longer(
    -callset,
    names_to = "component",
    values_to = "prop"
  ) %>%
  dplyr::mutate(
    component = factor(
      component,
      levels = c("removed_germline", "removed_unrelated_pon", "kept")
    )
  )

new_lev <- viz_rel_stacked$callset %>% levels(.) %>% str_replace_all(., "^call_", "")
viz_rel_stacked$callset <- viz_rel_stacked$callset %>% str_replace_all(., "^call_", "")
viz_rel_stacked$callset <- factor(viz_rel_stacked$callset, levels = new_lev)


lab_df <- viz_rel_stacked %>%
  dplyr::mutate(component = as.character(component)) %>%
  dplyr::group_by(callset) %>%
  dplyr::arrange(match(component, c("kept", "removed_unrelated_pon", "removed_germline")), .by_group = TRUE) %>%
  dplyr::mutate(
    y_center = cumsum(prop) - prop / 2,
    y_edge   = cumsum(prop),
    y_line   = y_edge + 0.010,
    y_label  = y_edge + 0.028,
    label = ifelse(
      prop >= 0.03,
      scales::percent(prop, accuracy = 0.1),
      scales::percent(prop, accuracy = 0.01)
    )
  ) %>%
  dplyr::ungroup()

ggplot(viz_rel_stacked, aes(x = callset, y = prop, fill = component)) +
  geom_col(width = 0.75) +

  # inside
  geom_text(
    data = lab_df %>% dplyr::filter(prop >= 0.03),
    aes(x = callset, y = y_center, label = label),
    inherit.aes = FALSE,
    size = 2.5
  ) +

  # thin connector line
  geom_segment(
    data = lab_df %>% dplyr::filter(prop < 0.03, prop > 0),
    aes(x = callset, xend = callset, y = y_edge, yend = y_line),
    inherit.aes = FALSE,
    linewidth = 0.3
  ) +

  # label
  geom_text(
    data = lab_df %>% dplyr::filter(prop < 0.03, prop > 0),
    aes(x = callset, y = y_label, label = label),
    inherit.aes = FALSE,
    hjust = 0,
    size = 2.4
  ) +

  coord_flip(clip = "off") +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.10))
  ) +
  theme_classic() +
  labs(x = NULL, y = "Proportion (%)")
```

```{r}
# nogerm -> filt : per-call counts (order = call_cols_full_depth 유지)

count_calls <- function(df, call_cols){
  df %>%
    summarise(dplyr::across(dplyr::all_of(call_cols), ~ sum(.x==1, na.rm=TRUE)))
}

cnt_before <- count_calls(nogerm, call_cols_full_depth) %>%
  mutate(stage = "before_unrelated")

cnt_after <- count_calls(filt, call_cols_full_depth) %>%
  mutate(stage = "after_unrelated")

unrelated_call_profile <- dplyr::bind_rows(cnt_before, cnt_after) %>%
  tidyr::pivot_longer(-stage, names_to = "callset", values_to = "n") %>%
  tidyr::pivot_wider(names_from = stage, values_from = n) %>%
  dplyr::mutate(
    removed = before_unrelated - after_unrelated,
    removed_frac = dplyr::if_else(
      before_unrelated > 0,
      removed / before_unrelated,
      NA_real_
    ),
    callset = factor(callset, levels = call_cols_full_depth)
  ) %>%
  dplyr::arrange(callset)

unrelated_call_profile



library(dplyr)
library(tidyr)
library(stringr)

platform_tokens <- list(
  call_Illumina = c(
    "ILLUMINA_6566_BLOOD",
    "ILLUMINA_6566_SPERM",
    "HG19_ILLUMINA_7669"
  ),
  call_ppmSeq_multiread = "DUPLEX_PPMSEQ_6566",
  call_ppmSeq_putative_multiread = "DUPLEX_PPMSEQ_6566",
  call_ppmSeq_singleton_HC = "DUPLEX_PPMSEQ_6566",
  call_UDSeq_multiread = "DUPLEX_UDSEQ_6566",
  call_UDSeq_singleton = "DUPLEX_UDSEQ_6566",
  call_NanoSeq_multiread = "DUPLEX_NANOSEQ_6566",
  call_NanoSeq_singleton = "DUPLEX_NANOSEQ_6566",
  call_PTA_singleton = c(
    "HG19_PTA_6566_1",
    "HG19_PTA_6566_2",
    "HG19_PTA_8420_1",
    "HG19_PTA_8420_2"
  ),
  call_PTA_multiple = c(
    "HG19_PTA_6566_1",
    "HG19_PTA_6566_2",
    "HG19_PTA_8420_1",
    "HG19_PTA_8420_2"
  )
)

platform_fdr <- lapply(call_cols_full_depth, function(cc){

  df_call <- nogerm %>%
    dplyr::filter(.data[[cc]] == 1)

  total_calls <- nrow(df_call)

  tokens_use <- platform_tokens[[cc]]

  df_token_hits <- df_call %>%
    tidyr::separate_rows(existing_in_unrelated_which, sep = ",") %>%
    dplyr::mutate(token = str_trim(existing_in_unrelated_which)) %>%
    #tidyr::separate_rows(altge1_in_unrelated_which, sep = ",") %>%
    #dplyr::mutate(token = str_trim(altge1_in_unrelated_which)) %>%
    dplyr::filter(token %in% tokens_use) %>%
    dplyr::distinct(chr, pos, ref, alt, token)

  df_token_hits %>%
    dplyr::count(token, name = "n_hit") %>%
    dplyr::mutate(
      callset = cc,
      total_calls = total_calls,
      FDR = n_hit / total_calls
    )

}) %>%
  dplyr::bind_rows()

platform_fdr %>% dplyr::select(callset, total_calls, token, n_hit, FDR)

```

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)

maf_map <- list(
  call_Illumina = "ILLUMINA_MAF",
  call_ppmSeq_multiread = "DUPLEX_PPMSEQ_MAF",
  call_ppmSeq_singleton_HC = "DUPLEX_PPMSEQ_MAF",
  call_UDSeq_multiread = "DUPLEX_UDSEQ_MAF",
  call_UDSeq_singleton = "DUPLEX_UDSEQ_MAF",
  call_NanoSeq_multiread = "DUPLEX_NANOSEQ_MAF",
  call_NanoSeq_singleton = "DUPLEX_NANOSEQ_MAF",
  call_HiDEFseq_double_stranded = "DUPLEX_HIDEFSEQ_MAF",
  call_HiDEFseq_single_stranded = "DUPLEX_HIDEFSEQ_MAF"
)

maf_plot_df <- lapply(names(maf_map), function(cc){

  maf_col <- maf_map[[cc]]

  nogerm %>%
    dplyr::filter(.data[[cc]] == 1) %>%
    dplyr::mutate(
      callset = stringr::str_replace(cc, "^call_", ""),
      group = ifelse(existing_in_unrelated == 1,
                     "unrelated_pon_removed",
                     "kept"),
      MAF = .data[[maf_col]]
    ) %>%
    dplyr::select(callset, group, MAF)

}) %>%
  dplyr::bind_rows() %>%
  dplyr::filter(!is.na(MAF), MAF > 0, MAF <= 0.3)

maf_plot_df$callset <- maf_plot_df$callset %>% factor(., levels=unique(maf_plot_df$callset))

ggplot(maf_plot_df,
       aes(x = MAF, fill = group)) +
  geom_histogram(aes(y = ..density..),
                 bins = 120,
                 alpha = 0.45,
                 position = "identity",
                 color = NA) +
  facet_wrap(~callset, scales = "free_y", ncol = 1) +
  scale_fill_manual(values = c(
    "kept" = "#4C72B0",
    "unrelated_pon_removed" = "#D62728"
  )) +
  theme_classic(base_size = 13) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(face = "bold"),
    legend.title = element_blank()
  ) +
  labs(
    x = "Platform-specific BAM MAF (0–0.3)",
    y = "Density"
  )
```

```{r}
library(dplyr)
library(stringr)

existing_cols_full <- colnames(nogerm)[
  str_detect(colnames(nogerm), "^is_existing_lower_ci_0001_") &
  !str_detect(colnames(nogerm), "30[xX]|100[xX]|200[xX]")
]



self_existing_map <- list(
  call_Illumina = "is_existing_lower_ci_0001_ILLUMINA",
  call_ppmSeq_multiread = "is_existing_lower_ci_0001_DUPLEX_PPMSEQ",
  call_ppmSeq_singleton_HC = "is_existing_lower_ci_0001_DUPLEX_PPMSEQ",
  call_UDSeq_multiread = "is_existing_lower_ci_0001_DUPLEX_UDSEQ",
  call_UDSeq_singleton = "is_existing_lower_ci_0001_DUPLEX_UDSEQ",
  call_NanoSeq_multiread = "is_existing_lower_ci_0001_DUPLEX_NANOSEQ",
  call_NanoSeq_singleton = "is_existing_lower_ci_0001_DUPLEX_NANOSEQ",
  call_HiDEFseq_double_stranded = "is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ",
  call_HiDEFseq_single_stranded = "is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ",
  call_PTA_singleton = "existing_any_PTA1_16"
)

overlap_df <- lapply(names(self_existing_map), function(cc){

  self_col <- self_existing_map[[cc]]

  # 모든 other BAM existing columns (자기 자신 제외)
  other_cols <- setdiff(existing_cols_full, self_col)

  # PTA는 existing_any_PTA1_16도 other evidence로 간주
  other_cols_all <- setdiff(c(existing_cols_full, "existing_any_PTA1_16"), self_col)

  nogerm %>%
    dplyr::filter(.data[[cc]] == 1) %>%
    dplyr::mutate(
      callset = stringr::str_replace(cc, "^call_", ""),
      group = ifelse(existing_in_unrelated == 1,
                     "unrelated_pon_removed",
                     "kept"),
      overlap_count = rowSums(dplyr::across(all_of(other_cols_all)) == 1,
                              na.rm = TRUE)
    ) %>%
    dplyr::select(callset, group, overlap_count)

}) %>%
  dplyr::bind_rows()



overlap_df$callset <- factor(
  overlap_df$callset,
  levels = stringr::str_replace(names(self_existing_map), "^call_", "")
)

ggplot(overlap_df,
       aes(x = group, y = overlap_count, fill = group)) +
  geom_violin(trim = TRUE, scale = "width", alpha = 0.7) +
  geom_boxplot(width = 0.15, outlier.size = 0.4, color = "black") +
  facet_wrap(~callset, ncol = 1, scales = "free_y") +
  scale_fill_manual(values = c(
    "kept" = "#4C72B0",
    "unrelated_pon_removed" = "#D62728"
  )) +
  theme_classic(base_size = 13) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(face = "bold"),
    legend.position = "none"
  ) +
  labs(
    x = NULL,
    y = "Number of other-platform BAM-level overlaps"
  )

```


```{r}
# Full-depth only: call_cols_full_depth
# - Source decomposition: platform-only vs multi-platform (within unrelated-hit variants)
# - FDR per token: fraction of ALL calls hit by each unrelated token

library(dplyr)
library(tidyr)
library(stringr)

tokens <- c(
  "DUPLEX_NANOSEQ_6566",
  "DUPLEX_PPMSEQ_6566",
  "DUPLEX_UDSEQ_6566",
  "HG19_ILLUMINA_7669",
  "HG19_PTA_6566_1",
  "HG19_PTA_6566_2",
  "HG19_PTA_8420_1",
  "HG19_PTA_8420_2",
  "ILLUMINA_6566_BLOOD",
  "ILLUMINA_6566_SPERM"
)

platform_group <- function(token_vec){
  groups <- character(0)

  if (any(token_vec %in% c("ILLUMINA_6566_BLOOD","ILLUMINA_6566_SPERM","HG19_ILLUMINA_7669")))
    groups <- c(groups, "ILLUMINA")
  if (any(token_vec == "DUPLEX_PPMSEQ_6566"))
    groups <- c(groups, "PPMSEQ")
  if (any(token_vec == "DUPLEX_UDSEQ_6566"))
    groups <- c(groups, "UDSEQ")
  if (any(token_vec == "DUPLEX_NANOSEQ_6566"))
    groups <- c(groups, "NANOSEQ")
  if (any(str_detect(token_vec, "^HG19_PTA_")))
    groups <- c(groups, "PTA")

  groups <- unique(groups)
  if (length(groups) == 0) return("None")
  if (length(groups) == 1) return(paste0(groups, "-only"))
  "Multi-platform"
}

analyze_callset_full_depth <- function(call_col){
  df_call <- nogerm %>% dplyr::filter(.data[[call_col]] == 1)
  total_calls <- nrow(df_call)

  # FDR per token (denominator = all calls in this callset)
  fdr_token <- df_call %>%
    tidyr::separate_rows(existing_in_unrelated_which, sep = ",") %>%
    dplyr::mutate(token = str_trim(existing_in_unrelated_which)) %>%
    dplyr::filter(token %in% tokens) %>%
    dplyr::count(token, name = "n_hit") %>%
    dplyr::mutate(
      callset = call_col,
      total_calls = total_calls,
      fdr = n_hit / total_calls
    )

  # Source decomposition among unrelated-hit variants
  source_decomp <- df_call %>%
    dplyr::filter(existing_in_unrelated_which != ".") %>%
    tidyr::separate_rows(existing_in_unrelated_which, sep = ",") %>%
    dplyr::mutate(token = str_trim(existing_in_unrelated_which)) %>%
    dplyr::group_by(chr, pos, ref, alt) %>%
    dplyr::summarise(tokens = list(token), .groups = "drop") %>%
    dplyr::mutate(source = vapply(tokens, platform_group, character(1))) %>%
    dplyr::count(source, name = "n") %>%
    dplyr::mutate(
      callset = call_col,
      frac_within_unrelated = n / sum(n)
    )

  list(fdr_token = fdr_token, source_decomp = source_decomp)
}

full_depth_results <- lapply(call_cols_full_depth, analyze_callset_full_depth)
names(full_depth_results) <- call_cols_full_depth

# Convenience binders
full_depth_fdr_token <- dplyr::bind_rows(lapply(full_depth_results, `[[`, "fdr_token"))
full_depth_source_decomp <- dplyr::bind_rows(lapply(full_depth_results, `[[`, "source_decomp"))

full_depth_fdr_token
full_depth_source_decomp


```


```{r}
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)

# ----------------------------
# 0) Factor orders (keep your call order)
# ----------------------------
call_order <- call_cols_full_depth

token_order <- c(
  "ILLUMINA_6566_BLOOD",
  "ILLUMINA_6566_SPERM",
  "HG19_ILLUMINA_7669",
  "DUPLEX_PPMSEQ_6566",
  "DUPLEX_UDSEQ_6566",
  "DUPLEX_NANOSEQ_6566",
  "HG19_PTA_6566_1",
  "HG19_PTA_6566_2",
  "HG19_PTA_8420_1",
  "HG19_PTA_8420_2"
)


# ----------------------------
# 2) Stacked bars: platform-only vs multi-platform composition (within unrelated-hit variants)
# ----------------------------
source_order <- c("ILLUMINA-only", "PPMSEQ-only", "UDSEQ-only", "NANOSEQ-only", "PTA-only", "Multi-platform")

source_plot_df <- full_depth_source_decomp %>%
  dplyr::mutate(
    callset = factor(callset, levels = call_order),
    callset_label = stringr::str_replace(callset, "^call_", ""),
    source = factor(source, levels = source_order)
  ) %>%
  dplyr::group_by(callset_label) %>%
  dplyr::mutate(frac_within_unrelated = n / sum(n)) %>%
  dplyr::ungroup()


source_plot_df$callset_label <- factor(source_plot_df$callset_label, levels=levels(source_plot_df$callset) %>% str_replace_all(., "^call_", "") %>% rev(.))
p_source <- ggplot(source_plot_df, aes(x = callset_label, y = frac_within_unrelated, fill = source)) +
  geom_col(width = 0.75, color = "white", linewidth = 0.2) +
  coord_flip() +
  theme_classic(base_size = 12) +
  theme(
    axis.title = element_blank(),
    legend.title = element_blank(),
    panel.grid = element_blank()
  ) +
  labs(y = "Fraction (within unrelated-hit variants)")

p_source


# ----------------------------
# 3) Optional: compact summary table to accompany plots (per callset totals)
# ----------------------------
summary_tbl <- full_depth_fdr_token %>%
  dplyr::group_by(callset) %>%
  dplyr::summarise(total_calls = dplyr::first(total_calls), .groups = "drop") %>%
  dplyr::left_join(
    full_depth_source_decomp %>%
      dplyr::group_by(callset) %>%
      dplyr::summarise(unrelated_hit_variants = sum(n), .groups = "drop"),
    by = "callset"
  ) %>%
  dplyr::mutate(
    callset = factor(callset, levels = call_order),
    callset_label = stringr::str_replace(as.character(callset), "^call_", ""),
    unrelated_hit_rate = unrelated_hit_variants / total_calls
  ) %>%
  dplyr::arrange(callset)

summary_tbl

```

```{r}

```


### Call-to-call

```{r}
suppressPackageStartupMessages({
  library(readr)
  library(dplyr)
  library(stringr)
  library(UpSetR)
})


# ---- Build matrix for UpSetR (missing cols -> 0) ----
make_mat_strata <- function(df, cols) {
  out <- lapply(cols, function(x) {
    if (x %in% colnames(df)) as.integer(df[[x]] %in% 1) else rep(0L, nrow(df))
  })
  mat <- as.data.frame(out, check.names = FALSE)
  colnames(mat) <- cols
  rownames(mat) <- NULL
  mat
}

mat_calls <- make_mat_strata(filt, call_cols_full_depth)

colnames(mat_calls) <- str_replace(colnames(mat_calls), "^call_", "")

# ---- Plot ----
upset(
  mat_calls,
  nsets = ncol(mat_calls),
  nintersects = 200,
  keep.order = TRUE,
  order.by = "freq",
  mainbar.y.label = "Intersection size",
  sets.x.label = "Set size"
)
```


### Call-to-BAM

```{r}

call_cols_full_depth <- c(
  "call_Illumina",
  "call_ppmSeq_multiread",
  #"call_ppmSeq_putative_multiread",
  "call_ppmSeq_singleton_HC",
  "call_UDSeq_multiread",
  "call_UDSeq_singleton",
  "call_NanoSeq_multiread",
  "call_NanoSeq_singleton",
  "call_HiDEFseq_double_stranded",
  "call_HiDEFseq_single_stranded",
  "call_PTA_singleton"
)

call_cols_200x <- c(
  "call_Illumina_200x",
  "call_ppmSeq_multiread_200x",
  "call_ppmSeq_singleton_HC_200x",
  "call_UDSeq_multiread",
  "call_UDSeq_singleton",
  "call_NanoSeq_multiread",
  "call_NanoSeq_singleton",
  "call_HiDEFseq_double_stranded",
  "call_HiDEFseq_single_stranded",
  "call_PTA_singleton"
)

call_cols_100x <- c(
  "call_Illumina_100x",
  "call_ppmSeq_multiread_100x",
  #"call_ppmSeq_putative_multiread_100x",
  "call_ppmSeq_singleton_HC_100x",
  "call_UDSeq_multiread",
  "call_UDSeq_singleton",
  "call_NanoSeq_multiread",
  "call_NanoSeq_singleton",
  "call_HiDEFseq_double_stranded",
  "call_HiDEFseq_single_stranded",
  "call_PTA_singleton"
)

call_cols_30x <- c(
  "call_Illumina_30x",
  "call_ppmSeq_multiread_30x",
  #"call_ppmSeq_putative_multiread_30x",
  "call_ppmSeq_singleton_HC_30x",
  "call_UDSeq_30x_multiread",
  "call_UDSeq_30x_singleton",
  "call_NanoSeq_multiread",
  "call_NanoSeq_singleton",
  "call_HiDEFseq_double_stranded",
  "call_HiDEFseq_single_stranded",
  "call_PTA_singleton"
)


call_cols_ppmSeq <- c(
  "call_ppmSeq_multiread",
  #"call_ppmSeq_putative_multiread",
  "call_ppmSeq_singleton_HC",
  "call_ppmSeq_multiread_30x",
  #"call_ppmSeq_putative_multiread_30x",
  "call_ppmSeq_singleton_HC_30x",
  "call_ppmSeq_multiread_100x",
  #"call_ppmSeq_putative_multiread_100x",
  "call_ppmSeq_singleton_HC_100x",
  "call_ppmSeq_multiread_200x",
  #"call_ppmSeq_putative_multiread_200x",
  "call_ppmSeq_singleton_HC_200x"
)

call_cols_illumina <- c(
  "call_Illumina",
  "call_Illumina_30x",
  "call_Illumina_100x",
  "call_Illumina_200x"
)


call_cols_full_depth_clonal_with_PTA <- c(
  "call_Illumina",
  "call_ppmSeq_multiread",
  "call_UDSeq_multiread",
  "call_NanoSeq_multiread",
  "call_HiDEFseq_double_stranded",
  "call_PTA_singleton"
)
call_cols_full_depth_clonal <- c(
  "call_Illumina",
  "call_ppmSeq_multiread",
  "call_UDSeq_multiread",
  "call_NanoSeq_multiread",
  "call_HiDEFseq_double_stranded"
)

call_cols_200x_clonal_with_PTA <- c(
  "call_Illumina_200x",
  "call_ppmSeq_multiread_200x",
  "call_UDSeq_multiread",
  "call_NanoSeq_multiread",
  "call_HiDEFseq_double_stranded",
  "call_PTA_singleton"
)


call_cols_200x_clonal <- c(
  "call_Illumina_200x",
  "call_ppmSeq_multiread_200x",
  "call_UDSeq_multiread",
  "call_NanoSeq_multiread",
  "call_HiDEFseq_double_stranded"
)


call_cols_100x_clonal_with_PTA <- c(
  "call_Illumina_100x",
  "call_ppmSeq_multiread_100x",
  "call_UDSeq_multiread",
  "call_NanoSeq_multiread",
  "call_HiDEFseq_double_stranded",
  "call_PTA_singleton"
)


call_cols_100x_clonal <- c(
  "call_Illumina_100x",
  "call_ppmSeq_multiread_100x",
  "call_UDSeq_multiread",
  "call_NanoSeq_multiread",
  "call_HiDEFseq_double_stranded"
)

call_cols_30x_clonal_with_PTA <- c(
  "call_Illumina_30x",
  "call_ppmSeq_multiread_30x",
  "call_UDSeq_30x_multiread",
  "call_NanoSeq_multiread",
  "call_HiDEFseq_double_stranded",
  "call_PTA_singleton"
)

call_cols_30x_clonal <- c(
  "call_Illumina_30x",
  "call_ppmSeq_multiread_30x",
  "call_UDSeq_30x_multiread",
  "call_NanoSeq_multiread",
  "call_HiDEFseq_double_stranded"
)


exist_cols_full <- c(
  "is_existing_lower_ci_0001_ILLUMINA",
  "is_existing_lower_ci_0001_DUPLEX_PPMSEQ",
  "is_existing_lower_ci_0001_DUPLEX_UDSEQ",
  "is_existing_lower_ci_0001_DUPLEX_NANOSEQ",
  "is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ",
  "existing_any_PTA1_16_singleton",
  "existing_any_PTA1_16_multiple"
)
exist_cols_full_altge1 <- c(
  "is_altge1_ILLUMINA",
  "is_altge1_DUPLEX_PPMSEQ",
  "is_altge1_DUPLEX_UDSEQ",
  "is_altge1_DUPLEX_NANOSEQ",
  "is_altge1_DUPLEX_HIDEFSEQ",
  "existing_any_PTA1_16_singleton",
  "existing_any_PTA1_16_multiple"
)

exist_cols_200x <- c(
  "is_existing_lower_ci_0001_ILLUMINA_200x",
  "is_existing_lower_ci_0001_DUPLEX_PPMSEQ_200X",
  "is_existing_lower_ci_0001_DUPLEX_UDSEQ",
  "is_existing_lower_ci_0001_DUPLEX_NANOSEQ",
  "is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ",
  "existing_any_PTA1_16_singleton",
  "existing_any_PTA1_16_multiple"
)

exist_cols_100x <- c(
  "is_existing_lower_ci_0001_ILLUMINA_100x",
  "is_existing_lower_ci_0001_DUPLEX_PPMSEQ_100X",
  "is_existing_lower_ci_0001_DUPLEX_UDSEQ",
  "is_existing_lower_ci_0001_DUPLEX_NANOSEQ",
  "is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ",
  "existing_any_PTA1_16_singleton",
  "existing_any_PTA1_16_multiple"
)

exist_cols_30x <- c(
  "is_existing_lower_ci_0001_ILLUMINA_30x",
  "is_existing_lower_ci_0001_DUPLEX_PPMSEQ_30X",
  "is_existing_lower_ci_0001_DUPLEX_UDSEQ_30X",
  "is_existing_lower_ci_0001_DUPLEX_NANOSEQ",
  "is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ",
  "existing_any_PTA1_16_singleton",
  "existing_any_PTA1_16_multiple"
)

get_self_exist_cols_full <- function(callset) {

  if (callset == "call_Illumina")
    return("is_existing_lower_ci_0001_ILLUMINA")

  if (callset %in% c("call_ppmSeq_multiread","call_ppmSeq_singleton_HC"))
    return("is_existing_lower_ci_0001_DUPLEX_PPMSEQ")

  if (callset %in% c("call_UDSeq_multiread","call_UDSeq_singleton"))
    return("is_existing_lower_ci_0001_DUPLEX_UDSEQ")

  if (callset %in% c("call_NanoSeq_multiread","call_NanoSeq_singleton"))
    return("is_existing_lower_ci_0001_DUPLEX_NANOSEQ")

  if (callset %in% c("call_HiDEFseq_double_stranded","call_HiDEFseq_single_stranded"))
    return("is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ")

  if (callset == "call_PTA_singleton")
    return(c("existing_any_PTA1_16_singleton","existing_any_PTA1_16_multiple"))

  character(0)
}


get_self_exist_cols_full_altge1 <- function(callset) {

  if (callset == "call_Illumina")
    return("is_altge1_ILLUMINA")

  if (callset %in% c("call_ppmSeq_multiread","call_ppmSeq_singleton_HC"))
    return("is_altge1_DUPLEX_PPMSEQ")

  if (callset %in% c("call_UDSeq_multiread","call_UDSeq_singleton"))
    return("is_altge1_DUPLEX_UDSEQ")

  if (callset %in% c("call_NanoSeq_multiread","call_NanoSeq_singleton"))
    return("is_altge1_DUPLEX_NANOSEQ")

  if (callset %in% c("call_HiDEFseq_double_stranded","call_HiDEFseq_single_stranded"))
    return("is_altge1_DUPLEX_HIDEFSEQ")

  if (callset == "call_PTA_singleton")
    return(c("existing_any_PTA1_16_singleton","existing_any_PTA1_16_multiple"))

  character(0)
}

get_self_exist_cols_30x <- function(callset) {

  if (callset == "call_Illumina_30x")
    return("is_existing_lower_ci_0001_ILLUMINA_30x")

  if (callset %in% c("call_ppmSeq_multiread_30x","call_ppmSeq_singleton_HC_30x"))
    return("is_existing_lower_ci_0001_DUPLEX_PPMSEQ_30X")

  if (callset %in% c("call_UDSeq_30x_multiread","call_UDSeq_30x_singleton"))
    return("is_existing_lower_ci_0001_DUPLEX_UDSEQ_30X")

  # NanoSeq / HiDEFseq: no downsample existence columns in your list
  if (callset %in% c("call_NanoSeq_multiread","call_NanoSeq_singleton"))
    return("is_existing_lower_ci_0001_DUPLEX_NANOSEQ")

  if (callset %in% c("call_HiDEFseq_double_stranded","call_HiDEFseq_single_stranded"))
    return("is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ")

  if (callset %in% c("call_PTA_singleton","call_PTA_multiple"))
    return(c("existing_any_PTA1_16_singleton","existing_any_PTA1_16_multiple"))

  character(0)
}

get_self_exist_cols_100x <- function(callset) {

  if (callset == "call_Illumina_100x")
    return("is_existing_lower_ci_0001_ILLUMINA_100x")

  if (callset %in% c("call_ppmSeq_multiread_100x","call_ppmSeq_singleton_HC_100x","call_ppmSeq_putative_multiread_100x"))
    return("is_existing_lower_ci_0001_DUPLEX_PPMSEQ_100X")

  # UDSeq/Nano/HiDEF: no downsample existence columns in your list
  if (callset %in% c("call_UDSeq_multiread","call_UDSeq_singleton"))
    return("is_existing_lower_ci_0001_DUPLEX_UDSEQ")

  if (callset %in% c("call_NanoSeq_multiread","call_NanoSeq_singleton"))
    return("is_existing_lower_ci_0001_DUPLEX_NANOSEQ")

  if (callset %in% c("call_HiDEFseq_double_stranded","call_HiDEFseq_single_stranded"))
    return("is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ")

  if (callset %in% c("call_PTA_singleton","call_PTA_multiple"))
    return(c("existing_any_PTA1_16_singleton","existing_any_PTA1_16_multiple"))

  character(0)
}
get_self_exist_cols_200x <- function(callset) {

  if (callset == "call_Illumina_200x")
    return("is_existing_lower_ci_0001_ILLUMINA_200x")

  if (callset %in% c("call_ppmSeq_multiread_200x","call_ppmSeq_singleton_HC_200x","call_ppmSeq_putative_multiread_200x"))
    return("is_existing_lower_ci_0001_DUPLEX_PPMSEQ_200X")

  # UDSeq/Nano/HiDEF: no downsample existence columns in your list
  if (callset %in% c("call_UDSeq_multiread","call_UDSeq_singleton"))
    return("is_existing_lower_ci_0001_DUPLEX_UDSEQ")

  if (callset %in% c("call_NanoSeq_multiread","call_NanoSeq_singleton"))
    return("is_existing_lower_ci_0001_DUPLEX_NANOSEQ")

  if (callset %in% c("call_HiDEFseq_double_stranded","call_HiDEFseq_single_stranded"))
    return("is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ")

  if (callset %in% c("call_PTA_singleton","call_PTA_multiple"))
    return(c("existing_any_PTA1_16_singleton","existing_any_PTA1_16_multiple"))

  character(0)
}

```


```{r}
library(dplyr)
library(stringr)
library(ggplot2)

call_cols <- call_cols_full_depth
#call_cols <- call_cols_30x
#call_cols <- call_cols_100x
#call_cols <- call_cols_200x

call_cols_clonal <- call_cols_full_depth_clonal
#call_cols_clonal <- call_cols_30x_clonal
#call_cols_clonal <- call_cols_100x_clonal
#call_cols_clonal <- call_cols_200x_clonal

#exist_cols <- exist_cols_full
exist_cols <- exist_cols_full_altge1
#exist_cols <- exist_cols_30x
#exist_cols <- exist_cols_100x
#exist_cols <- exist_cols_200x

#get_self_exist_cols <- get_self_exist_cols_full 
get_self_exist_cols <-get_self_exist_cols_full_altge1
#get_self_exist_cols <- get_self_exist_cols_30x 
#get_self_exist_cols <- get_self_exist_cols_100x 
#get_self_exist_cols <- get_self_exist_cols_200x 

call_levels <- str_replace(call_cols, "^call_", "")

# main
call_to_exist_prop <- lapply(call_cols, function(cc) {

  #df_call <- filt %>% dplyr::filter(.data[[cc]] == 1)
  df_call <- filt_altge1 %>% dplyr::filter(.data[[cc]] == 1)

  if (nrow(df_call) == 0) return(NULL)

  self_cols <- get_self_exist_cols(cc)
  other_exist_cols <- setdiff(exist_cols, self_cols)

  other_n <- rowSums(df_call[, other_exist_cols, drop=FALSE] == 1, na.rm = TRUE)

  overlap_class <- case_when(
    other_n == 0 ~ "private",
    other_n == 1 ~ "overlap_1",
    other_n >= 2 ~ "overlap_ge_2"
  )

  tibble(
    callset = str_replace(cc, "^call_", ""),
    overlap_class = overlap_class
  )
}) %>%
  bind_rows() %>%
  count(callset, overlap_class, name = "n") %>%
  group_by(callset) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup() %>%
  mutate(
    overlap_class = factor(overlap_class,
                           levels = c("private","overlap_1","overlap_ge_2")),
    callset = factor(callset,
                     levels = str_replace(call_cols, "^call_", ""))
  )



call_to_exist_prop$callset <- factor(
  call_to_exist_prop$callset,
  levels = rev(call_levels)  # coord_flip 때문에 reverse
)

call_to_exist_prop <- call_to_exist_prop %>%
  dplyr::mutate(
    overlap_class = factor(
      overlap_class,
      levels = c("overlap_ge_2", "overlap_1", "private")  # ← reverse
    )
  )

lab_df <- call_to_exist_prop %>%
  dplyr::mutate(overlap_class = as.character(overlap_class)) %>%
  dplyr::group_by(callset) %>%
  dplyr::arrange(match(overlap_class, c("private", "overlap_1", "overlap_ge_2")), .by_group = TRUE) %>%
  dplyr::mutate(
    y_center = cumsum(prop) - prop / 2,
    y_edge   = cumsum(prop),
    y_line   = y_edge + 0.020,
    y_label  = y_edge + 0.035,
    label = ifelse(
      prop >= 0.042,
      scales::percent(prop, accuracy = 0.1),
      scales::percent(prop, accuracy = 0.01)
    )
  ) %>%
  dplyr::ungroup()

ggplot(call_to_exist_prop,
       aes(x = callset, y = prop, fill = overlap_class)) +
  geom_col(width = 0.75) +

  # inside (>=1%)
  geom_text(
    data = lab_df %>% dplyr::filter(prop >= 0.042),
    aes(x = callset, y = y_center, label = label),
    inherit.aes = FALSE,
    size = 2.7
  ) +

  # line
  geom_segment(
    data = lab_df %>% dplyr::filter(prop < 0.042, prop > 0),
    aes(x = callset, xend = callset, y = y_edge, yend = y_line),
    inherit.aes = FALSE,
    linewidth = 0.3
  ) +

  # outside label (<1%)
  geom_text(
    data = lab_df %>% dplyr::filter(prop < 0.042, prop > 0),
    aes(x = callset, y = y_label, label = label),
    inherit.aes = FALSE,
    hjust = 0,
    size = 2.5
  ) +

  coord_flip(clip = "off") +
  scale_y_continuous(
    labels = scales::percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.10))
  ) +
  theme_classic() +
  labs(x = NULL, y = "Proportion (%)", fill = NULL)
```



```{r}
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(scales)




#call_cols <- call_cols_full_depth
#exist_cols <- exist_cols_full
#get_self_exist_cols <- get_self_exist_cols_full



# -----------------------------
# Name prettifier
# -----------------------------
pretty_callset <- function(x) {
  #x <- str_remove(x, "^call_")
  depth <- str_match(x, "_(\\d+[Xx])$")[,2]
  depth <- ifelse(is.na(depth), "", paste0("_", tolower(depth)))
  x0 <- str_remove(x, "_(\\d+[Xx])$")

  case_when(
#    x0 == "call_Illumina" ~ paste0("Illumina", depth),
#    x0 == "call_ppmSeq_multiread" ~ paste0("ppmSeq_multi", depth),
#    x0 == "call_ppmSeq_singleton_HC" ~ paste0("ppmSeq_single", depth),
#    x0 == "call_UDSeq_multiread" ~ "UDSeq_multi",
#    x0 == "call_UDSeq_singleton" ~ "UDSeq_single",
#    x0 == "call_UDSeq_30x_multiread" ~ paste0("UDSeq_multi", depth),
#    x0 == "call_UDSeq_30x_singleton" ~ paste0("UDSeq_single", depth),
#    x0 == "call_NanoSeq_multiread" ~ "NanoSeq_multi",
#    x0 == "call_NanoSeq_singleton" ~ "NanoSeq_single",
#    x0 == "call_HiDEFseq_double_stranded" ~ "HiDEFseq_ds",
#    x0 == "call_HiDEFseq_single_stranded" ~ "HiDEFseq_ss",
#    x0 == "call_PTA_singleton" ~ "PTA_single",
    TRUE ~ x
  )
}

pretty_exist <- function(x) {
  depth <- str_match(x, "_(\\d+[Xx])$")[,2]
  depth <- ifelse(is.na(depth), "", paste0("_", tolower(depth)))
  x0 <- str_remove(x, "_(\\d+[Xx])$")

  case_when(
    str_detect(x0, "ILLUMINA") ~ paste0("Illumina", depth),
    str_detect(x0, "DUPLEX_PPMSEQ") ~ paste0("ppmSeq", depth),
    str_detect(x0, "DUPLEX_UDSEQ") ~ paste0("UDSeq", depth),
    str_detect(x0, "DUPLEX_NANOSEQ") ~ paste0("NanoSeq", depth),
    str_detect(x0, "DUPLEX_HIDEFSEQ") ~ paste0("HiDEFseq", depth),
    str_detect(x0, "PTA") & str_detect(x0, "singleton") ~ "PTA_single",
    str_detect(x0, "PTA") & str_detect(x0, "multiple") ~ "PTA_multiple",
    TRUE ~ x
  )
}

# -----------------------------
# Core: for each callset, make "Total + overlaps with each other BAM evidence"
# Overlaps are NOT mutually exclusive (separate bars).
# -----------------------------
make_callset_overlap_long <- function(filt, call_cols, exist_cols, get_self_exist_cols,
                                      mode = c("prop","count")) {
  mode <- match.arg(mode)

  bind_rows(lapply(call_cols, function(cc) {

    df_call <- filt %>% dplyr::filter(.data[[cc]] == 1)
    total <- nrow(df_call)
    if (total == 0) return(NULL)

    self_cols <- get_self_exist_cols(cc)
    other_exist_cols <- setdiff(exist_cols, self_cols)

    other_n <- vapply(other_exist_cols, function(ec) {
      sum(df_call[[ec]] == 1, na.rm = TRUE)
    }, numeric(1))

    tibble(
      callset = cc,
      overlap_with = c("Total", other_exist_cols),
      n = c(total, other_n)
    ) %>%
      mutate(
        prop = n / total,
        value = if (mode == "prop") prop else n,
        callset_pretty = pretty_callset(callset),
        overlap_with_pretty = if_else(overlap_with == "Total",
                                      "Total (calls)",
                                      pretty_exist(overlap_with))
      )
  }))
}


make_callset_overlap_long <- function(filt, call_cols, exist_cols, get_self_exist_cols,
                                      mode = c("prop","count")) {
  mode <- match.arg(mode)

  bind_rows(lapply(call_cols, function(cc) {

    df_call <- filt %>% dplyr::filter(.data[[cc]] == 1)
    total <- nrow(df_call)
    if (total == 0) return(NULL)

    self_cols <- get_self_exist_cols(cc)
    other_exist_cols <- setdiff(exist_cols, self_cols)

    # Overlap with each other BAM (NOT mutually exclusive)
    other_n <- vapply(other_exist_cols, function(ec) {
      sum(df_call[[ec]] == 1, na.rm = TRUE)
    }, numeric(1))

    # Private: no other BAM evidence (excluding self)
    if (length(other_exist_cols) == 0) {
      private_n <- total
    } else {
      other_mat <- as.matrix(df_call[, other_exist_cols, drop = FALSE] == 1)
      other_mat[is.na(other_mat)] <- FALSE
      private_n <- sum(rowSums(other_mat) == 0)
    }

    tibble(
      callset = cc,
      overlap_with = c("Total", "Private", other_exist_cols),
      n = c(total, private_n, other_n)
    ) %>%
      mutate(
        prop = n / total,
        value = if (mode == "prop") prop else n,
        callset_pretty = pretty_callset(callset),
        overlap_with_pretty = case_when(
          overlap_with == "Total" ~ "Total (calls)",
          overlap_with == "Private" ~ "Private",
          TRUE ~ pretty_exist(overlap_with)
        )
      )
  }))
}

# -----------------------------
# Plot without tidytext::reorder_within
# - Order within each facet: Total first, then overlaps by value desc
# -----------------------------

# =========================================================
# 1) full depth: relative proportion (call + BAM both full depth)


long_full_rel <- make_callset_overlap_long(
  #filt = filt,
  filt = filt_altge1,
  call_cols = call_cols,
  exist_cols = exist_cols,
  get_self_exist_cols = get_self_exist_cols,
  mode = "prop"
)

tbl_full_rel <- long_full_rel %>%
  dplyr::select(callset_pretty, overlap_with_pretty, n, prop) %>%
  group_by(callset_pretty) %>%
  arrange(desc(prop), .by_group = TRUE) %>%
  ungroup()
tbl_full_rel

plot_callset_overlap <- function(long_df, mode = c("prop","count")) {
  mode <- match.arg(mode)

  dfp <- long_df %>%
    group_by(callset_pretty) %>%
    mutate(
      ord = if_else(overlap_with == "Total", Inf, value),
      overlap_with_pretty2 = paste0(overlap_with_pretty, "___", callset_pretty),
      label = if (mode == "prop") scales::percent(value, accuracy = 0.1) else as.character(n)
    ) %>%
    ungroup()

  # Global factor levels across all facets, respecting per-facet ordering (your original)
  lvl <- dfp %>%
    group_by(callset_pretty) %>%
    arrange(desc(ord), .by_group = TRUE) %>%
    summarize(levels = list(unique(overlap_with_pretty2)), .groups = "drop") %>%
    pull(levels) %>%
    unlist(use.names = FALSE)

  dfp$overlap_with_pretty2 <- factor(dfp$overlap_with_pretty2, levels = lvl)

  ggplot(dfp, aes(x = overlap_with_pretty2, y = value, fill = overlap_with_pretty)) +
    geom_col(width = 0.75) +
    geom_text(
      aes(label = label),
      hjust = -0.05,              # put label slightly to the right of bar end
      size = 3,
      show.legend = FALSE
    ) +
    coord_flip() +
    facet_wrap(~ callset_pretty, scales = "free_y", nrow = 2, ncol = 5) +
    scale_x_discrete(labels = function(x) sub("___.*$", "", x)) +
    scale_y_continuous(
      labels = if (mode == "prop") percent_format(accuracy = 1) else waiver(),
      expand = expansion(mult = c(0, 0.18))  # add headroom for labels (prevents axis collision)
    ) +
    theme_classic(base_size = 13) +
    theme(legend.position = "none") +
    labs(x = NULL, y = ifelse(mode == "prop", "Proportion within callset", "Count"))
}
# use as-is
long_full_rel$callset_pretty <- long_full_rel$callset_pretty %>% factor(., levels=unique(.))
plot_callset_overlap(long_full_rel, mode = "prop")


```




```{r}
#call_cols_clonal <- call_cols_full_depth_clonal
#exist_cols <- exist_cols_full
#get_self_exist_cols <- get_self_exist_cols_full 

call_levels <- str_replace(call_cols_clonal, "^call_", "")

call_to_exist_abs <- lapply(call_cols_clonal, function(cc) {

  df_call <- filt %>% dplyr::filter(.data[[cc]] == 1)
  if (nrow(df_call) == 0) return(NULL)

  self_cols <- get_self_exist_cols(cc)
  other_exist_cols <- setdiff(exist_cols, self_cols)

  other_n <- rowSums(df_call[, other_exist_cols, drop=FALSE] == 1, na.rm = TRUE)

  overlap_class <- case_when(
    other_n == 0 ~ "private",
    other_n == 1 ~ "overlap_1",
    other_n >= 2 ~ "overlap_ge_2"
  )

  tibble(
    callset = str_replace(cc, "^call_", ""),
    overlap_class = overlap_class
  )
}) %>%
  bind_rows() %>%
  count(callset, overlap_class, name = "n") %>%
  mutate(
    overlap_class = factor(overlap_class,
                           levels = c("private","overlap_1","overlap_ge_2")),
    callset = factor(callset,
                     levels = rev(call_levels))
  )

# absolute stacked
ggplot(call_to_exist_abs,
       aes(x = callset, y = n, fill = overlap_class)) +
  geom_col(width = 0.75) +
  coord_flip() +
  theme_classic() +
  labs(x = NULL, y = "Count", fill = NULL)


call_to_bam_overlap_tbl <- call_to_exist_abs %>%  mutate(
    callset = factor(callset, levels = rev(levels(callset)))
  ) %>%  arrange(callset, overlap_class) %>%
    group_by(callset) %>%
    mutate(prop = n / sum(n)) %>%
    ungroup()

call_to_bam_overlap_tbl
```






### Variant universe

```{r}

exist_cols <- exist_cols_full
#exist_cols <- exist_cols_30x
#exist_cols <- exist_cols_100x
#exist_cols <- exist_cols_200x


support_cols <- setdiff(exist_cols,
          c("existing_any_PTA1_16_singleton",
            "existing_any_PTA1_16_multiple"))


#filt2 <- filt %>%
filt2 <- filt %>%
  mutate(
    n_support = rowSums(
      dplyr::select(., dplyr::all_of(support_cols)) == 1,
      na.rm = TRUE
    )
  ) %>%
  dplyr::filter(n_support >= 1, n_support <= 5) %>%
  mutate(n_support = factor(n_support, levels = 1:5)) #%>%
  #filter(n_support<5)

med_tbl <- filt2 %>%
  group_by(n_support) %>%
  summarize(med = median(ILLUMINA_MAF, na.rm = TRUE), .groups = "drop")

ggplot(filt2, aes(x = n_support, y = ILLUMINA_MAF, fill = n_support)) +
  geom_violin(trim = TRUE, scale = "width", alpha = 0.5, color = NA) +
  geom_boxplot(width = 0.18, outlier.shape = NA, alpha = 0.95) +
  coord_flip() +
  theme_classic(base_size = 14) +
  theme(legend.position = "none") +
  labs(
    x = "# Platforms supporting variant",
    y = "Illumina MAF"
  ) +
   scale_y_continuous(
    limits = c(0, 0.3),
    breaks = seq(0, 0.3, by = 0.05),
    expand = expansion(mult = c(0.02, 0.18))
  )

stat_tbl <- filt2 %>%
  group_by(n_support) %>%
  summarize(
    mean_maf   = mean(ILLUMINA_MAF, na.rm = TRUE),
    median_maf = median(ILLUMINA_MAF, na.rm = TRUE),
    .groups = "drop"
  )
stat_tbl
```

```{r}
library(dplyr)
library(ggplot2)
library(scales)

exist_cols <- exist_cols_full_altge1
support_cols <- c(
  setdiff(exist_cols,
          c("existing_any_PTA1_16_singleton",
            "existing_any_PTA1_16_multiple"))
)

filt2 <- filt %>%
  mutate(
    n_support = rowSums(
      dplyr::select(., dplyr::all_of(support_cols)) == 1,
      na.rm = TRUE
    ),
    PTA_class = case_when(
      existing_any_PTA1_16_number == 0 ~ "no_PTA_BAM_support",
      existing_any_PTA1_16_number == 1 ~ "PTA_BAM_support_1",
      existing_any_PTA1_16_number >= 2 ~ "PTA_BAM_support_ge_2",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(PTA_class)) %>%
  filter(n_support >= 1, n_support <= 5) %>%
  mutate(
    n_support = factor(n_support, levels = 1:5),
    PTA_class = factor(PTA_class,
                       levels = c("no_PTA_BAM_support",
                                  "PTA_BAM_support_1",
                                  "PTA_BAM_support_ge_2"))
  )

plot_df <- filt2 %>%
  count(n_support, PTA_class, name = "n") %>%
  group_by(n_support) %>%
  mutate(frac = n / sum(n)) %>%
  ungroup()

ggplot(plot_df, aes(x = n_support, y = frac, fill = PTA_class)) +
  geom_col(width = 0.85) +
  coord_flip() +
  
  geom_text(
    aes(label = percent(frac, accuracy = 1)),
    position = position_stack(vjust = 0.5),  # center of each stack
    size = 4
  ) +
  
  theme_classic(base_size = 14) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "N_platform", y = "Fraction", fill = "PTA")

```


```{r}
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
exist_cols_30x_altge1 <- c("is_altge1_ILLUMINA_30x", "is_altge1_DUPLEX_PPMSEQ_30X", "is_altge1_DUPLEX_UDSEQ_30X", "is_altge1_DUPLEX_NANOSEQ", "is_altge1_DUPLEX_HIDEFSEQ", "existing_any_PTA1_16_singleton", "existing_any_PTA1_16_multiple")

#exist_cols <- exist_cols_full_altge1
exist_cols <- exist_cols_30x_altge1

#exist_cols <- exist_cols_30x
#exist_cols <- exist_cols_100x
#exist_cols <- exist_cols_200x

support_cols <- c(
  setdiff(exist_cols,
          c("existing_any_PTA1_16_singleton",
            "existing_any_PTA1_16_multiple"))
)

#support_cols <- filt %>% colnames(.) %>% .[str_detect(., "^is_existing")] %>% .[str_detect(., "ILLUMINA")]
# ---- settings ----
# map evidence col -> MAF col (adjust here if your column names differ)
# 1) evidence col -> raw platform token (remove prefix only)
#platform_raw <- stringr::str_remove(support_cols, "^is_existing_lower_ci_0001_")
platform_raw <- stringr::str_remove(support_cols, "^is_altge1_")

# 2) evidence col -> MAF col (rule: <platform_raw>_MAF)
# e.g. is_existing_lower_ci_0001_DUPLEX_PPMSEQ -> DUPLEX_PPMSEQ_MAF
evidence_to_maf <- stats::setNames(
  paste0(platform_raw, "_MAF"),
  support_cols
)

# 3) pretty platform label for facets (remove prefix + optional DUPLEX_)
pretty_plat <- function(x) {
  #stringr::str_remove(x, "^is_existing_lower_ci_0001_") %>%
  stringr::str_remove(x, "^is_altge1_") %>%
    stringr::str_remove("^DUPLEX_")
}

plats <- pretty_plat(support_cols)

# ---- build long MAF table (only variants with evidence in that platform) ----
maf_long <- bind_rows(lapply(support_cols, function(ec) {
  maf_col <- evidence_to_maf[[ec]]
  stopifnot(!is.null(maf_col))
  if (!maf_col %in% colnames(filt)) stop(paste("Missing MAF column:", maf_col))

  filt %>%
    transmute(
      platform = pretty_plat(ec),
      maf = .data[[maf_col]],
      has_evidence = .data[[ec]] == 1
    ) %>%
    filter(has_evidence, !is.na(maf))
}), .id = NULL) %>%
  select(-has_evidence)

# ---- make vid once ----
if (all(c("chr","pos","ref","alt") %in% colnames(filt))) {
  filt <- filt %>% mutate(vid = paste(chr, pos, ref, alt, sep=":"))
} else {
  filt <- filt %>% mutate(vid = row_number())
}

# ---- build long MAF table (only variants with evidence) ----
maf_long <- bind_rows(lapply(support_cols, function(ec) {
  maf_col <- evidence_to_maf[[ec]]
  if (!maf_col %in% colnames(filt)) stop(paste("Missing MAF column:", maf_col))

  filt %>%
    transmute(
      vid = vid,
      platform = pretty_plat(ec),
      maf = .data[[maf_col]],
      has_evidence = .data[[ec]] == 1
    ) %>%
    filter(has_evidence, !is.na(maf))
})) %>%
  dplyr::select(-has_evidence)

# ---- wide for pairwise (shared variants only; both platforms have evidence) ----
maf_wide <- maf_long %>%
  mutate(platform = factor(platform, levels = plats)) %>%
  tidyr::pivot_wider(names_from = platform, values_from = maf)

# ---- build all ordered pairs (x,y), drop self ----
pairs <- expand.grid(
  x = plats,
  y = plats,
  stringsAsFactors = FALSE
) %>%
  filter(x != y) %>%
  as_tibble()

# ---- collect pairwise points + stats ----
pair_df <- bind_rows(lapply(seq_len(nrow(pairs)), function(i) {
  x <- pairs$x[i]
  y <- pairs$y[i]

  d <- maf_wide %>%
    select(vid, all_of(c(x, y))) %>%
    rename(x_maf = !!x, y_maf = !!y) %>%
    filter(!is.na(x_maf), !is.na(y_maf))

  n <- nrow(d)
  if (n >= 3) {
    ct <- suppressWarnings(cor.test(d$x_maf, d$y_maf, method = "pearson"))
    r <- unname(ct$estimate)
    p <- ct$p.value
  } else {
    r <- NA_real_
    p <- NA_real_
  }

  d %>%
    mutate(
      x_plat = x,
      y_plat = y,
      n = n,
      r = r,
      p = p
    )
}))

# text labels per facet (one row each)
lab_df <- pair_df %>%
  distinct(x_plat, y_plat, n, r, p) %>%
  mutate(
    label = paste0(
      "n=", n,
      "\nr=", ifelse(is.na(r), "NA", sprintf("%.3f", r)),
      "\np=", ifelse(is.na(p), "NA", format(p, digits = 2, scientific = TRUE))
    ),
    x_pos = 0.29,
    y_pos = 0.29
  )

pair_df <- pair_df %>%
  mutate(
    x_plat = factor(x_plat, levels = plats),
    y_plat = factor(y_plat, levels = plats)
  )

lab_df <- lab_df %>%
  mutate(
    x_plat = factor(x_plat, levels = plats),
    y_plat = factor(y_plat, levels = plats)
  )


#pdf("ilmn_maf_maf.pdf")
ggplot(pair_df, aes(x = x_maf, y = y_maf)) +
  geom_point(size = 0.6, alpha = 0.75) +
  geom_abline(slope = 1, intercept = 0, linetype = 2) +
  geom_text(
    data = lab_df,
    aes(x = x_pos, y = y_pos, label = label),
    inherit.aes = FALSE,
    hjust = 1, vjust = 1,
    size = 3
  ) +
  facet_grid(y_plat ~ x_plat, drop = FALSE) +
  scale_x_continuous(limits = c(0, 0.3), breaks = seq(0, 0.3, by = 0.1)) +
  scale_y_continuous(limits = c(0, 0.3), breaks = seq(0, 0.3, by = 0.1)) +
  coord_fixed() +
  theme_classic(base_size = 12) +
  theme(
    strip.background = element_blank(),
    panel.spacing = unit(0.8, "lines")
  ) +
  labs(x = "MAF (x platform)", y = "MAF (y platform)")
#dev.off()
```


```{r}
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)

df_to_use_maf_maf <- rdselect_call1000

exist_cols <- exist_cols_full
#exist_cols <- exist_cols_30x
#exist_cols <- exist_cols_100x
#exist_cols <- exist_cols_200x

support_cols <- c(
  setdiff(exist_cols,
          c("existing_any_PTA1_16_singleton",
            "existing_any_PTA1_16_multiple"))
)

#support_cols <- filt %>% colnames(.) %>% .[str_detect(., "^is_existing")] %>% .[str_detect(., "ILLUMINA")]
# ---- settings ----
# map evidence col -> MAF col (adjust here if your column names differ)
# 1) evidence col -> raw platform token (remove prefix only)
platform_raw <- stringr::str_remove(support_cols, "^is_existing_lower_ci_0001_")

# 2) evidence col -> MAF col (rule: <platform_raw>_MAF)
# e.g. is_existing_lower_ci_0001_DUPLEX_PPMSEQ -> DUPLEX_PPMSEQ_MAF
evidence_to_maf <- stats::setNames(
  paste0(platform_raw, "_MAF"),
  support_cols
)

# 3) pretty platform label for facets (remove prefix + optional DUPLEX_)
pretty_plat <- function(x) {
  stringr::str_remove(x, "^is_existing_lower_ci_0001_") %>%
    stringr::str_remove("^DUPLEX_")
}

plats <- pretty_plat(support_cols)

# ---- build long MAF table (only variants with evidence in that platform) ----
maf_long <- bind_rows(lapply(support_cols, function(ec) {
  maf_col <- evidence_to_maf[[ec]]
  stopifnot(!is.null(maf_col))
  if (!maf_col %in% colnames(filt)) stop(paste("Missing MAF column:", maf_col))

  filt %>%
    transmute(
      platform = pretty_plat(ec),
      maf = .data[[maf_col]],
      has_evidence = .data[[ec]] == 1
    ) %>%
    filter(has_evidence, !is.na(maf))
}), .id = NULL) %>%
  select(-has_evidence)

# ---- make vid once ----
if (all(c("chr","pos","ref","alt") %in% colnames(filt))) {
  filt <- filt %>% mutate(vid = paste(chr, pos, ref, alt, sep=":"))
} else {
  filt <- filt %>% mutate(vid = row_number())
}

# ---- build long MAF table (only variants with evidence) ----
maf_long <- bind_rows(lapply(support_cols, function(ec) {
  maf_col <- evidence_to_maf[[ec]]
  if (!maf_col %in% colnames(filt)) stop(paste("Missing MAF column:", maf_col))

  filt %>%
    transmute(
      vid = vid,
      platform = pretty_plat(ec),
      maf = .data[[maf_col]],
      has_evidence = .data[[ec]] == 1
    ) %>%
    filter(has_evidence, !is.na(maf))
})) %>%
  dplyr::select(-has_evidence)

# ---- wide for pairwise (shared variants only; both platforms have evidence) ----
maf_wide <- maf_long %>%
  mutate(platform = factor(platform, levels = plats)) %>%
  tidyr::pivot_wider(names_from = platform, values_from = maf)

# ---- build all ordered pairs (x,y), drop self ----
pairs <- expand.grid(
  x = plats,
  y = plats,
  stringsAsFactors = FALSE
) %>%
  filter(x != y) %>%
  as_tibble()

# ---- collect pairwise points + stats ----
pair_df <- bind_rows(lapply(seq_len(nrow(pairs)), function(i) {
  x <- pairs$x[i]
  y <- pairs$y[i]

  d <- maf_wide %>%
    select(vid, all_of(c(x, y))) %>%
    rename(x_maf = !!x, y_maf = !!y) %>%
    filter(!is.na(x_maf), !is.na(y_maf))

  n <- nrow(d)
  if (n >= 3) {
    ct <- suppressWarnings(cor.test(d$x_maf, d$y_maf, method = "pearson"))
    r <- unname(ct$estimate)
    p <- ct$p.value
  } else {
    r <- NA_real_
    p <- NA_real_
  }

  d %>%
    mutate(
      x_plat = x,
      y_plat = y,
      n = n,
      r = r,
      p = p
    )
}))

# text labels per facet (one row each)
lab_df <- pair_df %>%
  distinct(x_plat, y_plat, n, r, p) %>%
  mutate(
    label = paste0(
      "n=", n,
      "\nr=", ifelse(is.na(r), "NA", sprintf("%.3f", r)),
      "\np=", ifelse(is.na(p), "NA", format(p, digits = 2, scientific = TRUE))
    ),
    x_pos = 0.29,
    y_pos = 0.29
  )

pair_df <- pair_df %>%
  mutate(
    x_plat = factor(x_plat, levels = plats),
    y_plat = factor(y_plat, levels = plats)
  )

lab_df <- lab_df %>%
  mutate(
    x_plat = factor(x_plat, levels = plats),
    y_plat = factor(y_plat, levels = plats)
  )


ggplot(pair_df, aes(x = x_maf, y = y_maf)) +
  geom_point(size = 0.6, alpha = 0.75) +
  geom_abline(slope = 1, intercept = 0, linetype = 2) +
  geom_text(
    data = lab_df,
    aes(x = x_pos, y = y_pos, label = label),
    inherit.aes = FALSE,
    hjust = 1, vjust = 1,
    size = 3
  ) +
  facet_grid(y_plat ~ x_plat, drop = FALSE) +
  scale_x_continuous(limits = c(0, 0.3), breaks = seq(0, 0.3, by = 0.1)) +
  scale_y_continuous(limits = c(0, 0.3), breaks = seq(0, 0.3, by = 0.1)) +
  coord_fixed() +
  theme_classic(base_size = 12) +
  theme(
    strip.background = element_blank(),
    panel.spacing = unit(0.8, "lines")
  ) +
  labs(x = "MAF (x platform)", y = "MAF (y platform)")
#dev.off()
```


```{r}
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)

#exist_cols <- exist_cols_full
#exist_cols <- exist_cols_30x
#exist_cols <- exist_cols_100x
#exist_cols <- exist_cols_200x

#support_cols <- c(
#  setdiff(exist_cols,
#          c("existing_any_PTA1_16_singleton",
#            "existing_any_PTA1_16_multiple"))
#)

pretty_plat <- function(x) {
  x %>%
#    str_remove("^is_existing_lower_ci_0001_") %>%
    str_remove("^is_altge1_") %>%

    str_remove("^DUPLEX_")
}

plats <- pretty_plat(support_cols)
plats <- plats %>% factor(., levels=(.))
#plats <- factor(plats, levels = c("ILLUMINA","PPMSEQ","UDSEQ","NANOSEQ","HIDEFSEQ")) %>% as.character()

# binary matrix: rows=variants, cols=platforms
mat_bam <- as.data.frame(lapply(support_cols, function(x) as.integer(filt[[x]] %in% 1)),
                         check.names = FALSE)
colnames(mat_bam) <- pretty_plat(support_cols)
mat_bam <- mat_bam[rowSums(mat_bam) >= 2, , drop = FALSE]
# enforce order if present
keep <- intersect(levels(plats), colnames(mat_bam))
mat_bam <- mat_bam[, keep, drop = FALSE]

# Jaccard function
jaccard <- function(a, b) {
  a <- as.integer(a != 0)
  b <- as.integer(b != 0)
  inter <- sum(a & b)
  uni <- sum(a | b)
  if (uni == 0) NA_real_ else inter / uni
}

# compute full matrix (exclude self-self)
J <- matrix(NA_real_, ncol(mat_bam), ncol(mat_bam),
            dimnames = list(colnames(mat_bam), colnames(mat_bam)))

for (i in seq_len(ncol(mat_bam))) {
  for (j in seq_len(ncol(mat_bam))) {
    if (i == j) {
      J[i, j] <- NA_real_
    } else {
      J[i, j] <- jaccard(mat_bam[[i]], mat_bam[[j]])
    }
  }
}
# long for ggplot
J_long <- as.data.frame(J) %>%
  tibble::rownames_to_column("y") %>%
  pivot_longer(-y, names_to = "x", values_to = "jaccard") %>%
  mutate(
    x = factor(x, levels = colnames(J)),
    y = factor(y, levels = rev(colnames(J)))
  )
J_long <- J_long %>% dplyr::filter(x != y)

ggplot(J_long, aes(x = x, y = y, fill = jaccard)) +
  geom_tile(color = "grey85", linewidth = 0.4) +
  geom_text(aes(label = sprintf("%.2f", jaccard)), size = 3, color = "black") +
  coord_fixed() +
  theme_classic(base_size = 14) +
  theme(
    axis.title = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  scale_fill_gradient(low = "white", high = "steelblue", limits = c(0, 1)) +
  labs(fill = "Jaccard")
```

```{r}
library(dplyr)
library(stringr)
library(tidyr)

# ---- inputs ----
call_cols_ppmSeq_multiread <- c(
  "call_ppmSeq_multiread",
  "call_ppmSeq_multiread_30x",
  "call_ppmSeq_multiread_100x",
  "call_ppmSeq_multiread_200x"
)

call_cols_ppmSeq_singleton <- c(
  "call_ppmSeq_singleton_HC",
  "call_ppmSeq_singleton_HC_30x",
  "call_ppmSeq_singleton_HC_100x",
  "call_ppmSeq_singleton_HC_200x"
)

# truth axes: exclude PPM evidence + exclude PTA
#truth_cols <- c(
#  "is_existing_lower_ci_0001_ILLUMINA",
#  "is_existing_lower_ci_0001_DUPLEX_UDSEQ",
#  "is_existing_lower_ci_0001_DUPLEX_NANOSEQ",
#  "is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ"
#)

truth_cols <- c(
  "is_existing_lower_ci_0001_ILLUMINA",
  "is_existing_lower_ci_0001_DUPLEX_UDSEQ",
  "is_existing_lower_ci_0001_DUPLEX_NANOSEQ",
  "is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ"
)


# ---- helper: parse depth label ----
depth_label <- function(call_col) {
  if (str_detect(call_col, "_30x$"))  return("30x")
  if (str_detect(call_col, "_100x$")) return("100x")
  if (str_detect(call_col, "_200x$")) return("200x")
  "Full"
}

type_label <- function(call_col) {
  if (str_detect(call_col, "multiread")) "multiread" else "singleton"
}

# ---- compute truth once ----
df0 <- filt %>%
  mutate(
    truth = rowSums(dplyr::across(dplyr::all_of(truth_cols), ~ .x == 1), na.rm = TRUE) >= 1
  )

# ---- compute confusion + metrics for a set of call cols ----
score_calls <- function(call_cols) {
  bind_rows(lapply(call_cols, function(cc) {
    pred <- df0[[cc]] == 1

    TP <- sum(pred &  df0$truth, na.rm = TRUE)
    FP <- sum(pred & !df0$truth, na.rm = TRUE)
    FN <- sum(!pred &  df0$truth, na.rm = TRUE)
    TN <- sum(!pred & !df0$truth, na.rm = TRUE)

    precision <- if ((TP + FP) == 0) NA_real_ else TP / (TP + FP)
    recall    <- if ((TP + FN) == 0) NA_real_ else TP / (TP + FN)
    f1        <- if (is.na(precision) || is.na(recall) || (precision + recall) == 0) NA_real_
                 else 2 * precision * recall / (precision + recall)

    tibble(
      call_col = cc,
      depth = depth_label(cc),
      ppm_type = type_label(cc),
      TP = TP, FP = FP, FN = FN, TN = TN,
      precision = precision,
      recall = recall,
      F1 = f1
    )
  }))
}

res <- bind_rows(
  score_calls(call_cols_ppmSeq_multiread),
  score_calls(call_cols_ppmSeq_singleton)
) %>%
  mutate(depth = factor(depth, levels = c("30x","100x","200x","Full"))) %>%
  arrange(ppm_type, depth)

res


```


```{r}
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)

# -----------------------------
# helper
# -----------------------------
depth_label <- function(call_col) {
  if (str_detect(call_col, "_30x$"))  return("30x")
  if (str_detect(call_col, "_100x$")) return("100x")
  if (str_detect(call_col, "_200x$")) return("200x")
  "Full"
}

type_label <- function(call_col) {
  if (str_detect(call_col, "multiread")) "multiread" else "singleton_HC"
}

# -----------------------------
# truth definitions
# -----------------------------
truth_map <- list(
  Any4 = c("is_existing_lower_ci_0001_ILLUMINA",
           "is_existing_lower_ci_0001_DUPLEX_UDSEQ",
           "is_existing_lower_ci_0001_DUPLEX_NANOSEQ",
           "is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ"),
  Illumina = c("is_existing_lower_ci_0001_ILLUMINA"),
  UDSeq = c("is_existing_lower_ci_0001_DUPLEX_UDSEQ"),
  NanoSeq = c("is_existing_lower_ci_0001_DUPLEX_NANOSEQ"),
  HiDEFseq = c("is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ")
)

call_cols <- c(call_cols_ppmSeq_multiread,
               call_cols_ppmSeq_singleton)

# -----------------------------
# compute TP / FP only
# -----------------------------
res_prec <- bind_rows(lapply(call_cols, function(cc) {

  pred <- filt[[cc]] == 1
  n_calls <- sum(pred, na.rm = TRUE)

  bind_rows(lapply(names(truth_map), function(tname) {

    truth_mat <- filt[, truth_map[[tname]], drop = FALSE]
    truth <- rowSums(truth_mat == 1, na.rm = TRUE) >= 1

    TP <- sum(pred & truth, na.rm = TRUE)
    FP <- sum(pred & !truth, na.rm = TRUE)

    precision <- if ((TP+FP)==0) NA_real_ else TP/(TP+FP)

    tibble(
      call_col = cc,
      depth = depth_label(cc),
      ppm_type = type_label(cc),
      truth_def = tname,
      TP = TP,
      FP = FP,
      n_calls = TP + FP,
      precision = precision,
      FDR = 1 - precision
    )

  }))

})) %>%
  mutate(
    depth = factor(depth, levels = c("30x","100x","200x","Full"))
  )

# -----------------------------
# 1️⃣ Precision line plot
# -----------------------------
ggplot(res_prec,
       aes(x=depth, y=precision,
           color=truth_def, group=truth_def)) +
  geom_line(linewidth=0.9) +
  geom_point(size=2.5) +
  facet_wrap(~ppm_type) +
  theme_classic(base_size=14) +
  labs(y="Precision", x="Depth", color="Truth")

# -----------------------------
# 3️⃣ TP count
# -----------------------------
ggplot(res_prec,
       aes(x=depth, y=TP,
           color=truth_def, group=truth_def)) +
  geom_line(linewidth=0.9) +
  geom_point(size=2.5) +
  facet_wrap(~ppm_type) +
  theme_classic(base_size=14) +
  labs(y="BAM-level validation count (TP)", x="Depth", color="Truth")

```


```{r}
library(dplyr)
library(ggplot2)
library(scales)

# ---- depth tick labels with (calls=...) for each ppm_type ----
tick_calls <- res_prec %>%
  distinct(depth, ppm_type, n_calls) %>%
  arrange(ppm_type, depth) %>%
  mutate(tick = paste0(as.character(depth), "\n(calls=", n_calls, ")"))

tick_map <- tick_calls %>%
  select(ppm_type, depth, tick) %>%
  tidyr::pivot_wider(names_from = depth, values_from = tick)

make_depth_ticks <- function(ppm_type) {
  z <- tick_calls %>% filter(ppm_type == !!ppm_type)
  setNames(z$tick, z$depth)
}

# ---- common theme: add grid (not too heavy) ----
theme_grid <- function() {
  theme_classic(base_size = 14) +
    theme(
      panel.grid.major = element_line(color = "grey88", linewidth = 0.4),
      panel.grid.minor = element_line(color = "grey93", linewidth = 0.25),
      axis.text.x = element_text(lineheight = 0.95)
    )
}

plot_precision <- function(df, ppm_type_label, upper=1, step=0.1) {
  depth_ticks <- make_depth_ticks(ppm_type_label)

  ggplot(df %>% filter(ppm_type == ppm_type_label),
         aes(x = depth, y = precision, color = truth_def, group = truth_def)) +
    geom_line(linewidth = 0.9) +
    geom_point(size = 2.3) +
    scale_x_discrete(labels = depth_ticks) +
    scale_y_continuous(limits = c(0, upper),
                        breaks = seq(0, upper, by = step),
                        labels = label_number(accuracy = 0.01)) +
    labs(
      title = paste0("ppmSeq ", ppm_type_label, " - Precision"),
      x = "Depth",
      y = "Precision",
      color = "Truth"
    ) +
    theme_grid()
}

plot_tp <- function(df, ppm_type_label) {
  depth_ticks <- make_depth_ticks(ppm_type_label)

  ggplot(df %>% filter(ppm_type == ppm_type_label),
         aes(x = depth, y = TP, color = truth_def, group = truth_def)) +
    geom_line(linewidth = 0.9) +
    geom_point(size = 2.3) +
    scale_x_discrete(labels = depth_ticks) +
    labs(
      title = paste0("ppmSeq ", ppm_type_label, " - TP"),
      x = "Depth",
      y = "BAM-level validation count (TP)",
      color = "Truth"
    ) +
    theme_grid()
}

# ---- 4 plots ----
p_prec_multi <- plot_precision(res_prec, "multiread", 0.5, 0.1)
p_tp_multi   <- plot_tp(res_prec, "multiread")
p_prec_sing  <- plot_precision(res_prec, "singleton_HC", 0.1, 0.02)
p_tp_sing    <- plot_tp(res_prec, "singleton_HC")

p_prec_multi
p_tp_multi
p_prec_sing
p_tp_sing
```

```{r}
library(dplyr)
library(stringr)
library(ggplot2)

depth_label <- function(call_col) {
  if (str_detect(call_col, "_30x$"))  return("30x")
  if (str_detect(call_col, "_100x$")) return("100x")
  if (str_detect(call_col, "_200x$")) return("200x")
  "Full"
}

# ----- multiread only -----
pta_m <- bind_rows(lapply(call_cols_ppmSeq_multiread, function(cc) {
  filt %>%
    filter(.data[[cc]] == 1,
           existing_any_PTA1_16_number >= 1) %>%
    transmute(
      depth = depth_label(cc),
      pta_n = existing_any_PTA1_16_number
    )
}))

pta_m <- pta_m %>%
  mutate(
    depth = factor(depth, levels = c("30x","100x","200x","Full")),
    pta_n = factor(pta_n)
  )

pta_count <- pta_m %>%
  count(depth, pta_n)

ggplot(pta_count,
       aes(x = pta_n, y = n, fill = depth)) +
  geom_col(width = 0.8, color = "grey30") +
  facet_grid(depth ~ ., scales = "free_y") +
  scale_fill_manual(values = c(
    "30x"  = "#4E79A7",
    "100x" = "#F28E2B",
    "200x" = "#59A14F",
    "Full" = "#E15759"
  )) + scale_y_continuous(breaks = function(x) pretty(x))+
  theme_classic(base_size = 14) +
  theme(
    panel.grid.major.y = element_line(color="grey88", linewidth=0.4),
    panel.grid.minor.y = element_line(color="grey93", linewidth=0.25),
    legend.position = "none"
  ) +
  labs(
    x = "existing_any_PTA1_16_number (≥1)",
    y = "Absolute count"
  )

```




```{r}
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(scales)

depth_label <- function(call_col) {
  if (str_detect(call_col, "_30x$"))  return("30x")
  if (str_detect(call_col, "_100x$")) return("100x")
  if (str_detect(call_col, "_200x$")) return("200x")
  "Full"
}

# ---- truth definitions (your version) ----
truth_map <- list(
  Any4 = c("is_existing_lower_ci_0001_DUPLEX_PPMSEQ",
           "is_existing_lower_ci_0001_DUPLEX_UDSEQ",
           "is_existing_lower_ci_0001_DUPLEX_NANOSEQ",
           "is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ"),
  PPMSEQ = c("is_existing_lower_ci_0001_DUPLEX_PPMSEQ"),
  UDSeq = c("is_existing_lower_ci_0001_DUPLEX_UDSEQ"),
  NanoSeq = c("is_existing_lower_ci_0001_DUPLEX_NANOSEQ"),
  HiDEFseq = c("is_existing_lower_ci_0001_DUPLEX_HIDEFSEQ")
)

call_cols <- call_cols_illumina  # Illumina only

# ---- compute TP/FP/precision per depth ----
res_prec <- bind_rows(lapply(call_cols, function(cc) {

  pred <- filt[[cc]] == 1
  n_calls <- sum(pred, na.rm = TRUE)

  bind_rows(lapply(names(truth_map), function(tname) {

    truth_mat <- filt[, truth_map[[tname]], drop = FALSE]
    truth <- rowSums(truth_mat == 1, na.rm = TRUE) >= 1

    TP <- sum(pred & truth, na.rm = TRUE)
    FP <- sum(pred & !truth, na.rm = TRUE)

    precision <- if ((TP+FP)==0) NA_real_ else TP/(TP+FP)

    tibble(
      call_col = cc,
      depth = depth_label(cc),
      truth_def = tname,
      TP = TP,
      FP = FP,
      n_calls = TP + FP,
      precision = precision,
      FDR = 1 - precision
    )
  }))
})) %>%
  mutate(depth = factor(depth, levels = c("30x","100x","200x","Full"))) %>%
  arrange(depth, truth_def)

# ---- depth ticks: include calls per depth (Illumina only) ----
tick_calls <- res_prec %>%
  distinct(depth, n_calls) %>%
  arrange(depth) %>%
  mutate(tick = paste0(as.character(depth), "\n(calls=", n_calls, ")"))

depth_ticks <- setNames(tick_calls$tick, tick_calls$depth)

# ---- theme with grid ----
theme_grid <- function() {
  theme_classic(base_size = 14) +
    theme(
      panel.grid.major = element_line(color = "grey88", linewidth = 0.4),
      panel.grid.minor = element_line(color = "grey93", linewidth = 0.25),
      axis.text.x = element_text(lineheight = 0.95)
    )
}

# ---- plots (Illumina only) ----
plot_precision_illumina <- function(df, upper = 1, step = 0.1) {
  ggplot(df, aes(x = depth, y = precision, color = truth_def, group = truth_def)) +
    geom_line(linewidth = 0.9) +
    geom_point(size = 2.3) +
    scale_x_discrete(labels = depth_ticks) +
    scale_y_continuous(
      limits = c(0, upper),
      breaks = seq(0, upper, by = step),
      labels = label_number(accuracy = 0.01)
    ) +
    labs(
      title = "Illumina calls - Precision",
      x = "Depth",
      y = "Precision",
      color = "Truth"
    ) +
    theme_grid()
}

plot_tp_illumina <- function(df) {
  ggplot(df, aes(x = depth, y = TP, color = truth_def, group = truth_def)) +
    geom_line(linewidth = 0.9) +
    geom_point(size = 2.3) +
    scale_x_discrete(labels = depth_ticks) +
    labs(
      title = "Illumina calls - TP",
      x = "Depth",
      y = "BAM-level validation count (TP)",
      color = "Truth"
    ) +
    theme_grid()
}

p_prec <- plot_precision_illumina(res_prec, upper = 0.5, step = 0.05)
p_tp   <- plot_tp_illumina(res_prec)

p_prec
p_tp
```

```{r}
library(dplyr)
library(stringr)
library(ggplot2)

depth_label <- function(call_col) {
  if (str_detect(call_col, "_30x$"))  return("30x")
  if (str_detect(call_col, "_100x$")) return("100x")
  if (str_detect(call_col, "_200x$")) return("200x")
  "Full"
}

# ----- multiread only -----
pta_m <- bind_rows(lapply(call_cols_illumina, function(cc) {
  filt %>%
    filter(.data[[cc]] == 1,
           existing_any_PTA1_16_number >= 1) %>%
    transmute(
      depth = depth_label(cc),
      pta_n = existing_any_PTA1_16_number
    )
}))

pta_m <- pta_m %>%
  mutate(
    depth = factor(depth, levels = c("30x","100x","200x","Full")),
    pta_n = factor(pta_n)
  )

pta_count <- pta_m %>%
  count(depth, pta_n)

ggplot(pta_count,
       aes(x = pta_n, y = n, fill = depth)) +
  geom_col(width = 0.8, color = "grey30") +
  facet_grid(depth ~ ., scales = "free_y") +
  scale_fill_manual(values = c(
    "30x"  = "#4E79A7",
    "100x" = "#F28E2B",
    "200x" = "#59A14F",
    "Full" = "#E15759"
  )) + scale_y_continuous(breaks = function(x) pretty(x))+
  theme_classic(base_size = 14) +
  theme(
    panel.grid.major.y = element_line(color="grey88", linewidth=0.4),
    panel.grid.minor.y = element_line(color="grey93", linewidth=0.25),
    legend.position = "none"
  ) +
  labs(
    x = "existing_any_PTA1_16_number (≥1)",
    y = "Absolute count"
  )


```

```{r}
library(dplyr)
library(stringr)

depth_label <- function(call_col) {
  if (str_detect(call_col, "_30x$"))  return("30x")
  if (str_detect(call_col, "_100x$")) return("100x")
  if (str_detect(call_col, "_200x$")) return("200x")
  "Full"
}

amp_col <- "AmpliSeq"  # <-- 실제 컬럼명으로 수정

amp_tbl <- bind_rows(lapply(call_cols_illumina, function(cc) {
  filt %>%
    summarise(
      depth = depth_label(cc),
      n_calls = sum(.data[[cc]] == 1, na.rm = TRUE),
      n_amp_plus = sum(.data[[cc]] == 1 & .data[[amp_col]] == 1, na.rm = TRUE)
    )
})) %>%
  mutate(
    depth = factor(depth, levels = c("30x","100x","200x","Full")),
    frac = n_amp_plus / n_calls
  ) %>%
  arrange(depth)

amp_tbl

```


```{r}

library(readr)
library(dplyr)
library(pscl)
library(ggplot2)
library(tidyr)

# ---- load ----
hist_df <- read_tsv("coverage/ppmseq_7614_coverage_hist.merged.tsv",
                    show_col_types = FALSE) %>%
  dplyr::select(depth, count) %>%
  mutate(depth = as.integer(depth),
         count = as.numeric(count)) %>%
  arrange(depth)

total <- sum(hist_df$count)
zero_count <- hist_df$count[match(0L, hist_df$depth)]
if (is.na(zero_count)) zero_count <- 0
obs_coverage <- 1 - (zero_count / total)

# ---- fit ----
dat <- hist_df %>% mutate(y = depth, w = count)

m_zip  <- zeroinfl(y ~ 1 | 1, data = dat, weights = w, dist = "poisson")
m_zinb <- zeroinfl(y ~ 1 | 1, data = dat, weights = w, dist = "negbin")

zip_pi     <- plogis(unname(m_zip$coefficients$zero[1]))
zip_lambda <- exp(unname(m_zip$coefficients$count[1]))

zinb_pi    <- plogis(unname(m_zinb$coefficients$zero[1]))
zinb_mu    <- exp(unname(m_zinb$coefficients$count[1]))
zinb_theta <- m_zinb$theta
zinb_alpha <- 1 / zinb_theta
# ---- pmf ----
pmf_zip <- function(k, pi, lambda){
  ifelse(k == 0,
         pi + (1 - pi) * dpois(0, lambda),
         (1 - pi) * dpois(k, lambda))
}

pmf_zinb <- function(k, pi, mu, theta){
  ifelse(k == 0,
         pi + (1 - pi) * dnbinom(0, mu = mu, size = theta),
         (1 - pi) * dnbinom(k, mu = mu, size = theta))
}

max_depth <- max(hist_df$depth)
ks <- 0:max_depth

pred_df <- tibble(
  depth = ks,
  ZIP  = pmf_zip(ks, zip_pi, zip_lambda),
  ZINB = pmf_zinb(ks, zinb_pi, zinb_mu, zinb_theta)
) %>%
  pivot_longer(cols = c(ZIP, ZINB),
               names_to = "model",
               values_to = "prob")

plot_df <- hist_df %>%
  mutate(frac = count / sum(count))

ggplot(plot_df, aes(depth, frac)) +
  geom_col(fill="grey80") +
  geom_line(data=pred_df,
            aes(depth, prob, colour=model),
            linewidth=1) +
  scale_y_continuous(labels=function(x) sprintf("%.0f%%",100*x)) +
  labs(title="ZIP vs ZINB fit",
       subtitle=paste0(
         "ZIP: pi=",round(zip_pi,3),
         ", lambda=",round(zip_lambda,2),
         " | ZINB: pi=",round(zinb_pi,3),
         ", mu=",round(zinb_mu,2),
         ", alpha=",round(zinb_alpha,3),
         " | coverage=",round(100*obs_coverage,1),"%"
       ),
       x="Depth",
       y="Fraction of loci") +
  theme_bw()

```

```{r}
library(readr)
library(dplyr)
library(glmmTMB)

hist_df <- read_tsv("coverage/ppmseq_7614_coverage_hist.merged.tsv",
                    show_col_types = FALSE) %>%
  dplyr::select(depth, count) %>%
  mutate(depth = as.integer(depth),
         count = as.numeric(count)) %>%
  arrange(depth)

# Expand histogram into per-locus vector (can be big; use weighted approach instead)
# glmmTMB supports weights directly
dat <- hist_df %>% transmute(y = depth, w = count)

m_zinb <- glmmTMB(
  y ~ 1,
  ziformula = ~ 1,
  family = nbinom2(),   # NB2: Var = mu + mu^2/theta  => alpha = 1/theta
  weights = w,
  data = dat
)

# Parameters
zinb_pi <- plogis(unname(fixef(m_zinb)$zi[1]))     # zero-inflation mixing pi
zinb_mu <- exp(unname(fixef(m_zinb)$cond[1]))      # mean mu
zinb_theta <- sigma(m_zinb)                        # theta in nbinom2
zinb_alpha <- 1 / zinb_theta                       # alpha (NB2)

c(pi = zinb_pi, mu = zinb_mu, theta = zinb_theta, alpha = zinb_alpha)

```


```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(stringr)


in_tsv <- "all_fits.zip_zinb.tsv"
out_png <- "coverage_vs_alpha.pretty.png"

# Fix header issue: missing tab between ZINB_alpha and ZINB_loglik
lines <- readLines(in_tsv, warn = FALSE)
stopifnot(length(lines) > 0)
lines[1] <- gsub("ZINB_alphaZINB_loglik", "ZINB_alpha\tZINB_loglik", lines[1], fixed = TRUE)

tmp <- tempfile(fileext = ".tsv")
writeLines(lines, tmp)

df <- read_tsv(tmp, show_col_types = FALSE)
#df <- df %>%
#  mutate(
#    sample = ifelse(stringr::str_detect(sample, "7614"), 
#                    sample, 
#                    paste0(sample, "_7614"))
#  )

df <- df %>%
  mutate(
    sample = ifelse(
      !stringr::str_detect(sample, "7614") & 
      !stringr::str_detect(sample, "6566"),
      paste0(sample, "_7614"),
      sample
    )
  )

df2 <- df %>%
  mutate(
    coverage_pct = as.numeric(coverage_pct),
    ZINB_alpha = as.numeric(ZINB_alpha),
    family = as.character(family),
    sample = as.character(sample)
  ) %>%
  filter(is.finite(coverage_pct), is.finite(ZINB_alpha))
df2 <- df2 %>%
  mutate(
    shape_group = dplyr::case_when(
      stringr::str_detect(sample, "6566") ~ "6566",
      stringr::str_detect(sample, "7614") ~ "7614",
      TRUE ~ "other"
    )
  )


# R-valid colors (hex)
fam_cols <- c(
  illumina = "#1f77b4",  # blue
  nanoseq  = "#ff7f0e",  # orange
  ppmseq   = "#2ca02c",  # green
  udseq    = "#d62728",  # red
  pta      = "#9467bd",  # purple
  hidef    = "#000000",  # black
  other    = "#7f7f7f"   # gray
)

# Auto log-y if alpha dynamic range is large
alpha_ratio <- max(df2$ZINB_alpha, na.rm = TRUE) / max(min(df2$ZINB_alpha, na.rm = TRUE), 1e-12)
use_logy <- is.finite(alpha_ratio) && alpha_ratio > 20

p <- ggplot(df2, aes(x = coverage_pct, y = ZINB_alpha, color = family, shape = shape_group)) +
  geom_point(size = 3, alpha = 0.9) +
  ggrepel::geom_text_repel(
    aes(label = sample),
    size = 3,
    max.overlaps = Inf,
    box.padding = 0.35,
    point.padding = 0.25,
    min.segment.length = 0,
    seed = 1
  ) +
  scale_shape_manual(
    values = c(
      "6566" = 17,   # triangle
      "7614" = 16,   # circle
      "other" = 16
    )
  ) +
  scale_color_manual(values = fam_cols, drop = FALSE) +
  labs(
    x = "Theoretical genomic coverage (%) = non-zero / total",
    y = if (use_logy) "Over-dispersion parameter (log10)" else "Over-dispersion parameter ",
    title = "Coverage vs overdispersion of covered read number"
  ) +
  theme_bw(base_size = 12) +
  theme(
    legend.position = "right",
    panel.grid.minor = element_blank()
  )

if (use_logy) p <- p + scale_y_log10()

#ggsave(out_png, p, width = 8.5, height = 6.5, dpi = 300)
p


```

```{r}
set.seed(42)



rdselect_call300_30x <- purrr::map_dfr(call_cols_30x, function(cc) {

  idx <- which(filt[[cc]] == 1)
  if (length(idx) == 0) return(NULL)

  take_n <- min(300, length(idx))
  sel_idx <- sample(idx, take_n, replace = FALSE)

  filt[sel_idx, ] %>%
    dplyr::mutate(source_call = cc)
}) %>%
  dplyr::distinct(chr, pos, ref, alt, .keep_all = TRUE)


rdselect_call1000 <- purrr::map_dfr(call_cols_full_depth, function(cc) {

  idx <- which(filt[[cc]] == 1)
  if (length(idx) == 0) return(NULL)

  take_n <- min(1000, length(idx))
  sel_idx <- sample(idx, take_n, replace = FALSE)

  filt[sel_idx, ] %>%
    dplyr::mutate(source_call = cc)
}) %>%
  dplyr::distinct(chr, pos, ref, alt, .keep_all = TRUE)




rdselect_bam5000 <- purrr::map_dfr(support_cols, function(cc) {
  
  idx <- which(filt[[cc]] == 1)
  if (length(idx) == 0) return(NULL)
  
  take_n <- min(5000, length(idx))
  sel_idx <- sample(idx, take_n, replace = FALSE)
  
  filt[sel_idx, ] %>%
    dplyr::mutate(source_bam = cc)
})

rdselect_bam5000 <- rdselect_bam5000 %>%
  dplyr::distinct(chr, pos, ref, alt, .keep_all = TRUE)
```


```{r}
library(dplyr)
library(stringr)
library(UpSetR)

# support_cols is given by you (e.g. is_existing_lower_ci_0001_* columns)
support_cols <- exist_cols_full
#support_cols <- exist_cols_200x
#support_cols <- exist_cols_100x
#support_cols <- exist_cols_30x


# ---- Build matrix for UpSetR (missing cols -> 0) ----
make_mat_strata <- function(df, cols) {
  out <- lapply(cols, function(x) {
    if (x %in% colnames(df)) as.integer(df[[x]] %in% 1) else rep(0L, nrow(df))
  })
  mat <- as.data.frame(out, check.names = FALSE)
  colnames(mat) <- cols
  rownames(mat) <- NULL
  mat
}

# ---- Pretty BAM labels from support_cols ----
pretty_plat <- function(x) {
  x %>%
    str_remove("^is_existing_lower_ci_0001_") %>%
    str_remove("^DUPLEX_")
}

# ---- BAM evidence matrix ----

big_tsv <- filt
#big_tsv <- rdselect_call300
#big_tsv <- rdselect_call500
#big_tsv <- rdselect_call1000
#big_tsv <- rdselect_call300_30x
mat_bam <- make_mat_strata(big_tsv, support_cols)

# rename columns to platform labels (ILLUMINA, PPMSEQ, UDSEQ, NANOSEQ, HIDEFSEQ)
colnames(mat_bam) <- pretty_plat(colnames(mat_bam))

# enforce your order (if you want)
#bam_order <- c("ILLUMINA","PPMSEQ","UDSEQ","NANOSEQ","HIDEFSEQ")
#colnames(mat_bam) <- factor(colnames(mat_bam), levels = bam_order) %>% as.character()
#mat_bam <- mat_bam[, bam_order, drop = FALSE]
mat_bam_filt <- mat_bam %>%
  mutate(.k = rowSums(dplyr::select(., everything()) > 0)) %>%  # number of supporting platforms per row
  #filter(.k >= 2) %>%
  dplyr::select(-.k)

# ---- Plot ----
upset(
  mat_bam_filt,
  nsets = ncol(mat_bam_filt),
  nintersects = 200,
  keep.order = TRUE,
  order.by = "freq",
  mainbar.y.label = "Intersection size",
  sets.x.label = "Set size"
)

big_tsv %>% dplyr::select(call_cols_full_depth) %>% colSums(.)

big_tsv %>% dplyr::select(support_cols) %>% dplyr::select(-existing_any_PTA1_16_multiple ) %>% colSums(.)

#rdselect_call1000 %>% write_tsv("rdselect_call1000.tsv")
```


```{r}
library(dplyr)
r_overpois <- function(n, lambda, alpha) {
  shape <- 1 / alpha
  rate  <- 1 / (alpha * lambda)
  
  lam_g <- rgamma(n, shape = shape, rate = rate)
  rpois(n, lam_g)
}
#%>% ggplot(., aes(x=count))+geom_histogram(, binwidth = 10)

k_poi <- rpois(1000, 100) %>% as.data.frame(.) %>% setNames("count") %>% dplyr::mutate(distn="poisson")
k_overpoi_003 <- r_overpois(1000,100,0.03) %>% as.data.frame(.) %>% setNames("count") %>% dplyr::mutate(distn="negbinom")


k_poi_overpoi_003 <- bind_rows(k_poi, k_overpoi_003)

k_poi_overpoi_003 %>% ggplot(., aes(x=count, fill=distn))+geom_histogram(binwidth = 5, position="dodge")


k_poi <- rpois(1000, 100) %>% as.data.frame(.) %>% setNames("count") %>% dplyr::mutate(distn="poisson")

k_overpoi_012 <- r_overpois(1000,100,0.12) %>% as.data.frame(.) %>% setNames("count") %>% dplyr::mutate(distn="negbinom")


k_poi_overpoi_012 <- bind_rows(k_poi, k_overpoi_012)

k_poi_overpoi_012 %>% ggplot(., aes(x=count, fill=distn))+geom_histogram(binwidth = 5, position="dodge")

```


```{r}

library(dplyr)
library(ComplexHeatmap)
library(circlize)
library(stringr)


trin <- read_tsv("trinuc32_all_methods_wide.tsv")

# ---- 1) matrix + normalize (each column sums to 1) ----
mat_raw <- trin %>%
  tibble::column_to_rownames("trinuc") %>%
  as.matrix()

mat_norm <- apply(mat_raw, 2, function(x) x / sum(x))

# ---- 2) cosine similarity + distance(for clustering) ----
cos_sim <- function(m) {
  m_scaled <- apply(m, 2, function(x) x / sqrt(sum(x^2)))
  t(m_scaled) %*% m_scaled
}

sim_mat  <- cos_sim(mat_norm)
dist_mat <- 1 - sim_mat  # cosine distance

# ---- 3) clustering by cosine distance ----
hc <- hclust(as.dist(dist_mat), method = "average")

# ---- 4) annotation (same as you had) ----
samples <- colnames(dist_mat)

tissue <- dplyr::case_when(
  str_detect(samples, "7614") ~ "7614",
  str_detect(samples, "6566_blood") ~ "6566_blood",
  str_detect(samples, "6566_sperm") ~ "6566_sperm",
  str_detect(samples, "(^|_)6566($|_)") ~ "6566_sperm",
  str_detect(samples, "ref_hg38") ~ "reference",
  TRUE ~ "unknown"
)

platform <- dplyr::case_when(
  str_detect(samples, "^ppmseq") ~ "ppmseq",
  str_detect(samples, "^illumina") ~ "illumina",
  str_detect(samples, "^nanoseq") ~ "nanoseq",
  str_detect(samples, "^udseq") ~ "udseq",
  str_detect(samples, "^hidef") ~ "hidef",
  str_detect(samples, "^PTA") ~ "PTA",
  str_detect(samples, "ref_hg38") ~ "reference",
  TRUE ~ "other"
)

ha <- HeatmapAnnotation(
  tissue = tissue,
  platform = platform,
  col = list(
    tissue = c(
      "7614" = "#1b9e77",
      "6566_sperm" = "#d95f02",
      "6566_blood" = "#7570b3",
      "reference" = "grey40",
      "unknown" = "grey80"
    ),
    platform = c(
      ppmseq = "#66c2a5",
      illumina = "#fc8d62",
      nanoseq = "#8da0cb",
      udseq = "#e78ac3",
      hidef = "#a6d854",
      PTA = "#ffd92f",
      reference = "black",
      other = "grey50"
    )
  ),
  annotation_name_side = "left"
)

# ---- 5) log-scale coloring on distance, but legend labels in distance units ----

# Make diagonal NA (so it doesn't dominate; dist=0 => log blow-up)
dist_mat2 <- dist_mat
diag(dist_mat2) <- NA_real_

# z is what we actually plot for colors
z_mat <- -log10(dist_mat2)
diag(z_mat) <- NA_real_

# Robust range from off-diagonal only
z_off <- z_mat[upper.tri(z_mat)]
z_off <- z_off[is.finite(z_off)]
z_lo <- as.numeric(quantile(z_off, 0.05, names = FALSE))
z_hi <- as.numeric(quantile(z_off, 0.995, names = FALSE))

col_fun <- colorRamp2(
  c(z_lo, (z_lo + z_hi) / 2, z_hi),
  c("#440154", "#fdae61", "#ffffbf")
)

# Legend ticks: choose integer z ticks within range (cap to <=5 ticks)
k_min <- max(0, floor(z_lo))
k_max <- max(k_min + 1, ceiling(z_hi))
k_all <- seq(k_min, k_max)

if (length(k_all) > 5) {
  k_at <- unique(round(seq(k_min, k_max, length.out = 5)))
} else {
  k_at <- k_all
}

legend_labels <- ifelse(
  k_at == 0,
  "1",
  paste0("1e-", k_at)
)

alpha <- 0.25
hc2 <- hc
hc2$height <- hc$height^alpha

ht <- Heatmap(
  z_mat,
  name = "Cosine Distance",
  col = col_fun,
  na_col = "white",
  cluster_rows = as.dendrogram(hc2),
  cluster_columns = as.dendrogram(hc2),
  row_dend_width = grid::unit(40, "mm"),        
  column_dend_height = grid::unit(40, "mm"),    
  top_annotation = ha,
  row_names_side = "left",
  column_names_rot = 90,
  heatmap_legend_param = list(
    title = "Cosine Distance",
    at = rev(k_at),
    labels = rev(legend_labels)
  )
)

pdf()
draw(ht)
```





```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
})

df <- tibble::tribble(
  ~denovo,   ~cosmic, ~pct,
  "SBS96A","SBS5",79.12,
  "SBS96A","SBS16",14.82,
  "SBS96A","SBS1",6.06,
  "SBS96B","SBS5",55.00,
  "SBS96B","SBS8",32.48,
  "SBS96B","SBS1",12.52,
  "SBS96C","SBS32",40.22,
  "SBS96C","SBS30",24.46,
  "SBS96C","SBS58",23.50,
  "SBS96C","SBS2",11.82
)

aetiology_map <- c(
  SBS1="Aging",
  SBS5="Aging",
  SBS2="APOBEC",
  SBS8="Unknown",
  SBS16="CWalsh_neuron_TC-NER",
  SBS30="Base_excision_repair",
  SBS32="CWalsh_oligodendrocyte",
  SBS58="Artifact"
)

df <- df %>%
  mutate(
    aetiology = aetiology_map[cosmic],
    cosmic = factor(cosmic,
                    levels=c("SBS1","SBS5","SBS8","SBS16",
                             "SBS30","SBS32","SBS58","SBS2")),
    denovo = factor(denovo,
                    levels=c("SBS96A","SBS96B","SBS96C"))
  )

cols <- c(
  Aging="#2f5d3a",
  APOBEC="#e67e22",
  Unknown="grey70",
  "CWalsh_neuron_TC-NER"="#7b3f8c",
  Base_excision_repair="#d98aa6",
  CWalsh_oligodendrocyte="#5b2a86",
  Artifact="#efe6b5"
)

ggplot(df, aes(cosmic, denovo)) +
  geom_point(aes(size=pct, fill=aetiology),
             shape=21, colour="black", stroke=0.4) +
  geom_text(aes(label=sprintf("%.0f", pct)),
            size=3.2, colour="black") +
  scale_fill_manual(values=cols) +
  scale_size(range=c(4,20)) +
  scale_y_discrete(limits=rev) +
  theme_bw(base_size=13) +
  theme(
    panel.grid.major = element_line(color="grey90"),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle=45, hjust=1),
    legend.position="right",
    legend.title=element_text(face="bold")
  ) +
  labs(x=NULL, y=NULL, fill="Aetiology", size="Contribution (%)")

```

```{r}
mutSig_vcf<-read_tsv("vcf_sig_denovo.tsv")
cosmic_refit_mutSig_vcf<-read_tsv("vcf_sig_cosmic_refit.tsv")
vcf_trinuc <- read_tsv("vcf_trinuc.tsv")
sig_vcf_trinuc <- read_tsv("SBS96_vcf_sig.tsv")


mutSig_bamEvidence<-read_tsv("bamEvidence_sig_denovo.tsv")
cosmic_refit_mutSig_bamEvidence<-read_tsv("bamEvidence_sig_cosmic_refit.tsv")
bamEvidence_trinuc <- read_tsv("bamEvidence_trinuc.tsv")
sig_bamEvidence_trinuc <- read_tsv("SBS96_bamEvidence_sig.tsv")
```

```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(stringr)
  library(tidyr)
  library(ggplot2)
})

# ---- 1) De novo rename map: SBS96A->Sig1 ... SBS96D->Sig4 ----
denovo_map <- c(
  "Signature 96-A" = "Sig1",
  "Signature 96-B" = "Sig2",
  "Signature 96-C" = "Sig3",
  "Signature 96-D" = "Sig4",
  "SBS96A" = "Sig1",
  "SBS96B" = "Sig2",
  "SBS96C" = "Sig3",
  "SBS96D" = "Sig4"
)

# ---- 2) Parse "Global NMF Signatures" into long df: (denovo, cosmic, pct) ----
# cosmic_refit_mutSig_vcf columns:
# `De novo extracted`, `Global NMF Signatures`, ...
df_cosmic_refit_mutSig_vcf <- cosmic_refit_mutSig_vcf %>%
  transmute(
    denovo_raw = `De novo extracted`,
    mix_raw = `Global NMF Signatures`
  ) %>%
  mutate(
    denovo = recode(denovo_raw, !!!denovo_map, .default = denovo_raw),
    # Split by '&'
    mix_raw = str_replace_all(mix_raw, "\\s*&\\s*", "&"),
    parts = str_split(mix_raw, "&", simplify = FALSE)
  ) %>%
  tidyr::unnest_longer(parts) %>%
  mutate(
    parts = str_trim(parts),
    # Extract SBS id + percent in parentheses
    cosmic = str_extract(parts, "SBS\\d+"),
    pct = as.numeric(str_match(parts, "\\(([-0-9.]+)%\\)")[,2])
  ) %>%
  filter(!is.na(cosmic), !is.na(pct)) %>%
  select(denovo, cosmic, pct)

# ---- 3) Plot: bubble, fill by cosmic (pastel), label pct ----
y_levels <- c("Sig1","Sig2","Sig3","Sig4")

cosmic_cols <- c(
  SBS1  = "#A8DADC",  # pastel cyan
  SBS2  = "#F4A261",  # soft orange
  SBS5  = "#8EC9B8",  # mint
  SBS8  = "#C9CCD6",  # light grey-lavender
  SBS16 = "#B8A1D9",  # pastel purple
  SBS31 = "#F6BD60",  # soft yellow-orange
  SBS32 = "#9DB4E0"   # pastel blue
)
x_levels <- cosmic_cols %>% names(.)

p1<-ggplot(
  df_cosmic_refit_mutSig_vcf %>%
    mutate(
      cosmic = factor(cosmic, levels = x_levels),
      denovo = factor(denovo, levels = y_levels)
    )
) +
  geom_point(
    aes(x = cosmic, y = denovo, size = pct, fill = cosmic),
    shape = 21,
    color = "grey20",
    stroke = 0.55
  ) +
  geom_text(
    aes(x = cosmic, y = denovo, label = sprintf("%.1f%%", pct)),
    size = 3.0,
    fontface = "bold",
    lineheight = 0.9
  ) +
  scale_fill_manual(values = cosmic_cols, name = "COSMIC") +
  scale_size(
    range = c(9, 18),
    breaks = c(10, 25, 50, 75),
    name = "Contribution (%)"
  ) +
  scale_y_discrete(limits = rev(levels(factor(y_levels)))) +
  coord_cartesian(clip = "off") +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.major = element_line(color = "grey92", linewidth = 0.5),
    panel.grid.minor = element_blank(),
    axis.title = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    axis.ticks = element_blank(),
    legend.title = element_text(face = "bold"),
    legend.position = "right"
  )
```


```{r}

suppressPackageStartupMessages({
  library(dplyr)
  library(tibble)
  library(ComplexHeatmap)
  library(circlize)
  library(grid)
})

# ---- make matrix: rows = 96 contexts, cols = samples ----
mat_96 <- vcf_trinuc %>%
  as.data.frame() %>%
  column_to_rownames("MutationType") %>%
  as.matrix()

# ---- cosine similarity between samples (columns) ----
cos_sim_cols <- function(m){
  m_scaled <- apply(m, 2, function(x){
    s2 <- sum(x^2)
    if (s2 == 0) return(rep(0, length(x)))
    x / sqrt(s2)
  })
  t(m_scaled) %*% m_scaled
}

sim_96 <- cos_sim_cols(mat_96)

# ---- cosine distance + clustering ----
dist_96 <- 1 - sim_96
dist_96[dist_96 < 0] <- 0
diag(dist_96) <- 0

hc_96 <- hclust(as.dist(dist_96), method = "average")


col_fun <- colorRamp2(
  c(0, 0.5, 0.75, 0.9, 1),
  c("#FFFFFF",
    "#E3EEF9",
    "#9ECAE1",
    "#2171B5",
    "#08306B")
)
ht <- Heatmap(
  sim_96,
  name = "Cosine\nSimilarity",
  col = col_fun,
  cluster_rows = as.dendrogram(hc_96),
  cluster_columns = as.dendrogram(hc_96),
  show_row_dend = FALSE,
  column_dend_height = unit(30, "mm"),
  row_names_side = "left",
  column_names_rot = 45,
  heatmap_legend_param = list(
    at = c(0, 0.5, 0.75, 0.9, 1),
    labels = c("0", "0.5", "0.75", "0.9", "1"),
    title = "Cosine similarity"
  )
)

draw(ht)

```


```{r}
suppressPackageStartupMessages({
  library(ComplexHeatmap)
  library(circlize)
  library(grid)
})

# ---- relative activity (% within each sample) ----
mat_sig_pct <- apply(mat_raw, 2, function(x){
  s <- sum(x)
  if (s == 0) return(rep(0, length(x)))
  100 * x / s
})
mat_sig_pct <- as.matrix(mat_sig_pct)
rownames(mat_sig_pct) <- rownames(mat_raw)   # Sig1-4
colnames(mat_sig_pct) <- colnames(mat_raw)   # Samples

# ---- red-ish palette ----
col_fun_red <- colorRamp2( c(0, 25, 50, 75, 100), c("#FFF5F0", "#FCAE91", "#FB6A4A", "#DE2D26", "#A50F15") )
col_fun_red <- colorRamp2(
  c(0, 25, 50, 75, 100),
  c("#FFFFFF",
    "#FDE0DD",
    "#FCAE91",
    "#FB6A4A",
    "#CB181D")
)

ht_sig <- Heatmap(
  mat_sig_pct,
  name = "Activity\n(%)",
  col = col_fun_red,
  cluster_rows = FALSE,
  cluster_columns = as.dendrogram(hc_96),   # << keep the 96D dendrogram
  show_row_dend = FALSE,
  column_dend_height = unit(30, "mm"),
  row_names_side = "left",
  column_names_rot = 90,
  heatmap_legend_param = list(
    at = c(0, 25, 50, 75, 100),
    title = "Activity (%)"
  ),
  cell_fun = function(j, i, x, y, w, h, fill){
    grid.text(sprintf("%.1f", mat_sig_pct[i, j]), x, y, gp = gpar(fontsize = 8))
  }
)

draw(ht_sig)

```

```{r}
mutSig_bamEvidence<-read_tsv("bamEvidence_sig_denovo.tsv")
cosmic_refit_mutSig_bamEvidence<-read_tsv("bamEvidence_sig_cosmic_refit.tsv")
bamEvidence_trinuc <- read_tsv("bamEvidence_trinuc.tsv")
sig_bamEvidence_trinuc <- read_tsv("SBS96_bamEvidence_sig.tsv")
```


```{r}
## ---- BAM Evidence: bubble (SigA-E vs COSMIC) ----
suppressPackageStartupMessages({
  library(dplyr)
  library(stringr)
  library(tidyr)
  library(ggplot2)
})

# ---- rename map: SBS96A->SigA ... SBS96E->SigE ----
denovo_map_bam <- c(
  "Signature 96-SBS96A" = "SigA",
  "Signature 96-SBS96B" = "SigB",
  "Signature 96-SBS96C" = "SigC",
  "Signature 96-SBS96D" = "SigD",
  "Signature 96-SBS96E" = "SigE",
  "SBS96A" = "SigA",
  "SBS96B" = "SigB",
  "SBS96C" = "SigC",
  "SBS96D" = "SigD",
  "SBS96E" = "SigE"
)

# ---- COSMIC colors (reuse overlaps; add new) ----
cosmic_cols_bam <- c(
  SBS1  = "#A8DADC",
  SBS2  = "#F4A261",
  SBS3  = "#BFD3E6",
  SBS4  = "#C7E9C0",
  SBS5  = "#8EC9B8",
  SBS12 = "#FDD0A2",
  SBS16 = "#B8A1D9",
  SBS31 = "#F6BD60",
  SBS32 = "#9DB4E0",
  SBS51 = "#D9D9D9"

  # new in BAM

)

# ---- parse refit table into long df ----
df_cosmic_refit_mutSig_bamEvidence <- cosmic_refit_mutSig_bamEvidence %>%
  transmute(
    denovo_raw = `De novo extracted`,
    mix_raw = `Global NMF Signatures`
  ) %>%
  mutate(
    denovo = recode(denovo_raw, !!!denovo_map_bam, .default = denovo_raw),
    mix_raw = str_replace_all(mix_raw, "\\s*&\\s*", "&"),
    parts = str_split(mix_raw, "&", simplify = FALSE)
  ) %>%
  tidyr::unnest_longer(parts) %>%
  mutate(
    parts = str_trim(parts),
    cosmic = str_extract(parts, "SBS\\d+"),
    pct = as.numeric(str_match(parts, "\\(([-0-9.]+)%\\)")[,2])
  ) %>%
  filter(!is.na(cosmic), !is.na(pct)) %>%
  select(denovo, cosmic, pct)

y_levels_bam <- c("SigA","SigB","SigC","SigD","SigE")
x_levels_bam <- names(cosmic_cols_bam)

p1_bam <- ggplot(
  df_cosmic_refit_mutSig_bamEvidence %>%
    mutate(
      cosmic = factor(cosmic, levels = x_levels_bam),
      denovo = factor(denovo, levels = y_levels_bam)
    )
) +
  geom_point(
    aes(x = cosmic, y = denovo, size = pct, fill = cosmic),
    shape = 21,
    color = "grey20",
    stroke = 0.55
  ) +
  geom_text(
    aes(x = cosmic, y = denovo, label = sprintf("%.1f%%", pct)),
    size = 3.0,
    fontface = "bold",
    lineheight = 0.9
  ) +
  scale_fill_manual(values = cosmic_cols_bam, name = "COSMIC") +
  scale_size(
    range = c(9, 18),
    breaks = c(10, 25, 50, 75),
    name = "Contribution (%)"
  ) +
  scale_y_discrete(limits = rev(levels(factor(y_levels_bam)))) +
  coord_cartesian(clip = "off") +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.major = element_line(color = "grey92", linewidth = 0.5),
    panel.grid.minor = element_blank(),
    axis.title = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    axis.ticks = element_blank(),
    legend.title = element_text(face = "bold"),
    legend.position = "right"
  )

p1_bam

```

```{r}
## ---- BAM Evidence: sample-sample cosine similarity heatmap (96D) ----
suppressPackageStartupMessages({
  library(dplyr)
  library(tibble)
  library(ComplexHeatmap)
  library(circlize)
  library(grid)
})

# ---- matrix: rows = 96 contexts, cols = samples ----
mat_96_bam <- bamEvidence_trinuc %>%
  as.data.frame() %>%
  column_to_rownames("MutationType") %>%
  as.matrix()

# ---- cosine similarity between samples (columns) ----
cos_sim_cols <- function(m){
  m_scaled <- apply(m, 2, function(x){
    s2 <- sum(x^2)
    if (s2 == 0) return(rep(0, length(x)))
    x / sqrt(s2)
  })
  t(m_scaled) %*% m_scaled
}

sim_96_bam <- cos_sim_cols(mat_96_bam)

# ---- cosine distance + clustering ----
dist_96_bam <- 1 - sim_96_bam
dist_96_bam[dist_96_bam < 0] <- 0
diag(dist_96_bam) <- 0
hc_96_bam <- hclust(as.dist(dist_96_bam), method = "average")

# ---- color (0-1) ----
col_fun_blue <- colorRamp2(
  c(0, 0.5, 0.75, 0.9, 1),
  c("#FFFFFF", "#E3EEF9", "#9ECAE1", "#2171B5", "#08306B")
)

ht_bam <- Heatmap(
  sim_96_bam,
  name = "Cosine\nSimilarity",
  col = col_fun_blue,
  cluster_rows = as.dendrogram(hc_96_bam),
  cluster_columns = as.dendrogram(hc_96_bam),
  show_row_dend = FALSE,
  column_dend_height = unit(30, "mm"),
  row_names_side = "left",
  column_names_rot = 45,
  heatmap_legend_param = list(
    at = c(0, 0.5, 0.75, 0.9, 1),
    labels = c("0", "0.5", "0.75", "0.9", "1"),
    title = "Cosine similarity"
  )
)

draw(ht_bam)
```

```{r}
## ---- BAM Evidence: sig activity heatmap (SigA-E x Samples), dendrogram = hc_96_bam ----
suppressPackageStartupMessages({
  library(dplyr)
  library(tibble)
  library(ComplexHeatmap)
  library(circlize)
  library(grid)
})

# ---- matrix: rows = sig, cols = sample ----
mat_raw_bam <- mutSig_bamEvidence %>%
  as.data.frame() %>%
  column_to_rownames("Samples") %>%
  t()

rownames(mat_raw_bam) <- dplyr::recode(rownames(mat_raw_bam), !!!denovo_map_bam, .default = rownames(mat_raw_bam))

# ---- relative activity (% within each sample) ----
mat_sig_pct_bam <- apply(mat_raw_bam, 2, function(x){
  s <- sum(x)
  if (s == 0) return(rep(0, length(x)))
  100 * x / s
})
mat_sig_pct_bam <- as.matrix(mat_sig_pct_bam)
rownames(mat_sig_pct_bam) <- rownames(mat_raw_bam)  # SigA-E
colnames(mat_sig_pct_bam) <- colnames(mat_raw_bam)  # Samples

# ---- red-ish but not too aggressive ----
col_fun_red <- colorRamp2(
  c(0, 25, 50, 75, 100),
  c("#FFFFFF", "#FDE0DD", "#FCAE91", "#FB6A4A", "#CB181D")
)

ht_sig_bam <- Heatmap(
  mat_sig_pct_bam,
  name = "Activity\n(%)",
  col = col_fun_red,
  cluster_rows = FALSE,
  cluster_columns = as.dendrogram(hc_96_bam),
  show_row_dend = FALSE,
  column_dend_height = unit(30, "mm"),
  row_names_side = "left",
  column_names_rot = 90,
  heatmap_legend_param = list(
    at = c(0, 25, 50, 75, 100),
    title = "Activity (%)"
  ),
  cell_fun = function(j, i, x, y, w, h, fill){
    grid.text(sprintf("%.1f", mat_sig_pct_bam[i, j]), x, y, gp = gpar(fontsize = 8))
  }
)

draw(ht_sig_bam)
```